### 2023-08-28 paper.tex
First draft

\documentclass{article}
\pdfpagewidth=8.5in
\pdfpageheight=11in

\usepackage{ijcai23}

\usepackage{times}
\usepackage{soul}
\usepackage{url}
\usepackage[hidelinks]{hyperref}
\usepackage[utf8]{inputenc}
\usepackage[small]{caption}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{booktabs}
\usepackage[switch]{lineno}

\usepackage{listings}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage[capitalize,noabbrev]{cleveref}
\usepackage{microtype}

\lstset{breaklines=true}
\lstset{language=C++,
        basicstyle=\ttfamily,
        stringstyle=\color{red},
        commentstyle=\color{green},
        breaklines=true,
        showstringspaces=false}

% Comment out this line in the camera-ready submission
\linenumbers

\urlstyle{same}

% the following package is optional:
%\usepackage{latexsym}

% See https://www.overleaf.com/learn/latex/theorems_and_proofs
% for a nice explanation of how to define new theorems, but keep
% in mind that the amsthm package is already included in this
% template and that you must *not* alter the styling.
\newtheorem{example}{Example}
\newtheorem{theorem}{Theorem}

% PDF Info Is REQUIRED.
% Please **do not** include Title and Author information
\pdfinfo{
/TemplateVersion (IJCAI.2023.0)
}

\title{Towards Practical First-Order Model Counting}

% \author{
% First Author$^1$
% \and
% Second Author$^2$\and
% Third Author$^{2,3}$\And
% Fourth Author$^4$
% \affiliations
% $^1$First Affiliation\\
% $^2$Second Affiliation\\
% $^3$Third Affiliation\\
% $^4$Fourth Affiliation
% \emails
% \{first, second\}@example.com,
% third@other.example.com,
% fourth@example.com
% }

\begin{document}

\maketitle

% <= 200 words
\begin{abstract}
\end{abstract}

\section{Introduction}

\section{Methodology}

\begin{itemize}
  \item Added support for infinite precision integers using the \texttt{gmp}
        library to the numerical evaluation part. Now, all functions take
        arguments (domain sizes) as unsigned integers and return
        \texttt{mpz\_class} objects (defined in the \texttt{gmpxx.h} header).
  \item Added a CLI option \texttt{--sizes} to specify the domain sizes on the
        command line.
  \item Converting the recursive equations into a \texttt{C++} program, which
        can then be compiled and executed to obtain numerical values.
  \item The translation of a set of equations($E$) into \texttt{C++} works as
        follows.
        \begin{enumerate}
          \item First, we create a cache for each function in $E$. For a
                function with name \texttt{func}, this cache is called
                \texttt{func\_cache}. This is implemented as a multi-dimensional
                vector containing objects of \texttt{class cahce\_elem} defined
                as shown in the example code. The default initialization of this
                object is to $-1$ which is useful for recognizing unevaluated
                cases.
          \item Next, we create a function definition for each lhs in $E$,
                including all functions and base cases. The signatures of these
                functions is decided as follows - a function call containing
                only variable arguments is named as the function itself, and
                ones with constants in their arguments are suffixed with a
                string that contains \texttt{'x'} at the $i$th place if the
                $i$th argument is variable and the $i$th argument if that
                argument is a constant. For example, $f(x1, x2, x3)$ is declared
                as \texttt{int f(int x1, int x2, int x3);} and $f(1, x2, x3)$ is
                declared as \texttt{int f\_1xx(int x2, int x3);} (the constant
                arguments are removed from the signature).
          \item The rhs of each equation in $E$ is used to define the body of
                the equation corresponding to the lhs of that equation. The
                function body (for a function \texttt{func} corresponding to
                equation $e$) is formed as follows-
                \begin{enumerate}
                  \item First, we check if the evaluation is already present in
                        the cache. If so, then we return the cache element. The
                        cache accesses are done using the \texttt{get\_elem}
                        function(definition given in the example), which resizes
                        the cache if the accessed index is out of range.
                  \item If the element is absent, then we decide if the
                        arguments corresponding to $e$ or one of the functions
                        corresponding to the base cases, based on the value of
                        the arguments. If it corresponds to the base cases, then
                        we directly call the base case function and return its
                        value. Else, we evaluate the value using the rhs, store
                        the evaluated value in the cache and return the
                        evaluated value. Note that in this step, we only call
                        the base case function with one more constant argument
                        that \texttt{func}. For example, \texttt{f0(x, y)} would
                        call \texttt{f0\_0x(y)} if \texttt{x == 0} and
                        \texttt{f0\_x0(x)} if \texttt{y == 0}.
                  \item In order to translate the rhs, we do the following
                        changes-
                        \begin{itemize}
                          \item \verb|a^b| is converted to \texttt{power(a,b)}.
                          \item \texttt{Sum[exp, x, a, b]} = ($\sum_{x=a}^{b}exp$) is converted to -\\
                                \begin{lstlisting}
([y,z,...](){
    int sum{0};
    for(int x{a}; x <= b; x++)
        sum += exp;
    return sum;
})()
\end{lstlisting}
                                where y, z, ... are the free variables present
                                in \texttt{exp}.
                        \end{itemize}
                \end{enumerate}
        \end{enumerate}
  \item Implemented the evaluation of basecases. The evaluation is done by
        simplifying the clauses and then using CRANE to find the basecases.
        First, in the \texttt{SimplifyUsingWolfram} class, while traversing the
        graph to find the equations, we store 2 \texttt{Map} objects -
        \texttt{clause\_func\_map}(which stores the mapping from the function
        names to the formulae, whose model count they represent) and
        \texttt{var\_domain\_map}(which stores the mapping from the variable
        names to the domains whose sizes they represent). Then, a particular
        domain is selected(using the algorithm described in previous reports)
        and the clauses are simplified. Then, CRANE is called on those clauses
        to evaluate the base cases using \texttt{WeightedCNF.SimplifyInWolfram}.
        After that, we change the function names and variable to make it
        consistent with the previous domain to variable mapping, and append
        these basecases to the set of equations.
  \item Ideas for finding which base cases are needed. The goal is to find the
        sufficient set of base cases. We know that if say, on the rhs of all
        equations, the domain size appears as $m-c_1, m-c_2, \dots, m-c_k$, then
        finding $f(0, x_1, x_2, \dots)$, $f(1, x_1, x_2, \dots)$,
        $\dots f(m_0, x_1, x_2, \dots)$ for every function $f$, where
        $m_0 = \textnormal{max}(c_1, c_2, \dots c_k) - 1$ forms a sufficient set
        of base cases. Hence, in order to do the same efficiently, we can take
        that domain for which $m_0$ is the minimum, i.e.
        $argmin(\textnormal{max}(c_1, c_2, \dots c_k))$.
  \item If a domain is set to $0$, the model count can be found by simplifying
        the universal and existential quantifiers over that domain, and then
        using crane. However, if a domain is set to contain some non-empty
        number of elements, we can apply GDR over that domain and proceed with
        crane, substituting the known values of preveiously calculated model
        counts in the process.
  \item Implemented a function to simplify functions (recursive and
        non-recursive) using the Wolfram Engine. Modified the wolfram
        simplification functions to incorporate binomial and indicator
        functions.
\end{itemize}

\subsection{Foo}

The target is to generate C++ code that can evaluate numerical values of the
model counts based on the equations generated by \textsc{Crane}. There are two
ways to do the same.
\begin{enumerate}
  \item Generate C++ code by traversing the FCG, similar to what is done in
        \texttt{OutputVisitor.scala}.
  \item Parse the equations generated by \textsc{Crane} after simplifying in
        wolfram and then generate C++ code.
\end{enumerate}

The problem with the first approach is that while generating base cases, the
subsequent calls for \textsc{Crane} do not necessarily have the same meanings
for the function arguments and the functions. For example, if \texttt{f1[x0, x1,
  x2]} represents the model count of a constrained formula F, where f1 is an
auxilliary formula and $x0 = |A|, x1 = |B|, x2 = |C|$, ($A,B,C$ are domains) and
we want to evaluate \texttt{f1[0, x1, x2]}($|A| = 0$), then \textsc{Crane} may
return the required model count as \texttt{f0[x0, x1]}, where
$x0 = |B|, x1 = |C|$. We'll need to translate this to \texttt{f1[0,x1,x2]},
which has already been implemented in \texttt{Basecases.scala}. Also, the second
approach can be done in linear time in the length of the formula using
the
\href{https://en.wikipedia.org/wiki/Shunting_yard_algorithm#:~:text=In%20computer%20science%2C%20the%20shunting,abstract%20syntax%20tree%20(AST).}{Shunting
  Yard Algorithm}. Hence, we are implementing the second approach (have already
implemented the parser).

\subsection{Bar}

The previous method of base case evaluation on setting a domain to size 0 or 1
had the following error - in case a domain size was set to 0, it assumed that
those predicates which were deleted from the clauses could take any truth value
over the entire domain and the rest were fully covered by the remaining clauses.

For example, consider the constrained CNF formula -
\begin{align}
  \forall x \in \Delta \forall y, z \in \Gamma &: P(x) \lor Q(y, z) \\
  \forall y \in \Gamma^\top &: Q(y, z)
\end{align}
In this case, if we set $|\Delta|$ to $0$, then based on the previous idea, the
new set of clauses would be
\begin{equation}
  \forall y, z \in \Gamma^\top : Q(y, z)
\end{equation}
and the set of removed predicates is $\phi$, and hence the model count returned
would be $1$. However, the actual model count shoud be
$2^{|\Gamma|^2 - |\Gamma^\top|^2}$.

There are 2 ideas which we could think of to solve this -
\begin{enumerate}
  \item \textit{(Currently, we have implemented this)} Convert the clauses
        having universal quantification over the null domain to tautology
        clauses, hence retaining all the predicates which don't have an argument
        belonging to the null domain. For example, we would convert the above
        mentioned CNF formula to
        \begin{align}
          \forall y, z \in \Gamma &: Q(y, z) \lor \neg Q(y, z) \\
          \forall y, z \in \Gamma^\top &: Q(y, z)
        \end{align}
        The model count returned by this will also consider the truth value of Q
        over $y \not \in \Gamma^\top$ or $z \not \in \Gamma^\top$.
  \item We find the domain size over which the predicate has not been defined by
        the clauses and multiply the required factor. For example, in the above
        example, the multiplier would be $2^{|\Gamma|^2 - |\Gamma^\top|^2}$ and
        the simplified set of clauses would be
        \begin{equation}
          \forall y, z \in \Gamma^\top : Q(y, z)
        \end{equation}
        which has a model count of $1$. Hence, the obtained value of model count
        would be $2^{|\Gamma|^2 - |\Gamma^\top|^2} \times 1$.
\end{enumerate}

\section{Algorithms}

\subsection{Algorithm to Find a Sufficient Set of Base Cases}

\begin{algorithm}[t]
  \caption{Algorithm For Sufficient Base Cases}\label{alg1}
  \KwIn{dependencies}
  \KwOut{base\_cases}

        % Algorithm steps
        base\_cases $\leftarrow$ Set()

        \ForEach{dependency $\in$ dependencies}{
          \ForEach{elem $\in$ dependency.value}{
            \If{dependency.key.func\_name = elem.func\_name}{
              \ForEach{arg $\in$ elem.args}{
                \tcp{eg : lim is 2 for the arg (x-3)}
                lim $\leftarrow$ arg.const\_subtracted - 1 \\
                \For{l $\in$ 0 to lim}{
                  \tcp{eg : for dependency.key = f(x,y), arg = x and l = 0, add f(0, y) to base\_cases}
                  base\_cases += dependency.key.substitute\_arg(arg\_var, l)
                }
              }
            }
            \Else{
              \ForEach{arg $\in$ elem.args}{
                \tcp{eg : lim is 2 for the arg (x-3)}
                lim $\leftarrow$ arg.const\_subtracted - 1 \\
                \For{l $\in$ 0 to lim}{
                  \tcp{eg : for dependency.key = f(x,y), arg = x and l = 0, add f(0, y) to base\_cases}
                  base\_cases += dependency.key.substitute\_arg(arg, l)\\
                  \tcp{signature for f(x-1, y-2) is f(x,y)}
                  base\_cases += elem.signature.substitute\_arg(arg, l)
                }
              }
            }
          }
        }
\end{algorithm}

The pseudo-code of the algorithm is given as \cref{alg1}. Here,
\texttt{dependencies} maps the function call on the lhs of each equation to the
set of function calls on the rhs of the equation. For example, for the equations
\begin{equation*}
  f(m, n) = g(m-1, n) + f(m-2, n-1)
\end{equation*}
\begin{equation*}
  g(m, n) = f(m-1, n-2) + g(m-1, n-1)
\end{equation*}
the dependencies are
\begin{equation*}
  f(m,n) \mapsto \{g(m-1,n), f(m-2, n-1)\}
\end{equation*}
\begin{equation*}
  g(m, n) \mapsto \{f(m-1, n-2), g(m-1, n-1)\}
\end{equation*}

\subsection{Algorithm to Transform CNF Based on Domain Sizes}

\begin{algorithm}[t]
  \caption{Algorithm For Transforming CNF Based on Domain Sizes}\label{alg2}
  \KwIn{cnf, dom, dom\_size}
  \KwOut{transformed\_cnf, multiplier}
  removed\_predicates $\leftarrow$ Set()
  retained\_predicates $\leftarrow$ Set()
  \If{dom\_size = 0}{
    \ForEach{clause $\in$ cnf}{
      \If{dom $\in$ clause.domains}{
        removed\_predicates += clause.predicates
      }
      \Else{
        retained\_predicates += clause.predicates
        transformed\_cnf += clause
      }
    }
    removed\_predicates = removed\_predicates $\setminus$ retained\_predicates
    \ForEach{pred $\in$ removed\_predicates}{
      multiplier *= $2^{(pred.domain\_sizes.join("*"))}$
    }
  }
  \ElseIf{dom\_size = 1}{
    const $\leftarrow$ newConst()
    \ForEach{clause $\in$ cnf}{
      transformed\_clause $\leftarrow$ clause
      \ForEach{v $\in$ clause.vars | v.domain = dom}{
        transformed\_cnf $\leftarrow$ transformed\_clause.substituteVar(v, const)
      }
      transformed\_cnf += transformed\_clause
    }
  }
\end{algorithm}

The pseudo code is given as \cref{alg2}. We use \cref{alg2} to find the
transformed CNF corresponding to each base case obtained using \cref{alg1} and
call \textsc{Crane} on the CNF to obtain the required base cases.

\subsection{Baz}

The following steps were followed while finding the basecases-
\begin{enumerate}
  \item Expand the summations in each equation - Here we expand the summations
        of the form:
        $\sum_{x=0}^{x1}<\textnormal{something}> \cdot [a \le x < b]$, or
        similar inequalities where x is bounded by constants and $a$ and $b$ are
        constants, by substituting the value of x from $a$ to $b-1$. For
        example, we replace
        $\sum_{x=0}^{x_1} {{x_1}\choose{x}} f(x_1-x)\cdot[0 \le x < 2]$ by
        ${{x_1}\choose{0}}f(x_1) + {{x_1}\choose{1}}f(x_1-1)$.
  \item Next, we find the dependencies of those functions that appear on the rhs
        of any equation. Consider, for example the following set of equations-
        \begin{align}
          f0(m, n) &= f1(m-1, n) + f2(m, n-1)\\
          f1(m, n) &= f1(m-1, n-1) \times f2(m-2, n-1)\\
          f2(m, n) &= 2 \times f1(m-3, n-1)
        \end{align}
        In this case, the dependencies computed are
        \begin{align*}
          f1(m, n) &\rightarrow f1(m-1, n-1), f2(m, n-1)\\
          f2(m, n) &\rightarrow f1(m-3, n-1)
        \end{align*}
  \item Now, based on idea II of the previous report, we find a domain that has
        only terms of the form $x-1$ appearing on the rhs of the dependencies.
        The base cases are then calculated by setting this domain size to 0. For
        the above example, $n$ is the selected domain, and not $m$ since there
        are $m-2$ and $m-3$ terms appearing in the arguments.
\end{enumerate}

\paragraph{Limitations of the current implementation.} Ideally, we should
calculate the base cases based on idea II of the previous report by finding the
basecases upto max$(c1, c2, ...) - 1$. However, we are yet to implement code for
finding the basecases for the non-null domain case.

\bibliographystyle{named}
\bibliography{paper}

\end{document}