### 2024-04-25 doc/paper.tex
The rest of Section 2

\documentclass{article}
\pdfpagewidth=8.5in
\pdfpageheight=11in

\usepackage{kr}

\usepackage{times}
\usepackage{soul}
\usepackage{url}
\usepackage[hidelinks]{hyperref}
\usepackage[utf8]{inputenc}
\usepackage[small]{caption}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{booktabs}
\urlstyle{same}

\usepackage[backgroundcolor=lightgray]{todonotes} % TODO: remove
\usepackage{listings}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage[capitalize,noabbrev]{cleveref}
\usepackage{microtype}
\usepackage{mathtools}
\usepackage{forest}
\usepackage{siunitx}
\usepackage{amsfonts}
\usepackage{pifont}
\usepackage[inline]{enumitem}

\usetikzlibrary{arrows.meta}

\newtheorem{theorem}{Theorem}
\newtheorem{fact}{Fact}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}

\pdfinfo{
/TemplateVersion (KR.2022.0, KR.2023.0, KR.2024.0)
}

\forestset{
  sn edges/.style={for tree={edge={-Latex}}}
}

\lstset{breaklines=true}
\lstset{language=C++,
        basicstyle=\ttfamily,
        stringstyle=\color{red},
        commentstyle=\color{green},
        breaklines=true,
        showstringspaces=false}

\newcommand{\crefrangeconjunction}{--}

\crefname{line}{line}{lines}

\crefalias{enumi}{type}
\crefname{type}{Type}{Types}
\creflabelformat{type}{#2\textup{#1}#3}

\crefalias{clause}{equation}
\crefname{clause}{Clause}{Clauses}
\creflabelformat{clause}{#2\textup{(#1)}#3}

\crefalias{formula}{equation}
\crefname{formula}{Formula}{Formulas}
\creflabelformat{formula}{#2\textup{(#1)}#3}

\DeclareMathOperator{\CR}{CR}
\DeclareMathOperator{\DR}{DR}
\DeclareMathOperator{\Reff}{Ref}

\DeclareMathOperator{\Doms}{Doms}

\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\ding{55}}

\newcommand{\expr}{\mathtt{expr}}
\newcommand{\Ctwo}{$\mathsf{C}^{2}$}
\newcommand{\FO}{$\mathsf{FO}$}
\newcommand{\UFO}{$\mathsf{UFO}^{2} + \mathsf{EQ}$}

\newcommand{\Cranetwo}{\textsc{Crane2}}
\newcommand{\Cranebfs}{\textsc{Crane2-BFS}}
\newcommand{\Cranegreedy}{\textsc{Crane2-Greedy}}

\SetKwFunction{CompileWithBaseCases}{CompileWithBaseCases}
\SetKwFunction{Crane}{Compile}
\SetKwFunction{Propagate}{Propagate}
\SetKwFunction{FindBaseCases}{FindBaseCases}
\SetKwFunction{Simplify}{Simplify}

\title{Towards Practical First-Order Model Counting}

\author{%
  Ananth K. Kidambi$^1$\footnote{The first and second authors contributed equally and were affiliated with the National University of Singapore during the completion of this work.}\and
Guramrit Singh$^1$\and
Paulius Dilkas$^2$\and
Kuldeep S. Meel$^3$ \\
\affiliations
$^1$Indian Institute of Technology Bombay, Mumbai, India\\
$^2$National University of Singapore, Singapore, Singapore\\
$^3$University of Toronto, Toronto, Canada\\
\emails
\{210051002, 210050061\}@iitb.ac.in,
paulius.dilkas@nus.edu.sg,
meel@cs.toronto.edu
}

\begin{document}

\maketitle

\begin{abstract}
  First-order model counting (FOMC) is the problem of counting the number of
  models of a sentence in first-order logic. Recently, a new algorithm for FOMC
  was proposed that, instead of simply providing the final count, generates
  definitions of (possibly recursive) functions, which can be evaluated with
  different arguments to compute the model count for any domain size. However,
  the algorithm did not include base cases in the recursive definitions. This
  work makes three contributions. First, we demonstrate how to construct
  function definitions that include base cases by modifying the logical formulas
  used in the FOMC algorithm. Second, we extend the well-known circuit
  modification technique in knowledge compilation, known as smoothing, to work
  with the formulas corresponding to base cases and the recently proposed node
  types, such as domain recursion and constraint removal. Third, we introduce a
  compilation algorithm that transforms the function definitions into C++ code,
  equipped with arbitrary-precision arithmetic. These additions allow the new
  FOMC algorithm to scale to domain sizes over \num{9000} times larger than the
  current state of the art, as demonstrated through experimental results.
\end{abstract}

\section{Introduction (TODO)}

\todo[inline,caption={}]{
  \begin{itemize}
    \item 9 pages including appendices but not acknowledgements
    \item Add some papers mentioned in:
          \begin{itemize}
            \item very recent work
            \item my previous paper, including:
                  \begin{itemize}
                    \item other liftable fragments
                    \item some more theory papers, e.g., LICS 2018
                  \end{itemize}
          \end{itemize}
    \item introduce \Cranetwo{}
    \item (maybe) introduce lifted/liftable
  \end{itemize}
}

\paragraph{Papers To Cite}
\begin{itemize}
  \item overviews
        \begin{itemize}
          \item lifted probabilistic inference~\cite{DBLP:conf/ecai/Kersting12}
          \item recent overview paper~\cite{DBLP:conf/ijcai/Kuzelka23}
        \end{itemize}
  \item Alternative definition~\cite{DBLP:journals/cacm/GogateD16}
  \item relevant theoretical work~\cite{DBLP:conf/aaai/MalhotraS22}
  \item original domain recursion~\cite{DBLP:conf/nips/Broeck11}
  \item algorithms
        \begin{itemize}
          \item \textsc{ForcLift}~\cite{DBLP:conf/ijcai/BroeckTMDR11}
          \item \textsc{Crane}~\cite{DBLP:conf/kr/DilkasB23}
          \item \textsc{FastWFOMC}~\cite{DBLP:conf/uai/BremenK21}
          \item \textsc{L2C}~\cite{DBLP:conf/kr/KazemiP16} (similarly to us
                compiles to C++ code, but (probably) doesn't work on as many
                formulas)
          \item approximate~\cite{DBLP:conf/ijcai/BremenK20}
          \item for Markov logic networks~\cite{DBLP:journals/ml/RichardsonD06}
                \begin{itemize}
                  \item \textsc{Magician}~\cite{DBLP:conf/aaai/VenugopalSG15}
                  \item \textsc{Tuffy}~\cite{DBLP:journals/pvldb/NiuRDS11}
                  \item \textsc{Alchemy}~\cite{DBLP:journals/cacm/GogateD16}
                        (same as the alternative definition)
                \end{itemize}
        \end{itemize}
  \item complexity
        \begin{itemize}
          \item liftability~\cite{DBLP:conf/starai/JaegerB12}
          \item hardness for three variables~\cite{DBLP:conf/pods/BeameBGS15}
          \item liftable fragments
                \begin{itemize}
                  \item \Ctwo~\cite{DBLP:journals/jair/Kuzelka21}
                  \item tree axioms~\cite{DBLP:journals/ai/BremenK23}
                  \item linear order axioms~\cite{DBLP:conf/aaai/TothK23}
                  \item some liftable fragments~\cite{DBLP:conf/nips/KazemiKBP16}
                \end{itemize}
        \end{itemize}
  \item applications
        \begin{itemize}
          \item extensions to
                sampling~\cite{DBLP:conf/aaai/WangB0K22,DBLP:conf/lics/WangP0K23}
          \item discovery of combinatorial
                sequences~\cite{DBLP:conf/ijcai/SvatosJT0K23}
          \item conjecturing recurrence
                relations~\cite{DBLP:conf/ilp/BarvinekB0ZK21}
          \item probabilistic logic
                programming~\cite{DBLP:journals/ijar/RiguzziBZCL17} (WFOMC was
                shown to be supreme)
          \item probabilistic databases~\cite{DBLP:journals/debu/GribkoffSB14}
        \end{itemize}
  \item lifted inference elsewhere
        \begin{itemize}
          \item constraint satisfaction~\cite{DBLP:journals/jair/TotisDRK23}
          \item answer set programming~\cite{DBLP:journals/ijar/AzzoliniR23}
        \end{itemize}
\end{itemize}

\begin{figure*}[t]
  \centering
  \begin{tikzpicture}
    \node at (-1, 0) (formula) {$\phi$};
    \node[draw,rounded rectangle] at (3, 0) (compilewithbasecases) {\CompileWithBaseCases};
    \node[draw,rounded rectangle] at (9, 0) (compilation) {Compile to C++};

    \node[draw,rounded rectangle,dashed] at (12, 0) (cpp) {C++ code};
    \node at (12, -1) (sizes) {Domain sizes};

    \node at (15, 0) (count) {Model count};

    \node[draw,rounded rectangle] at (3, -2) (findbasecases) {\FindBaseCases};
    \node[draw,rounded rectangle,left = 0.1cm of findbasecases] (crane) {\Crane};
    \node[draw,rounded rectangle,right = 0.1cm of findbasecases] (propagate) {\Propagate};
    \node[draw,rounded rectangle,right = 0.1cm of propagate] (simplify) {\Simplify};

    \node[draw,fit={(compilewithbasecases) (compilation) (crane) (findbasecases) (propagate)},inner ysep=7pt,yshift=5pt] {};
    \node at (0.2, 0.5) {\Cranetwo};

    \draw[-Latex] (formula) -- (compilewithbasecases);
    \draw[-Latex] (compilewithbasecases) -- node[above] {$\mathcal{E}$} (compilation);
    \draw[-Latex] (compilation) -- (cpp);
    \draw[-Latex] (sizes) -- (cpp);
    \draw[-Latex] (cpp) -- (count);

    \draw[-Latex,dashed] (compilewithbasecases) -- node[midway,left] {uses} (crane);
    \draw[-Latex,dashed] (compilewithbasecases) -- node[midway,left] {uses} (findbasecases);
    \draw[-Latex,dashed] (compilewithbasecases) -- node[midway,left] {uses} (propagate);
    \draw[-Latex,dashed] (compilewithbasecases) -- node[midway,right] {uses} (simplify);
  \end{tikzpicture}
  \caption[]{The outline of using \Cranetwo{} to compute the model count of a
    formula $\phi$. The formula is compiled into a set of equations
    $\mathcal{E}$ that are then compiled to a C++ program. This program can then
    be run with different command line arguments to compute the model count of
    $\phi$ for various domain sizes. \CompileWithBaseCases makes use
    of: \begin{enumerate*}[label=(\roman*)]
      \item the knowledge compilation algorithm of \textsc{Crane} (denoted by \Crane),
      \item a procedure for identifying a sufficient set of base cases (denoted by \FindBaseCases),
      \item a procedure for constructing a formula that corresponds to a given base case (denoted by \Propagate), and
      \item algebraic simplification techniques (denoted by \Simplify).
    \end{enumerate*}
  }\label{fig:overview}
\end{figure*}

\todo[inline,caption={}]{
  \begin{itemize}
    \item By pointing to the outline in \cref{fig:overview}, make forward
          references to the sections of the paper.
    \item Somewhere: mention that we're primarily concerned with computing
          unweighted model counts, but everything trivially extends to the
          weighted setting as well.
    \item Introduce greedy search and breadth-first search (BFS), together with
          \Cranegreedy{} and \Cranebfs{}. The knowledge compilation part of both
          \textsc{Crane} and \Cranetwo{} can be executed using either greedy
          (similar to \textsc{ForcLift}) or breadth-first search. We use both in
          our experiments, denoting them as \Cranegreedy{} and \Cranebfs{},
          respectively.
    \item Introduce the WFOMC acronym.
    \item Introduce first-order knowledge compilation as a class of WFOMC
          algorithms, defining/mentioning compilation rules, FCG, etc.
  \end{itemize}
}

\paragraph{Contributions}
\begin{itemize}
  \item Completing the definitions of recursive functions by:
  \begin{itemize}
    \item identifying a sufficient set of base cases (\cref{sec:identifying})
    \item constructing formulas that correspond to these base cases
          (\cref{sec:simplifying})
    \item and recursing on these subproblems
  \end{itemize}
  \item Compiling these function definitions into a C++ program that can be
        executed independently for any domain size values \cref{sec:cpp}
        \begin{itemize}
          \item including support for infinite precision arithmetic via GNU
                Multiple Precision Arithmetic Library
        \end{itemize}
  \item Experiments comparing \Cranetwo{} with the main alternative approach
        demonstrate the ability of \Cranetwo{} to scale to domain sizes\dots
\end{itemize}

\Cref{sec:identifying,sec:cpp} deal with algebraic constructs whereas
\cref{sec:simplifying} deals with logic.

\section{Preliminaries}

In \cref{sec:logic}, we summarise the fundamentals of first-order logic. Then,
in \cref{sec:threelogics}, we formally define WFOMC and discuss the differences
among three variants of first-order logic used for WFOMC.\@ Finally, in
\cref{sec:algebra}, we introduce the terminology used to describe the output of
the original \textsc{Crane} algorithm, i.e., functions and equations that define
them.

We write $\mathbb{N}_{0}$ for the set of non-negative integers. In the context
of both algebra and logic, we write $S\sigma$ for a \emph{substitution} $\sigma$
applied to an expression $S$, where
$\sigma = [x_{1} \mapsto y_{1}, x_{2} \mapsto y_{2}, \dots, x_{n} \mapsto y_{n}]$
denotes the replacement of all occurrences of $x_{i}$ with $y_{i}$ for all
$i = 1, \dots, n$.

\subsection{First-Order Logic}\label{sec:logic}

In this section, we review the basic concepts of first-order logic as they are
used in first-order knowledge compilation algorithms. There are two key
differences between the logic used within such algorithms and the logic
supported as input. First, Skolemization~\cite{DBLP:conf/kr/BroeckMD14} is used
to eliminate existential quantifiers by introducing additional predicates (some
of which have negative weights). Second, the input formula is rewritten as a
conjunction of clauses, each of which is in \emph{prenex normal
  form}~\cite{hinman2018fundamentals}.

A \emph{formula} is a conjunction of clauses. A \emph{clause} is of the form
$\forall x_{1} \in \Delta_{1}\text{.}\forall x_{2} \in \Delta_{2}\dots\text{
}\forall x_{n} \in \Delta_{n}\text{.}\phi(x_{1}, x_{2}, \dots, x_{n})$, where
$\phi$ is a disjunction of literals that only contain variables
$x_{1}, \dots, x_{n}$ (and any constants). We say that a clause is a
\emph{(positive) unit clause} if
\begin{enumerate*}[label=(\roman*)]
  \item there is only one literal with a predicate, and
  \item it is a positive literal.
\end{enumerate*}
A \emph{literal} is either an atom (i.e., a \emph{positive} literal) or its
negation (i.e., a \emph{negative} literal). An \emph{atom} is either
\begin{enumerate*}[label=(\roman*)]
  \item $P(t_{1}, \dots, t_{m})$ for some predicate $P$ and terms
  $t_{1}, \dots, t_{m}$ (written $P(\mathbf{t})$ for short) or
  \item $x=y$ for some terms $x$ and $y$.
\end{enumerate*}
An atom is \emph{ground} if it contains no variables, i.e., only constants. The
\emph{arity} of a predicate is the number of arguments it takes, i.e., $m$ in
the case of predicate $P$ above. When we want to denote a predicate together
with its arity, we write $P/m$. A \emph{term} is either a variable or a
constant. Throughout the paper, we use set-theoretic notation, interpreting a
formula as a set of clauses and a clause as a set of literals. Moreover, for
readability, clauses written on separate lines are implicitly conjoined.

\subsection{WFOMC Algorithms and Their Logics}\label{sec:threelogics}

\begin{table*}
  \centering
  \begin{tabular}{llclll}
    \toprule
    Logic & Sorts & Constants & Variables & Quantifiers & Additional atoms\\
    \midrule
    \FO & one or more & \cmark & unlimited & $\forall$, $\exists$ & $x = y$\\
    \Ctwo & one & \xmark & two & $\forall$, $\exists$, $\exists^{= k}$, $\exists^{\le k}$, $\exists^{\ge k}$ & ---\\
    \UFO & one & \xmark & two & $\forall$ & $|P| = m$\\
    \bottomrule
  \end{tabular}
  \caption[]{A comparison of the three logics used in WFOMC in terms of:
    \begin{enumerate*}[label=(\roman*)]
      \item the number of sorts,
      \item support for constants,
      \item the maximum number of variables,
      \item supported quantifiers, and
      \item supported atoms in addition to those of the form $P(\mathbf{t})$ for some predicate $P/n$ and $n$-tuple of terms $\mathbf{t}$.
    \end{enumerate*}
    Here:
    \begin{enumerate*}[label=(\roman*)]
      \item $k, m \in \mathbb{N}_{0}$, the latter of which can depend on the domain size,
      \item $P$ is a predicate, and
      \item $x$ and $y$ are terms.
    \end{enumerate*}
  }\label{tbl:logics}
\end{table*}

In \cref{tbl:logics}, we outline the differences among three first-order logics
commonly used in WFOMC:
\begin{enumerate*}[label=(\roman*)]
  \item \FO{} is the input format for
  \textsc{ForcLift}\footnote{\url{https://github.com/UCLA-StarAI/Forclift}} and
  its extensions
  \textsc{Crane}\footnote{\url{https://doi.org/10.5281/zenodo.8004077}} and
  \Cranetwo{},
  \item \Ctwo{} is often used in the literature on \textsc{FastWFOMC} and
  related
  methods~\cite{DBLP:journals/jair/Kuzelka21,DBLP:conf/aaai/MalhotraS22}, and
  \item \UFO{} is the input format supported by a private version of
  \textsc{FastWFOMC} obtained directly from the authors.
\end{enumerate*}
Note that the publicly available
version\footnote{\url{https://comp.nus.edu.sg/~tvanbr/software/fastwfomc.tar.gz}}
does not support any cardinality constraints. The notation we use to refer to
each logic is standard in the case of \Ctwo{}, new in the case of \UFO{}, and
redefined to be more specific in the case of \FO{}. All three logics are
function-free, and domains are always assumed to be finite.

In \FO{}, each term is assigned to a \emph{sort}, and each predicate $P/n$ is
assigned to a sequence of $n$ sorts. Each sort has its corresponding domain.
Most of these assignments to sorts are typically left implicit and can be
reconstructed from the quantifiers. For example, $\forall x,y \in \Delta\text{.
}P(x, y)$ implies that variables $x$ and $y$ have the same sort. On the other
hand, $\forall x \in \Delta\text{. }\forall y \in \Gamma\text{. } P(x, y)$
implies that $x$ and $y$ have different sorts, and it would be improper to
write, e.g., $\forall x \in \Delta\text{. }\forall y \in \Gamma\text{.
} P(x, y) \lor x = y$. \FO{} is also the only logic to support constants,
formulas with more than two variables, and the equality predicate.

\begin{remark}
  In the case of \textsc{ForcLift} and its extensions, support for a formula as
  valid input does not imply that the algorithm can compile the formula into a
  circuit or graph suitable for lifted model counting. However, it is known that
  \textsc{ForcLift} compilation is guaranteed to succeed on any \FO{} formula
  without constants and with at most two
  variables~\cite{DBLP:conf/nips/Broeck11}.
\end{remark}

Both \Ctwo{} and \UFO{} are single-sorted (i.e., variables are always quantified
over the same domain) and only include formulas with at most two variables and
no constants. Unlike \UFO{}, \Ctwo{} supports existential quantifiers, including
\emph{counting quantifiers}. For example, $\exists^{=1} x\text{. }\phi(x)$ means
that there exists \emph{exactly one} $x$ such that $\phi(x)$, and
$\exists^{\le 2} x\text{. }\phi(x)$ means that there exist \emph{at most two}
such $x$. The advantage \UFO{} brings over \Ctwo{} is the support for
\emph{(equality) cardinality constraints}. For example, $|P| = 3$ constrains all
models to have \emph{exactly three positive literals with predicate $P$}.

\begin{definition}[Model]\label{def:model}
  Let $\phi$ be a formula in \FO{}. For every predicate $P/n$ in $\phi$, let
  ${(\Delta_{i}^{P})}_{i=1}^{n}$ be a list of the corresponding domains (not
  necessarily distinct). Let $\sigma$ be a map from the domains of $\phi$ to
  their interpretations as sets such that:
  \begin{enumerate*}[label=(\roman*)]
    \item the sets are pairwise disjoint, and
    \item the constants in $\phi$ are included in the corresponding domains.
  \end{enumerate*}
  Then a \emph{structure} of $\phi$ (with respect to $\sigma$) is a set $M$ of
  ground literals defined by adding either $P(\mathbf{t})$ or
  $\neg P(\mathbf{t})$ for every predicate $P/n$ in $\phi$ and $n$-tuple
  $\mathbf{t} \in \prod_{i=1}^{n} \sigma(\Delta_{i}^{P})$. A structure is a
  \emph{model} if it satisfies $\phi$ (see Appendix~A of~Dilkas and
  Belle~\shortcite{DBLP:journals/corr/abs-2306-04189} for more details).
\end{definition}

\begin{definition}[WFOMC~\cite{DBLP:conf/ijcai/BroeckTMDR11}]
  Continuing from \cref{def:model}, for every predicate $P$ in $\phi$, let
  $w^{+}(P), w^{-}(P) \in \mathbb{Q}$ be its \emph{weights}. Unless specified
  otherwise, we assume weights to be equal to one. The \emph{(symmetric)
    weighted first-order model count} (WFOMC) of $\phi$ (with respect to
  $\sigma$, $w^{+}$, and $w^{-}$) is
  \[
    \sum_{M \models \phi} \prod_{P(\mathbf{t}) \in M} w^{+}(P) \prod_{\neg P(\mathbf{t}) \in M} w^{-}(P),
  \]
  where the sum is over all models of $\phi$.
\end{definition}

\begin{example}[Counting functions]\label{example:functions}
  To define predicate $P$ as an endofunction on $\Delta$, in \Ctwo{} one would
  write $\forall x \in \Delta\text{. }\exists^{=1} y \in \Delta\text{.
  }P(x, y)$. In \UFO{}, the same could be written as
  \begin{equation}\label[formula]{eq:functions1}
    \begin{gathered}
      \forall x, y \in \Delta\text{. }S(x) \lor \neg P(x, y)\\
      |P| = |\Delta|,
    \end{gathered}
  \end{equation}
  where $w^{-}(S) = -1$. Although \cref{eq:functions1} has more models compared
  to its counterpart in \Ctwo{}, the weight function makes it so the
  \emph{weighted} model count of \cref{eq:functions1} is still equal to the
  number of endofunctions on $\Delta$.

  Equivalently, in \FO{} we would write
  \begin{equation}\label[formula]{eq:fo}
    \begin{gathered}
      \forall x \in \Gamma\text{. }\exists y \in \Delta\text{. }P(x, y)\\
      \forall x \in \Gamma\text{. }\forall y, z \in \Delta\text{. }P(x, y) \land P(x, z) \Rightarrow y = z,
    \end{gathered}
  \end{equation}
  first stating that each $x$ must be mapped to at least one $y$ and then adding
  that if $x$ is mapped to both $y$ and $z$, then $y=z$. Note that \cref{eq:fo}
  is written with two domains instead of one. Clearly, one can still count the
  number of endofunctions by setting $|\Gamma| = |\Delta|$. As was previously
  observed by Dilkas and Belle~\shortcite{DBLP:conf/kr/DilkasB23}, such a
  reformulation can help first-order knowledge compilation algorithms find
  efficient solutions.
\end{example}

\subsection{Algebra}\label{sec:algebra}

We write $\expr{}$ for an arbitrary algebraic expression. Some terms have
different meanings in the context of algebra compared to logic. Here, a
\emph{constant} is a non-negative integer. Similarly, a \emph{variable} is
either a parameter of a function or a variable introduced by a summation, e.g.,
$i$ in $\sum_{i=1}^{n} \expr$. A (function) \emph{signature} is
$f(x_{1}, \dots, x_{n})$ (written $f(\mathbf{x})$ for short), where $f$ is an
$n$-ary function, and each $x_{i}$ is a variable. An \emph{equation} is
$f(\mathbf{x}) = \expr{}$, where $f(\mathbf{x})$ is a signature.

\begin{definition}\label{def:basecase}
  Let $f(\mathbf{x})$ be a function call where each $x_{i}$ is either a constant
  or a variable (note that signatures are included in this definition). Then a
  function call $f(\mathbf{y})$ is a \emph{base case} of $f(\mathbf{x})$ if
  $f(\mathbf{y}) = f(\mathbf{x})\sigma$, where $\sigma$ is a substitution that
  replaces one or more $x_{i}$ with a constant.
\end{definition}

\section{Completing the Definitions of Recursive Functions}\label{sec:main}

\begin{algorithm}[t]
  \caption{\protect\CompileWithBaseCases{$\phi$}}\label{alg:compilewithbasecases}
  \KwIn{formula $\phi$}
  \KwOut{set $\mathcal{E}$ of equations}
  $(\mathcal{E}, \mathcal{F}, \mathcal{D}) \gets \Crane{$\phi$}$\;\label{line:first}
  $\mathcal{E} \gets \Simplify{$\mathcal{E}$}$\;\label{line:second}
  \ForEach{base case $f(\mathbf{x}) \in \FindBaseCases{$\mathcal{E}$}$}{
    $\psi \gets \mathcal{F}(f)$\;
    \ForEach{$i$ such that $x_{i} \in \mathbb{N}_{0}$}{
      $\psi \gets \Propagate{$\psi$, $\mathcal{D}(f, i)$, $x_i$}$\;
    }
    $\mathcal{E} \gets \mathcal{E} \cup \CompileWithBaseCases{$\psi$}$\;
  }
\end{algorithm}

\Cref{alg:compilewithbasecases} outlines our overall approach for compiling a
formula into a set of equation that include the required base cases. In short,
we first use the knowledge compilation algorithm of the original
\textsc{Crane}~\cite{DBLP:conf/kr/DilkasB23} to compile the formula into:
\begin{enumerate*}[label=(\roman*)]
  \item set $\mathcal{E}$ of equations,
  \item map $\mathcal{F}$ from function names to formulas, and
  \item map $\mathcal{D}$ from function names and argument indices to domains.
\end{enumerate*}
After some algebraic simplification, $\mathcal{E}$ is passed to the
\FindBaseCases procedure (described in \cref{sec:identifying}) that returns a
set of base cases that we need to find solutions for. For each base case
$f(\mathbf{x})$, we identify the formula associated with $f$ and simplify it
using the \Propagate procedure (described in \cref{sec:simplifying}). The
algorithm then recurses on these simplified formulas and adds the resulting base
case equations to $\mathcal{E}$. \Cref{example:overall} explains
\cref{alg:compilewithbasecases} in more detail.

\begin{example}\label{example:overall}
  Consider the following formula (previously examined by Dilkas and
  Belle~\shortcite{DBLP:conf/kr/DilkasB23}) that defines predicate $P$ to be a
  bijection between two sets $\Gamma$ and $\Delta$:
  \[
    \begin{gathered}
      \forall x \in \Gamma\text{. }\exists y \in \Delta\text{. }P(x, y)\\
      \forall y \in \Delta\text{. }\exists x \in \Gamma\text{. }P(x, y)\\
      \forall x \in \Gamma\text{. }\forall y, z \in \Delta\text{. }P(x, y) \land P(x, z) \Rightarrow y = z\\
      \forall x, z \in \Gamma\text{. }\forall y \in \Delta\text{. }P(x, y) \land P(z, y) \Rightarrow x = z.
    \end{gathered}
  \]
  In particular, we examine the first solution that \Cranebfs{} returns for this
  formula.

  After \cref{line:first,line:second}, we have
  \begin{align*}
    \mathcal{E} &= \left\{\,\begin{aligned}f(m, n) &= \sum_{l=0}^{n} \binom{n}{l}{(-1)}^{n-l}g(l, m),\\ g(l, m) &= g(l-1, m) + mg(l-1, m-1)\end{aligned}\,\right\};\\
    \mathcal{D} &= \{\, (f, 1) \mapsto \Gamma, (f, 2) \mapsto \Delta, (g, 1) \mapsto \Delta^{\top}, (g, 2) \mapsto \Gamma \,\},
  \end{align*}
  where $\Delta^{\top}$ is a new domain introduced by \Crane. Then
  \FindBaseCases identifies two base cases: $g(0, m)$ and $g(l, 0)$. In both
  cases, \CompileWithBaseCases recurses on the formula $\mathcal{F}(g)$
  simplified by assuming that one of the domains is empty. In the first case, we
  recurse on the formula $\forall x \in \Gamma\text{. }S(x) \lor \neg S(x)$,
  where $S$ is a predicate introduced by Skolemization with weights
  $w^{+}(S) = 1$ and $w^{-}(S) = -1$. Hence, we get the base case
  $g(0, m) = 0^{m}$. In the case of $g(l, 0)$, \Propagate{$\psi$, $\Gamma$, $0$}
  returns an empty formula, giving us $g(l, 0) = 1$.
\end{example}

Note that these base cases overlap when $l = m = 0$ but are consistent with each
other since $0^{0} = 1$. More generally, let $\phi$ be a formula with two
domains $\Gamma$ and $\Delta$, and let $n, m \in \mathbb{N}_{0}$. Then the model
count of \Propagate{$\phi$, $\Delta$, $n$} assuming $|\Gamma| = m$ is the same
as the model count of \Propagate{$\phi$, $\Gamma$, $m$} assuming $|\Delta| = n$.

\todo[inline]{Can I turn the above into a theorem?}

Finally, we note that the \Simplify procedure plays a crucial role in
simplifying a common algebraic pattern $\sum_{m=0}^{n}[a \le m \le b] f(m)$.
Here:
\begin{enumerate*}[label=(\roman*)]
  \item $n$ is a variable,
  \item $a, b \in \mathbb{N}_{0}$ are constants,
  \item $f$ is an expression that may depend on $m$, and
  \item $[a \le m \le b] =
  \begin{cases}
    1 & \text{if $a \le m \le b$} \\
    0 & \text{otherwise}
  \end{cases}$ is the Iverson bracket.
\end{enumerate*}
\Simplify transforms this pattern into
$f(a) + f(a+1) + \cdots + f(\min\{\, n, b \,\})$. For instance, in the case of
\cref{example:overall}, \Simplify transforms
$g(l, m) = \sum_{k=0}^{m}[0 \le k \le 1]\binom{m}{k}g(l-1, m-k)$ into the
simpler form above.

\todo[inline]{A concluding sentence that introduces the remaining subsections?}

\subsection{Identifying a Sufficient Set of Base Cases}\label{sec:identifying}

\begin{algorithm}[t]
  \caption{\protect\FindBaseCases{$\mathcal{E}$}}\label{alg:findbasecases}
  \KwIn{set $\mathcal{E}$ of equations}
  \KwOut{set $\mathcal{B}$ of base cases}

  $\mathcal{B} \gets \emptyset$\;
  \ForEach{equation $(f(\mathbf{x}) = \expr{}) \in \mathcal{E}$}{
    \ForEach{function call $f(\mathbf{y}) \in \expr{}$}{\label{line:functioncall}
      \ForEach{$y_{i} \in \mathbf{y}$} {
        \uIf{$y_{i} \in \mathbb{N}_{0}$}{
          $\mathcal{B} \gets \mathcal{B} \cup \{\, f(\mathbf{x})[x_{i} \mapsto y_{i}] \,\}$\;
        }
        \ElseIf{$y_{i} = x_{i} - c_{i}$ for some $c_{i} \in \mathbb{N}_{0}$}{
          \For{$j \gets 0$ \KwTo $c_{i} - 1$}{\label{line:lim}
            $\mathcal{B} \gets \mathcal{B} \cup \{\, f(\mathbf{x})[x_{i} \mapsto j] \,\}$\;\label{line:insert}
          }
        }
      }
    }
  }
\end{algorithm}

\Cref{alg:findbasecases} summarises the implementation of \FindBaseCases. For
each recursive call from a function $f$ to itself, we consider two types of
arguments:
\begin{enumerate*}[label=(\roman*)]
  \item constants and
  \item arguments of the form $x_{i} - c_{i}$, where $c_{i} \in \mathbb{N}_{0}$
  is a constant, and $x_{i}$ is the $i$-th argument of the signature of $f$.
\end{enumerate*}
In the former case, we consider a base case specifically for that constant. In
the latter case, we consider a base case for constants from zero up to (but not
including) $c_{i}$. \Cref{thm:halting} below motivates this approach.

\begin{example}
  Consider the recursive function $g$ from \cref{example:overall}.
  \FindBaseCases{$\mathcal{E}$} iterates over two function calls: $g(l-1, m)$
  and $g(l-1, m-1)$. The former produces the base case $g(0, m)$, while the
  latter produces both $g(0, m)$ and $g(l, 0)$.
\end{example}

\todo[inline]{Prove (lots of notes, handwritten and elsewhere)}

\begin{theorem}[Termination]\label{thm:halting}
  Let us assume the following about the input set of equations $\mathcal{E}$:
  \begin{enumerate}
    \item Each function $f$ has exactly one equation with $f$ on the left-hand
          side.\@ We call this equation the \emph{definition} of $f$.
    \item There exists a topological ordering of all functions ${(f_{i})}_{i}$
          such that the definition of $f_{i}$ does not contain function calls to
          $f_{j}$ with $j > i$.\footnote{This condition excludes the possibility
          of mutual recursion and similar cyclic scenarios and is akin to
          stratified logic programs~\cite{DBLP:books/sp/Lloyd87}.}
    \item For every equation $(f(\mathbf{x}) = \expr) \in \mathcal{E}$, every
          recursive function call $f(\mathbf{y}) \in \expr$ satisfies the
          following:
          \begin{itemize}
            \item each $y_{i}$ is either $x_{i} - c_{i}$ or $c_{i}$ for some
                  constant $c_{i} \in \mathbb{N}_{0}$;
            \item there exists $i$ such that $y_{i} = x_{i} - c_{i}$ for some
                  $c_{i} > 0$.
          \end{itemize}
  \end{enumerate}
  Then the set of base cases $\mathcal{B}$ returned by
  \FindBaseCases{$\mathcal{E}$} is \emph{sufficient} for $\mathcal{E}$ in the
  following sense. The evaluation of any function call with non-negative
  arguments terminates as long as the evaluation of base cases is prioritised
  over the corresponding recursive definitions.\footnote{Recall that, as
    previously discussed, the order in which base cases are considered is
    immaterial.}
\end{theorem}

\subsection{Propagating Domain Size Assumptions}\label{sec:simplifying}

\begin{algorithm}[t]
  \caption{\protect\Propagate{$\phi$, $\Delta$, $n$}}\label{alg:propagate}
  \KwIn{formula $\phi$, domain $\Delta$, $n \in \mathbb{N}_{0}$}
  \KwOut{formula $\phi'$}
  $\phi' \gets \emptyset$\;
  \uIf{$n = 0$}{
    \ForEach{clause $C \in \phi$}{
      \lIf{$\Delta \not\in \Doms(C)$}{$\phi' \gets \phi' \cup \{\, C \,\}$}
      \Else{
        $C' \gets \{\, l \in C \mid \Delta \not\in \Doms(l) \,\}$\;
        \If{$C' \ne \emptyset$}{
          $l \gets \text{an arbitrary literal in } C'$\;\label{line:smoothing1}
          $\phi' \gets \phi' \cup \{\, C' \cup \{\, \neg l \,\} \,\} $\;\label{line:smoothing2}
        }
      }
    }
  }
  \Else{
    $D \gets \text{a set of $n$ new constants in $\Delta$}$\;
    \ForEach{clause $C \in \phi$}{
      ${(x_{i})}_{i=1}^{m} \gets \text{the variables in $C$ with domain $\Delta$}$\;
      \lIf{$m = 0$}{$\phi' \gets \phi' \cup \{\, C \,\}$}
      \Else{
        $\phi' \gets \phi' \cup \{\, C[x_{1} \mapsto c_{1}, \dots, x_{m} \mapsto c_{m}] \mid {(c_{i})}_{i=1}^{m} \in D^{m} \,\}$\;
      }
    }
  }
\end{algorithm}

\Propagate (\cref{alg:propagate}) modifies formula $\phi$ with the assumption
that domain $\Delta$ has size $n \in \mathbb{N}_{0}$. In the case of $n=0$, many
clauses become vacuously satisfied and can be removed. In the case of
$n > 0$\footnote{None of the formulas considered in this work had $n > 1$.}, we
perform partial grounding, using constants to replace all variables quantified
over $\Delta$. \Cref{alg:propagate} considers these two cases separately. For a
literal or a clause $C$, we write $\Doms(C)$ to denote the set of corresponding
domains.

In the case of $n = 0$, consider three types of clauses:
\begin{enumerate*}[label=(\roman*)]
  \item those that do not mention $\Delta$,\label{type1}
  \item those in which every literal contains variables quantified over
  $\Delta$, and\label{type2}
  \item those that have some literals with variables quantified over $\Delta$
  and some without.\label{type3}
\end{enumerate*}
\Cref{type1} clauses are transferred to the new formula $\phi'$ unchanged. For
\cref{type2} clauses, $C' = \emptyset$, so these clauses are filtered out. One
might think that the same should be done with \cref{type3} clauses, however,
\cref{line:smoothing1,line:smoothing2} perform a new kind of smoothing, the
explanation of which we defer to \cref{sec:smoothingbase}.

In the case of $n>0$, we introduce $n$ new constants. Consider an arbitrary
clause $C \in \phi$ and let $m \in \mathbb{N}_{0}$ be the number of variables in
$C$ quantified over $\Delta$. If $m=0$, then, similarly to the previous case, we
add $C$ directly to $\phi'$. Otherwise, we add a clause to $\phi'$ with every
possible way of replacing the $m$ variables in $C$ with some combination of the
$n$ new constants.

\begin{example}
  Let us consider the clause $C \equiv \forall x \in \Gamma\text{.
  }\forall y, z \in \Delta\text{. } \neg P(x, y) \lor \neg P(x, z) \lor y=z$.
  Then
  $\Doms(C) = \Doms(\neg P(x, y)) = \Doms(\neg P(x, z)) = \{\, \Gamma, \Delta \,\}$,
  and $\Doms(y=z) = \{\, \Delta \,\}$. A call to \Propagate{$\{\, C \,\}$,
    $\Delta$, $3$} would produce the following formula with nine clauses:
  \begin{align*}
    \forall x \in \Gamma\text{. }\neg P(x, c_{1}) \lor& \neg P(x, c_{1}) \lor c_{1}=c_{1}\\
    \forall x \in \Gamma\text{. }\neg P(x, c_{1}) \lor& \neg P(x, c_{2}) \lor c_{1}=c_{2}\\
    \vdots&\\
    \forall x \in \Gamma\text{. }\neg P(x, c_{3}) \lor& \neg P(x, c_{3}) \lor c_{3}=c_{3},\\
  \end{align*}
  where $c_{1}$, $c_{2}$, and $c_{3}$ are the new constants.
\end{example}

\section{Smoothing}

Smoothness originates in propositional knowledge compilation where it is defined
as the property that, for every disjunction node, all disjuncts (as subtrees of
the circuit) contain the same atoms~\cite{darwiche2001tractable}. Van den Broeck
et al.~\shortcite{DBLP:conf/ijcai/BroeckTMDR11} generalise smoothness to
first-order logic, adding set-disjunction and inclusion-exclusion nodes
alongside disjunction.

The motivation for smoothing is as follows. Whenever compilation rules such as
unit propagation and inclusion-exclusion simplify the formula, some ground atoms
might be eliminated from consideration (see \cref{example:basecasesmoothing}
below). To account for them during counting, smoothing nodes (i.e., tautological
clauses such as $P(c) \lor \neg P(c)$) are added to the FCG at an appropriate
location.

In the rest of this section, we extend the smoothing algorithm of Van den Broeck
et al.~\shortcite{DBLP:conf/ijcai/BroeckTMDR11} even further.
\Cref{sec:smoothingbase} describes the role smoothing plays in the
base-case-finding algorithm from \cref{sec:main}. Then, \cref{sec:smoothingfcg1}
shows how to adapt smoothing to the compilation rules introduced by Dilkas and
Belle~\shortcite{DBLP:conf/kr/DilkasB23}.

\subsection{Smoothing for Base Cases}\label{sec:smoothingbase}

In this section, we motivate and describe \cref{line:smoothing1,line:smoothing2}
of \cref{alg:propagate}. Suppose that \Propagate is called with arguments
$(\phi, \Delta, 0)$, i.e., we are simplifying formula $\phi$ by assuming that
domain $\Delta$ is empty. Informally, if there is a predicate $P$ in $\phi$ that
has nothing to do with domain $\Delta$, the role of smoothing is to preserve all
occurrences of $P$ even if all clauses with $P$ become vacuously satisfied. We
note that the approach presented in this section is far from unique and explain
it via an example below.

\begin{example}\label{example:basecasesmoothing}
  Let $\phi$ be
  \begin{align}
    \forall x \in \Delta\text{. }\forall y, z \in \Gamma&\text{. }Q(x) \lor P(y, z)\label[clause]{eq:example1}\\
    \forall y, z \in \Gamma'&\text{. }P(y, z)\label[clause]{eq:example2},
  \end{align}
  where $\Gamma' \subseteq \Gamma$ is a domain introduced by a compilation rule.
  Note that, as a relation, $P \subseteq \Gamma \times \Gamma$.

  Let us reason by hand about the model count of $\phi$ when
  $\Delta = \emptyset$. Predicate $Q$ can only take one value, i.e.,
  $Q = \emptyset$. The value of $P$ is fixed over $\Gamma' \times \Gamma'$ by
  \cref{eq:example2} but is allowed to vary freely over
  $(\Gamma \times \Gamma) \setminus (\Gamma' \times \Gamma')$ since
  \cref{eq:example1} is vacuously satisfied by all structures. Hence, the right
  model count should be $2^{|\Gamma|^2 - |\Gamma'|^2}$.

  However, without \cref{line:smoothing2}, \Propagate would simplify $\phi$ to
  $\forall y, z \in \Gamma'\text{. }P(y, z)$. Here, $P$ is defined as a subset
  of $\Gamma' \times \Gamma'$. Clearly, this simplified formula has only one
  model: $\{\, P(y, z) \mid y, z \in \Gamma' \,\}$.

  With \cref{line:smoothing2} included, $\phi$ is transformed to
  \begin{gather*}
    \forall y, z \in \Gamma\text{. }P(y, z) \lor \neg P(y, z)\\
    \forall y, z \in \Gamma'\text{. }P(y, z),
  \end{gather*}
  which retains the correct model count.
\end{example}

Note that the choice of $l$ on \cref{line:smoothing1} of \cref{alg:propagate} is
inconsequential because any choice achieves the same goal: constructing a
tautological clause that retains the literals in $C'$.

\subsection{Smoothing the FCG}\label{sec:smoothingfcg1}

Smoothing is a two-step process. First, positive unit clauses (denoting sets of
ground atoms that are accounted for in the FCG) are propagated `upwards', i.e.,
in the opposite direction of FCG arcs. Then, at nodes of certain types, missing
atoms are detected and additional nodes are added to account for them. In this
section, we:
\begin{enumerate*}[label=(\roman*)]
  \item describe the relevant node types from previous work,
  \item show how smoothing ought to work for these node types, and
  \item illustrate how the new smoothing techniques work on two example FCGs.
\end{enumerate*}

Before describing the proposed changes to smoothing, we briefly review the
compilation rules (and their corresponding node types) introduced by Dilkas and
Belle~\shortcite{DBLP:conf/kr/DilkasB23} while referring to the aforementioned
paper for precise definitions. \emph{Domain recursion} selects a domain
$\Delta$, introduces a new constant $c \in \Delta$, and, for each variable $x$
quantified over $\Delta$, considers two possibilities: $x = c$ or $x \ne c$,
modifying the formula accordingly. We denote the resulting node as
$\DR(c \in \Delta)$. \emph{Constraint removal} applies to formulas such that,
for some domain $\Delta$, each variable $x$ quantified over $\Delta$ is followed
by the inequality constraint $x \ne c$. In other words, the relevant clauses can
be rewritten to begin with $\forall x \in \Delta\text{.
}x \ne c \Rightarrow \ldots$ For such formulas, constraint removal replaces
$\Delta$ with a new domain $\Delta'$ and removes the $x \ne c$ constraints. We
denote the resulting node as $\CR(\Delta' \gets \Delta \setminus \{\,c\,\})$.
Finally, \emph{caching} detects when the input formula $\phi$ is equal to a
previously-encountered formula $\psi$ except for having different domains. We
denote the resulting node as $\Reff(\sigma)$, where $\sigma$ is the substitution
mapping the domains of $\psi$ to their corresponding domains in $\phi$.

\paragraph{Stage~1 for domain recursion.}
When visiting a $\DR(c \in \Delta)$ node, for each unit clause received from the
child node, we replace each occurrence of $\phi(c)$ or
$\forall x \in \Delta\text{. } x \ne c \Rightarrow \phi(x)$ with
$\forall x \in \Delta\text{. }\phi(x)$. In other words, if the subgraph below
the domain recursion node covers some of the ground atoms that are affected by
domain recursion, then it should cover all of them. If the relevant subgraph
indeed covers those ground atoms, Stage~2 will do nothing. Otherwise, smoothing
nodes will be added below the domain recursion node to cover the difference
between the sets of ground atoms assigned to the domain recursion node and its
child node.

\paragraph{Stage~1 for constraint removal and caching.}
When visiting a $\CR(\Delta' \gets \Delta \setminus \{\,c\,\})$ node, we reverse
constraint removal. In other words, we replace each
$\forall x \in \Delta'\text{. }\phi(x)$ with $\forall x \in \Delta\text{.
}x \ne c \Rightarrow \phi(x)$. When visiting a $\Reff(\sigma)$ node, we apply
substitution $\sigma$ to the unit clauses coming from the child node.

\paragraph{Stage~2 for domain recursion.}
We need not add smoothing nodes immediately below constraint removal or caching
nodes. However, for domain recursion we do the following.
\begin{enumerate}
  \item Suppose the set of unit clauses assigned to the child node during
        Stage~1 contains both $\phi(c)$ and $\forall x \in \Delta\text{.
        } x \ne c \Rightarrow \phi(x)$. In other words, the only difference
        between the two clauses is that one has the constant $c$ whereas the
        other one has a variable $x \ne c$. In such a case, we merge the two
        clauses into $\forall x \in \Delta\text{. }\phi(x)$.
  \item If necessary, we add smoothing nodes below the domain recursion node to
        cover the difference between the unit clauses assigned to the domain
        recursion node during Stage~1 and the unit clauses of the child node
        post-processed by the step above.
\end{enumerate}

\begin{figure}
  \centering
  \begin{forest}
    for tree={sn edges}
    [$\DR(c \in \Delta)$,name=dr,ellipse,draw
    [$\land$,ellipse,draw,dashed,edge=dashed
    [$P(c)$,rectangle,draw,fill=green!20]
    [$\forall x \in \Delta\text{. }x \ne c \Rightarrow P(x) \lor \neg P(x)$,rectangle,dashed,edge=dashed,draw,fill=blue!20]
    ]
    ]
  \end{forest}
  \caption{An artificial example of an FCG where a smoothing node must be added
    below a domain recursion node. The dashed nodes and arcs are added during
    Stage~2 of smoothing.}\label{fig:smoothing2}
\end{figure}

\begin{example}[An FCG that needs smoothing]
  \Cref{fig:smoothing2} shows an FCG with a domain recursion node
  $\DR(c \in \Delta)$ immediately followed by a single $P(c)$ node. In this
  case, Stage~1 of smoothing assigns $\{\, \forall x \in \Delta\text{.
  }P(x) \,\}$ to the former node and $\{\, P(c) \,\}$ to the latter. Since these
  two sets of unit clauses cover different ground atoms, Stage~2 adds a
  smoothing node to cover $P(x)$ for all $x \in \Delta \setminus \{\, c \,\}$.
\end{example}

\begin{figure}
  \centering
  \begin{forest}
    for tree={sn edges}
    [$\DR(c \in \Delta)$,name=dr,ellipse,draw,label={right:\textcolor{blue}{3. $\{\, \forall x \in \Delta\text{. }P(x) \,\}$}}
    [$\land$,ellipse,draw,label={left:\textcolor{blue}{2. $\{\, P(c) \,\}$}},label={right:\textcolor{blue}{6. $\{\, P(c), \forall x \in \Delta\text{. }x \ne c \Rightarrow P(x) \,\}$}}
    [$P(c)$,rectangle,draw,fill=green!20,label={272:\textcolor{blue}{1. $\{\, P(c) \,\}$}}]
    [$\CR(\Delta' \gets \Delta \setminus \{\, c \,\})$,ellipse,draw,label={87:\textcolor{blue}{5. $\{\, \forall x \in \Delta\text{. }x \ne c \Rightarrow P(x) \,\}$}}
    [$\Reff(\Delta \mapsto \Delta')$,name=ref,ellipse,draw,label={below:\textcolor{blue}{4. $\{\, \forall x \in \Delta'\text{. }P(x) \,\}$}}]
    ]
    ]
    ]
  \end{forest}
  \caption{A smooth FCG loosely based on \cref{example:overall}. The arc from
    the caching node to the domain recursion node is omitted for compactness.
    The labels next to nodes show the sets of unit clauses assigned to each node
    during Stage~1 and the order of these assignments, omitting empty
    assignments that replace a set $S$ with $S$ itself.}\label{fig:smoothing1}
\end{figure}

\begin{example}[A smooth FCG]\label{example:smooth}
  Stage~1 of smoothing is more involved in the case of the FCG in
  \cref{fig:smoothing1}. The unit clause $P(c)$ propagates to the conjunction
  node and is then generalised to $\forall x \in \Delta\text{. }P(x)$ by the
  domain recursion node. It then propagates to the caching node, changing form
  to $(\forall x \in \Delta\text{.
  }P(x))[\Delta \mapsto \Delta'] \equiv \forall x \in \Delta'\text{. }P(x)$. The
  constraint removal node re-introduces the constraints, transforming the clause
  to $\forall x \in \Delta\text{. }x \ne c \Rightarrow P(x)$, which is then
  propagated to the conjunction node, joining $P(c)$.

  In Stage~2, $P(c)$ and $\forall x \in \Delta\text{. }x \ne c \Rightarrow P(x)$
  are merged into $\forall x \in \Delta\text{. }P(x)$. Since the resulting
  clause matches the clause assigned to the domain recursion node, the FCG is
  already smooth.
\end{example}

The algebraic interpretation of \cref{fig:smoothing1} is an equation $e$ that
defines a recursive function. The right-hand side of $e$ already covers $P(c)$,
and the smoothing algorithm correctly recognises that the recursive call also
covers $P(x)$ for all $x \in \Delta \setminus \{\, c \,\}$.

\section{Generating C++ Code}\label{sec:cpp}

In this section, we describe the last step of \Cranetwo{} outlined in
\cref{fig:overview}: translating the set of equations $\mathcal{E}$ produced by
\CompileWithBaseCases into C++ code. The produced C++ program can then be
compiled and executed with different command-line arguments to compute the model
count of the formula for different (combinations of) domain sizes.

Each equation in $\mathcal{E}$ is compiled into a C++ function, together with
its own cache for memoisation. Let
$e = (f(\mathbf{x}) = \expr{}) \in \mathcal{E}$ be an arbitrary equation, and
let $\mathbf{c} \in \mathbb{N}_{0}^{n}$ represent the arguments of the
corresponding C++ function. The implementation of $e$ consists of three parts.
First, we check whether $\mathbf{c}$ is already in the cache of $e$ (in which
case the cached value is returned). Second, for each base case $f(\mathbf{y})$
of $f(\mathbf{x})$ (as in \cref{def:basecase}), we check whether $\mathbf{c}$
\emph{matches} $\mathbf{y}$, i.e., $c_{i} = y_{i}$ whenever
$y_{i} \in \mathbb{N}_{0}$. In this case, $\mathbf{c}$ is redirected to the C++
function that corresponds to the definition of base case $f(\mathbf{y})$.
Finally, if the above cases fail, we evaluate $\mathbf{c}$ according to
$\expr{}$, store the result in the cache, and return it.

\section{Experimental Evaluation}

We compare \Cranetwo{} (in both BFS and greedy modes) with \textsc{FastWFOMC}
and \textsc{ForcLift} on two problems previously considered by Dilkas and
Belle~\shortcite{DBLP:conf/kr/DilkasB23}: the function-counting problem from
\cref{example:functions} and the bijection-counting problem described below.
Note that comparing \Cranetwo{} and \textsc{FastWFOMC} on a larger set of
benchmarks is challenging because there is no automated way to translate a
formula in \FO{} or \Ctwo{} into \UFO{} (or even check if such an encoding is
possible). The experiments were run on an AMD~Ryzen~7~5800H processor with
\SI{16}{\gibi\byte} of memory and Arch Linux~6.8.2-arch2-1 operating system.
\textsc{FastWFOMC} was executed using Python~3.8.19 with Python-FLINT~0.5.0.

The \FO{} formula for bijections is described in \cref{example:overall}. The
equivalent formula in \Ctwo{} is
\begin{gather*}
  \forall x \in \Delta\text{. }\exists^{=1} y \in \Delta\text{. }P(x, y)\\
  \forall y \in \Delta\text{. }\exists^{=1} x \in \Delta\text{. }P(x, y).
\end{gather*}
Similarly, in \UFO{} the same formula can be written as
\begin{gather*}
  \forall x, y \in \Delta\text{. }R(x) \lor \neg P(x, y)\\
  \forall x, y \in \Delta\text{. }S(x) \lor \neg P(y, x)\\
  |P| = |\Delta|,
\end{gather*}
where $w^{-}(R) = w^{-}(S) = -1$.

\begin{figure}
  \centering
  \input{plot}
  \caption{The runtime of WFOMC algorithms. Note that the $y$ axis is on a
    logarithmic scale.}\label{fig:plot}
\end{figure}

In the first experiment, we run each of the four algorithms once on each
combination of benchmark and domain size, the latter ranging from one up to and
including 35. \Cref{fig:plot} shows total runtime values as a measure of the
overall performance of each algorithm. For the three knowledge compilation
algorithms, we also separately track compilation and inference times and comment
on them where relevant. Note that \Cranebfs{} is able to handle more instances
than either \textsc{ForcLift} or \Cranegreedy{}, i.e., the bijection-counting
problem in our experiments as well as similar formulas examined
previously~\cite{DBLP:conf/kr/DilkasB23}.

As shown in \cref{fig:plot}, the runtimes of all compilation-based algorithms
remain practically constant in contrast to the rapidly increasing runtimes of
\textsc{FastWFOMC}. Although the search/compilation part is slower in
\Cranetwo{} than in \textsc{ForcLift}, the difference is negligible. The
runtimes of the knowledge compilation algorithms appear constant because---for
these counting problems and domain sizes---compilation time dominates inference
time (recall that compilation time is independent of domain sizes). Indeed, the
maximum inference time of both \Cranebfs{} and \Cranegreedy{} across these
experiments is only \SI{4}{\milli\second}. The runtimes of \Cranetwo{} have
lower variation than those of \textsc{ForcLift} because with \textsc{ForcLift}
we compile the formula anew for each domain size whereas with \Cranetwo{} we
compile it once and reuse the resulting C++ program for all domain sizes.

\begin{table}
  \centering
  \begin{tabular}{lrr}
    \toprule
    Algorithm & Bijections & Functions \\
    \midrule
    \Cranebfs{} & \num{e4} & \num{3e5} \\
    \Cranegreedy{} & --- & \num{3e5} \\
    \textsc{FastWFOMC} & 28 & 32 \\
    \textsc{ForcLift} & --- & 143 \\
    \bottomrule
  \end{tabular}
  \caption{The maximum domain sizes that can be handled by each algorithm in at
    most \SI{45}{\second}. `---' indicates that the algorithm is unable to find
    a solution. The domain sizes for \Cranetwo{} are accurate up to the first
    digit while other domain sizes are exact.}\label{table:results}
\end{table}

For the second experiment, we examine the maximum domain size that the
algorithms can handle in at most \SI{45}{\second} of total runtime. As shown in
\cref{table:results}, \Cranebfs{} is able to scale to domain sizes \num{357} and
\num{9375} times larger than \textsc{FastWFOMC} for bijection-counting and
function-counting problems, respectively. Note that, while \textsc{ForcLift} may
have runtime performance comparable to \Cranetwo{}, its finite-precision
arithmetic cannot represent model counts for domain sizes higher than 143 and
returns $\infty$ instead.

\section{Conclusion (TODO)}

\todo[inline,caption={}]{
  \begin{itemize}
    \item Later on:
          \begin{itemize}
            \item Eliminate (e.g., Emacs) warnings.
            \item must run it by all 3 other coauthors
            \item re-check submission instructions and formatting guidelines
            \item cite the \textsc{Crane} paper (and other algorithms' papers)
                  where necessary
          \end{itemize}
  \end{itemize}
}

\bibliographystyle{kr}
\bibliography{paper}

\end{document}