### 2024-07-30 doc/paper/plot.tex
Everything related to the new set of experiments (except for the results, which are too big to be stored here)

% Created by tikzDevice version 0.12.6 on 2024-07-30 13:11:15
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (505.89,216.81);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.5pt,line join=round,line cap=round,fill=fillColor] ( -0.00,  0.00) rectangle (505.89,216.81);
\end{scope}
\begin{scope}
\path[clip] ( 33.14, 57.56) rectangle (186.23,198.75);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] ( 33.14, 57.56) rectangle (186.23,198.75);
\definecolor{drawColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 33.14, 70.17) --
	(186.23, 70.17);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 33.14,101.68) --
	(186.23,101.68);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 33.14,133.19) --
	(186.23,133.19);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 33.14,164.70) --
	(186.23,164.70);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 33.14,196.21) --
	(186.23,196.21);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 48.13, 57.56) --
	( 48.13,198.75);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 85.60, 57.56) --
	( 85.60,198.75);

\path[draw=drawColor,line width= 0.2pt,line join=round] (123.07, 57.56) --
	(123.07,198.75);

\path[draw=drawColor,line width= 0.2pt,line join=round] (160.53, 57.56) --
	(160.53,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 33.14, 85.92) --
	(186.23, 85.92);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 33.14,117.43) --
	(186.23,117.43);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 33.14,148.95) --
	(186.23,148.95);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 33.14,180.46) --
	(186.23,180.46);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 66.87, 57.56) --
	( 66.87,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (104.33, 57.56) --
	(104.33,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (141.80, 57.56) --
	(141.80,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (179.27, 57.56) --
	(179.27,198.75);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.10,106.36) --
	( 45.46,106.36) --
	( 50.81,106.36) --
	( 56.16,106.36) --
	( 61.51,106.36) --
	( 66.87,106.36) --
	( 72.22,106.37) --
	( 77.57,106.38) --
	( 82.92,106.42) --
	( 88.28,106.58) --
	( 93.63,107.23) --
	( 98.98,109.99) --
	(104.33,117.90);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 45.46,123.48) --
	( 50.81,112.22) --
	( 56.16,112.22) --
	( 61.51,112.42) --
	( 66.87,119.02) --
	( 72.22,162.96);
\definecolor{fillColor}{RGB}{27,158,119}

\path[fill=fillColor] ( 40.10,106.36) circle (  1.96);

\path[fill=fillColor] ( 45.46,106.36) circle (  1.96);

\path[fill=fillColor] ( 50.81,106.36) circle (  1.96);

\path[fill=fillColor] ( 56.16,106.36) circle (  1.96);

\path[fill=fillColor] ( 61.51,106.36) circle (  1.96);

\path[fill=fillColor] ( 66.87,106.36) circle (  1.96);

\path[fill=fillColor] ( 72.22,106.37) circle (  1.96);

\path[fill=fillColor] ( 77.57,106.38) circle (  1.96);

\path[fill=fillColor] ( 82.92,106.42) circle (  1.96);

\path[fill=fillColor] ( 88.28,106.58) circle (  1.96);

\path[fill=fillColor] ( 93.63,107.23) circle (  1.96);

\path[fill=fillColor] ( 98.98,109.99) circle (  1.96);

\path[fill=fillColor] (104.33,117.90) circle (  1.96);
\definecolor{fillColor}{RGB}{117,112,179}

\path[fill=fillColor] ( 43.49,121.52) --
	( 47.42,121.52) --
	( 47.42,125.44) --
	( 43.49,125.44) --
	cycle;

\path[fill=fillColor] ( 48.85,110.26) --
	( 52.77,110.26) --
	( 52.77,114.18) --
	( 48.85,114.18) --
	cycle;

\path[fill=fillColor] ( 54.20,110.26) --
	( 58.12,110.26) --
	( 58.12,114.18) --
	( 54.20,114.18) --
	cycle;

\path[fill=fillColor] ( 59.55,110.46) --
	( 63.47,110.46) --
	( 63.47,114.38) --
	( 59.55,114.38) --
	cycle;

\path[fill=fillColor] ( 64.90,117.06) --
	( 68.83,117.06) --
	( 68.83,120.99) --
	( 64.90,120.99) --
	cycle;

\path[fill=fillColor] ( 70.26,161.00) --
	( 74.18,161.00) --
	( 74.18,164.92) --
	( 70.26,164.92) --
	cycle;
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 34.75, 57.56) -- ( 34.75, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.10, 57.56) -- ( 40.10, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 45.46, 57.56) -- ( 45.46, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 50.81, 57.56) -- ( 50.81, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 56.16, 57.56) -- ( 56.16, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 61.51, 57.56) -- ( 61.51, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 66.87, 57.56) -- ( 66.87, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 72.22, 57.56) -- ( 72.22, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 77.57, 57.56) -- ( 77.57, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 82.92, 57.56) -- ( 82.92, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 88.28, 57.56) -- ( 88.28, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 93.63, 57.56) -- ( 93.63, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 98.98, 57.56) -- ( 98.98, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (104.33, 57.56) -- (104.33, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (109.69, 57.56) -- (109.69, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (115.04, 57.56) -- (115.04, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (120.39, 57.56) -- (120.39, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (125.74, 57.56) -- (125.74, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (131.10, 57.56) -- (131.10, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (136.45, 57.56) -- (136.45, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (141.80, 57.56) -- (141.80, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (147.15, 57.56) -- (147.15, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (152.51, 57.56) -- (152.51, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (157.86, 57.56) -- (157.86, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (163.21, 57.56) -- (163.21, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (168.56, 57.56) -- (168.56, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (173.92, 57.56) -- (173.92, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (179.27, 57.56) -- (179.27, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (184.62, 57.56) -- (184.62, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14, 64.92) -- ( 38.84, 64.92);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14, 75.42) -- ( 38.84, 75.42);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14, 85.92) -- ( 38.84, 85.92);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14, 96.43) -- ( 38.84, 96.43);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,106.93) -- ( 38.84,106.93);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,117.43) -- ( 38.84,117.43);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,127.94) -- ( 38.84,127.94);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,138.44) -- ( 38.84,138.44);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,148.95) -- ( 38.84,148.95);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,159.45) -- ( 38.84,159.45);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,169.95) -- ( 38.84,169.95);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,180.46) -- ( 38.84,180.46);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,190.96) -- ( 38.84,190.96);
\end{scope}
\begin{scope}
\path[clip] (190.73, 57.56) rectangle (343.81,198.75);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (190.73, 57.56) rectangle (343.81,198.75);
\definecolor{drawColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.2pt,line join=round] (190.73, 70.17) --
	(343.81, 70.17);

\path[draw=drawColor,line width= 0.2pt,line join=round] (190.73,101.68) --
	(343.81,101.68);

\path[draw=drawColor,line width= 0.2pt,line join=round] (190.73,133.19) --
	(343.81,133.19);

\path[draw=drawColor,line width= 0.2pt,line join=round] (190.73,164.70) --
	(343.81,164.70);

\path[draw=drawColor,line width= 0.2pt,line join=round] (190.73,196.21) --
	(343.81,196.21);

\path[draw=drawColor,line width= 0.2pt,line join=round] (205.71, 57.56) --
	(205.71,198.75);

\path[draw=drawColor,line width= 0.2pt,line join=round] (243.18, 57.56) --
	(243.18,198.75);

\path[draw=drawColor,line width= 0.2pt,line join=round] (280.65, 57.56) --
	(280.65,198.75);

\path[draw=drawColor,line width= 0.2pt,line join=round] (318.12, 57.56) --
	(318.12,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (190.73, 85.92) --
	(343.81, 85.92);

\path[draw=drawColor,line width= 0.5pt,line join=round] (190.73,117.43) --
	(343.81,117.43);

\path[draw=drawColor,line width= 0.5pt,line join=round] (190.73,148.95) --
	(343.81,148.95);

\path[draw=drawColor,line width= 0.5pt,line join=round] (190.73,180.46) --
	(343.81,180.46);

\path[draw=drawColor,line width= 0.5pt,line join=round] (224.45, 57.56) --
	(224.45,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (261.91, 57.56) --
	(261.91,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (299.38, 57.56) --
	(299.38,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (336.85, 57.56) --
	(336.85,198.75);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (197.68, 87.62) --
	(203.04, 87.62) --
	(208.39, 87.61) --
	(213.74, 87.61) --
	(219.09, 87.61) --
	(224.45, 87.61) --
	(229.80, 87.61) --
	(235.15, 87.62) --
	(240.50, 87.69) --
	(245.86, 88.58) --
	(251.21, 97.29) --
	(256.56,123.32) --
	(261.91,157.55) --
	(267.27,192.33);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (197.68, 87.37) --
	(203.04, 87.37) --
	(208.39, 87.36) --
	(213.74, 87.36) --
	(219.09, 87.36) --
	(224.45, 87.36) --
	(229.80, 87.36) --
	(235.15, 87.37) --
	(240.50, 87.44) --
	(245.86, 88.34) --
	(251.21, 97.16) --
	(256.56,123.35) --
	(261.91,157.54) --
	(267.27,192.33);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] (203.04,113.24) --
	(208.39,111.85) --
	(213.74,111.59) --
	(219.09,111.66) --
	(224.45,112.12) --
	(229.80,112.27) --
	(235.15,112.50) --
	(240.50,116.23) --
	(245.86,130.94) --
	(251.21,168.51);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (203.04, 68.42) --
	(208.39, 63.98) --
	(213.74, 63.98) --
	(219.09, 63.98);
\definecolor{fillColor}{RGB}{117,112,179}

\path[fill=fillColor] (201.07,111.28) --
	(205.00,111.28) --
	(205.00,115.20) --
	(201.07,115.20) --
	cycle;

\path[fill=fillColor] (206.43,109.89) --
	(210.35,109.89) --
	(210.35,113.82) --
	(206.43,113.82) --
	cycle;

\path[fill=fillColor] (211.78,109.63) --
	(215.70,109.63) --
	(215.70,113.55) --
	(211.78,113.55) --
	cycle;

\path[fill=fillColor] (217.13,109.70) --
	(221.06,109.70) --
	(221.06,113.62) --
	(217.13,113.62) --
	cycle;

\path[fill=fillColor] (222.49,110.16) --
	(226.41,110.16) --
	(226.41,114.09) --
	(222.49,114.09) --
	cycle;

\path[fill=fillColor] (227.84,110.31) --
	(231.76,110.31) --
	(231.76,114.23) --
	(227.84,114.23) --
	cycle;

\path[fill=fillColor] (233.19,110.53) --
	(237.11,110.53) --
	(237.11,114.46) --
	(233.19,114.46) --
	cycle;

\path[fill=fillColor] (238.54,114.27) --
	(242.47,114.27) --
	(242.47,118.20) --
	(238.54,118.20) --
	cycle;

\path[fill=fillColor] (243.90,128.97) --
	(247.82,128.97) --
	(247.82,132.90) --
	(243.90,132.90) --
	cycle;

\path[fill=fillColor] (249.25,166.55) --
	(253.17,166.55) --
	(253.17,170.47) --
	(249.25,170.47) --
	cycle;
\definecolor{fillColor}{RGB}{217,95,2}

\path[fill=fillColor] (197.68, 90.42) --
	(200.33, 85.84) --
	(195.04, 85.84) --
	cycle;

\path[fill=fillColor] (203.04, 90.42) --
	(205.68, 85.84) --
	(200.39, 85.84) --
	cycle;

\path[fill=fillColor] (208.39, 90.41) --
	(211.03, 85.84) --
	(205.75, 85.84) --
	cycle;

\path[fill=fillColor] (213.74, 90.41) --
	(216.38, 85.84) --
	(211.10, 85.84) --
	cycle;

\path[fill=fillColor] (219.09, 90.41) --
	(221.74, 85.84) --
	(216.45, 85.84) --
	cycle;

\path[fill=fillColor] (224.45, 90.41) --
	(227.09, 85.84) --
	(221.80, 85.84) --
	cycle;

\path[fill=fillColor] (229.80, 90.41) --
	(232.44, 85.84) --
	(227.16, 85.84) --
	cycle;

\path[fill=fillColor] (235.15, 90.42) --
	(237.79, 85.84) --
	(232.51, 85.84) --
	cycle;

\path[fill=fillColor] (240.50, 90.49) --
	(243.15, 85.92) --
	(237.86, 85.92) --
	cycle;

\path[fill=fillColor] (245.86, 91.39) --
	(248.50, 86.82) --
	(243.21, 86.82) --
	cycle;

\path[fill=fillColor] (251.21,100.21) --
	(253.85, 95.64) --
	(248.57, 95.64) --
	cycle;

\path[fill=fillColor] (256.56,126.40) --
	(259.20,121.82) --
	(253.92,121.82) --
	cycle;

\path[fill=fillColor] (261.91,160.59) --
	(264.56,156.02) --
	(259.27,156.02) --
	cycle;

\path[fill=fillColor] (267.27,195.39) --
	(269.91,190.81) --
	(264.62,190.81) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (200.26, 68.42) -- (205.81, 68.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (203.04, 65.64) -- (203.04, 71.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (205.61, 63.98) -- (211.16, 63.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (208.39, 61.20) -- (208.39, 66.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (210.97, 63.98) -- (216.52, 63.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (213.74, 61.20) -- (213.74, 66.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (216.32, 63.98) -- (221.87, 63.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (219.09, 61.20) -- (219.09, 66.75);
\definecolor{fillColor}{RGB}{27,158,119}

\path[fill=fillColor] (197.68, 87.62) circle (  1.96);

\path[fill=fillColor] (203.04, 87.62) circle (  1.96);

\path[fill=fillColor] (208.39, 87.61) circle (  1.96);

\path[fill=fillColor] (213.74, 87.61) circle (  1.96);

\path[fill=fillColor] (219.09, 87.61) circle (  1.96);

\path[fill=fillColor] (224.45, 87.61) circle (  1.96);

\path[fill=fillColor] (229.80, 87.61) circle (  1.96);

\path[fill=fillColor] (235.15, 87.62) circle (  1.96);

\path[fill=fillColor] (240.50, 87.69) circle (  1.96);

\path[fill=fillColor] (245.86, 88.58) circle (  1.96);

\path[fill=fillColor] (251.21, 97.29) circle (  1.96);

\path[fill=fillColor] (256.56,123.32) circle (  1.96);

\path[fill=fillColor] (261.91,157.55) circle (  1.96);

\path[fill=fillColor] (267.27,192.33) circle (  1.96);
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (192.33, 57.56) -- (192.33, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (197.68, 57.56) -- (197.68, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (203.04, 57.56) -- (203.04, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (208.39, 57.56) -- (208.39, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (213.74, 57.56) -- (213.74, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (219.09, 57.56) -- (219.09, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (224.45, 57.56) -- (224.45, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (229.80, 57.56) -- (229.80, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (235.15, 57.56) -- (235.15, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (240.50, 57.56) -- (240.50, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (245.86, 57.56) -- (245.86, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.21, 57.56) -- (251.21, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (256.56, 57.56) -- (256.56, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (261.91, 57.56) -- (261.91, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (267.27, 57.56) -- (267.27, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (272.62, 57.56) -- (272.62, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (277.97, 57.56) -- (277.97, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (283.32, 57.56) -- (283.32, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (288.68, 57.56) -- (288.68, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (294.03, 57.56) -- (294.03, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (299.38, 57.56) -- (299.38, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (304.73, 57.56) -- (304.73, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (310.09, 57.56) -- (310.09, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (315.44, 57.56) -- (315.44, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (320.79, 57.56) -- (320.79, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (326.14, 57.56) -- (326.14, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (331.50, 57.56) -- (331.50, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (336.85, 57.56) -- (336.85, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (342.20, 57.56) -- (342.20, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73, 64.92) -- (196.42, 64.92);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73, 75.42) -- (196.42, 75.42);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73, 85.92) -- (196.42, 85.92);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73, 96.43) -- (196.42, 96.43);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73,106.93) -- (196.42,106.93);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73,117.43) -- (196.42,117.43);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73,127.94) -- (196.42,127.94);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73,138.44) -- (196.42,138.44);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73,148.95) -- (196.42,148.95);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73,159.45) -- (196.42,159.45);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73,169.95) -- (196.42,169.95);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73,180.46) -- (196.42,180.46);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73,190.96) -- (196.42,190.96);
\end{scope}
\begin{scope}
\path[clip] (348.31, 57.56) rectangle (501.39,198.75);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (348.31, 57.56) rectangle (501.39,198.75);
\definecolor{drawColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.2pt,line join=round] (348.31, 70.17) --
	(501.39, 70.17);

\path[draw=drawColor,line width= 0.2pt,line join=round] (348.31,101.68) --
	(501.39,101.68);

\path[draw=drawColor,line width= 0.2pt,line join=round] (348.31,133.19) --
	(501.39,133.19);

\path[draw=drawColor,line width= 0.2pt,line join=round] (348.31,164.70) --
	(501.39,164.70);

\path[draw=drawColor,line width= 0.2pt,line join=round] (348.31,196.21) --
	(501.39,196.21);

\path[draw=drawColor,line width= 0.2pt,line join=round] (363.30, 57.56) --
	(363.30,198.75);

\path[draw=drawColor,line width= 0.2pt,line join=round] (400.76, 57.56) --
	(400.76,198.75);

\path[draw=drawColor,line width= 0.2pt,line join=round] (438.23, 57.56) --
	(438.23,198.75);

\path[draw=drawColor,line width= 0.2pt,line join=round] (475.70, 57.56) --
	(475.70,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (348.31, 85.92) --
	(501.39, 85.92);

\path[draw=drawColor,line width= 0.5pt,line join=round] (348.31,117.43) --
	(501.39,117.43);

\path[draw=drawColor,line width= 0.5pt,line join=round] (348.31,148.95) --
	(501.39,148.95);

\path[draw=drawColor,line width= 0.5pt,line join=round] (348.31,180.46) --
	(501.39,180.46);

\path[draw=drawColor,line width= 0.5pt,line join=round] (382.03, 57.56) --
	(382.03,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (419.50, 57.56) --
	(419.50,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (456.96, 57.56) --
	(456.96,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (494.43, 57.56) --
	(494.43,198.75);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (355.27, 89.52) --
	(360.62, 89.52) --
	(365.97, 89.52) --
	(371.32, 89.52) --
	(376.68, 89.52) --
	(382.03, 89.52) --
	(387.38, 89.52) --
	(392.73, 89.52) --
	(398.09, 89.52) --
	(403.44, 89.52) --
	(408.79, 89.52) --
	(414.14, 89.52) --
	(419.50, 89.52) --
	(424.85, 89.53) --
	(430.20, 89.55) --
	(435.55, 89.61) --
	(440.91, 89.71) --
	(446.26, 89.97) --
	(451.61, 90.65) --
	(456.96, 92.11) --
	(462.32, 95.47) --
	(467.67,101.07) --
	(473.02,108.80) --
	(478.37,119.46) --
	(483.73,132.70) --
	(489.08,144.57) --
	(494.43,157.31);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (355.27, 88.43) --
	(360.62, 88.43) --
	(365.97, 88.43) --
	(371.32, 88.43) --
	(376.68, 88.43) --
	(382.03, 88.43) --
	(387.38, 88.43) --
	(392.73, 88.43) --
	(398.09, 88.43) --
	(403.44, 88.43) --
	(408.79, 88.43) --
	(414.14, 88.43) --
	(419.50, 88.43) --
	(424.85, 88.44) --
	(430.20, 88.46) --
	(435.55, 88.51) --
	(440.91, 88.64) --
	(446.26, 88.92) --
	(451.61, 89.63) --
	(456.96, 91.16) --
	(462.32, 94.48) --
	(467.67,100.45) --
	(473.02,108.29) --
	(478.37,119.14) --
	(483.73,132.38) --
	(489.08,144.46) --
	(494.43,157.43);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] (360.62,114.65) --
	(365.97,111.15) --
	(371.32,111.49) --
	(376.68,111.10) --
	(382.03,115.68) --
	(387.38,162.17);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (360.62, 66.36) --
	(365.97, 64.61) --
	(371.32, 64.61) --
	(376.68, 64.61) --
	(382.03, 64.61) --
	(387.38, 64.92) --
	(392.73, 64.92);
\definecolor{fillColor}{RGB}{27,158,119}

\path[fill=fillColor] (355.27, 89.52) circle (  1.96);

\path[fill=fillColor] (360.62, 89.52) circle (  1.96);

\path[fill=fillColor] (365.97, 89.52) circle (  1.96);

\path[fill=fillColor] (371.32, 89.52) circle (  1.96);

\path[fill=fillColor] (376.68, 89.52) circle (  1.96);

\path[fill=fillColor] (382.03, 89.52) circle (  1.96);

\path[fill=fillColor] (387.38, 89.52) circle (  1.96);

\path[fill=fillColor] (392.73, 89.52) circle (  1.96);

\path[fill=fillColor] (398.09, 89.52) circle (  1.96);

\path[fill=fillColor] (403.44, 89.52) circle (  1.96);

\path[fill=fillColor] (408.79, 89.52) circle (  1.96);

\path[fill=fillColor] (414.14, 89.52) circle (  1.96);

\path[fill=fillColor] (419.50, 89.52) circle (  1.96);

\path[fill=fillColor] (424.85, 89.53) circle (  1.96);

\path[fill=fillColor] (430.20, 89.55) circle (  1.96);

\path[fill=fillColor] (435.55, 89.61) circle (  1.96);

\path[fill=fillColor] (440.91, 89.71) circle (  1.96);

\path[fill=fillColor] (446.26, 89.97) circle (  1.96);

\path[fill=fillColor] (451.61, 90.65) circle (  1.96);

\path[fill=fillColor] (456.96, 92.11) circle (  1.96);

\path[fill=fillColor] (462.32, 95.47) circle (  1.96);

\path[fill=fillColor] (467.67,101.07) circle (  1.96);

\path[fill=fillColor] (473.02,108.80) circle (  1.96);

\path[fill=fillColor] (478.37,119.46) circle (  1.96);

\path[fill=fillColor] (483.73,132.70) circle (  1.96);

\path[fill=fillColor] (489.08,144.57) circle (  1.96);

\path[fill=fillColor] (494.43,157.31) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (357.84, 66.36) -- (363.39, 66.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (360.62, 63.59) -- (360.62, 69.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (363.20, 64.61) -- (368.75, 64.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (365.97, 61.83) -- (365.97, 67.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (368.55, 64.61) -- (374.10, 64.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (371.32, 61.83) -- (371.32, 67.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (373.90, 64.61) -- (379.45, 64.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (376.68, 61.83) -- (376.68, 67.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (379.25, 64.61) -- (384.80, 64.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (382.03, 61.83) -- (382.03, 67.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (384.61, 64.92) -- (390.16, 64.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (387.38, 62.14) -- (387.38, 67.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (389.96, 64.92) -- (395.51, 64.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (392.73, 62.14) -- (392.73, 67.69);
\definecolor{fillColor}{RGB}{117,112,179}

\path[fill=fillColor] (358.66,112.68) --
	(362.58,112.68) --
	(362.58,116.61) --
	(358.66,116.61) --
	cycle;

\path[fill=fillColor] (364.01,109.19) --
	(367.93,109.19) --
	(367.93,113.11) --
	(364.01,113.11) --
	cycle;

\path[fill=fillColor] (369.36,109.53) --
	(373.29,109.53) --
	(373.29,113.45) --
	(369.36,113.45) --
	cycle;

\path[fill=fillColor] (374.71,109.13) --
	(378.64,109.13) --
	(378.64,113.06) --
	(374.71,113.06) --
	cycle;

\path[fill=fillColor] (380.07,113.72) --
	(383.99,113.72) --
	(383.99,117.64) --
	(380.07,117.64) --
	cycle;

\path[fill=fillColor] (385.42,160.21) --
	(389.34,160.21) --
	(389.34,164.13) --
	(385.42,164.13) --
	cycle;
\definecolor{fillColor}{RGB}{217,95,2}

\path[fill=fillColor] (355.27, 91.48) --
	(357.91, 86.91) --
	(352.62, 86.91) --
	cycle;

\path[fill=fillColor] (360.62, 91.48) --
	(363.26, 86.90) --
	(357.98, 86.90) --
	cycle;

\path[fill=fillColor] (365.97, 91.48) --
	(368.61, 86.90) --
	(363.33, 86.90) --
	cycle;

\path[fill=fillColor] (371.32, 91.48) --
	(373.97, 86.90) --
	(368.68, 86.90) --
	cycle;

\path[fill=fillColor] (376.68, 91.48) --
	(379.32, 86.90) --
	(374.03, 86.90) --
	cycle;

\path[fill=fillColor] (382.03, 91.48) --
	(384.67, 86.90) --
	(379.39, 86.90) --
	cycle;

\path[fill=fillColor] (387.38, 91.48) --
	(390.02, 86.90) --
	(384.74, 86.90) --
	cycle;

\path[fill=fillColor] (392.73, 91.48) --
	(395.38, 86.91) --
	(390.09, 86.91) --
	cycle;

\path[fill=fillColor] (398.09, 91.48) --
	(400.73, 86.90) --
	(395.44, 86.90) --
	cycle;

\path[fill=fillColor] (403.44, 91.48) --
	(406.08, 86.90) --
	(400.80, 86.90) --
	cycle;

\path[fill=fillColor] (408.79, 91.48) --
	(411.43, 86.91) --
	(406.15, 86.91) --
	cycle;

\path[fill=fillColor] (414.14, 91.48) --
	(416.79, 86.91) --
	(411.50, 86.91) --
	cycle;

\path[fill=fillColor] (419.50, 91.48) --
	(422.14, 86.91) --
	(416.85, 86.91) --
	cycle;

\path[fill=fillColor] (424.85, 91.50) --
	(427.49, 86.92) --
	(422.21, 86.92) --
	cycle;

\path[fill=fillColor] (430.20, 91.51) --
	(432.84, 86.93) --
	(427.56, 86.93) --
	cycle;

\path[fill=fillColor] (435.55, 91.57) --
	(438.20, 86.99) --
	(432.91, 86.99) --
	cycle;

\path[fill=fillColor] (440.91, 91.69) --
	(443.55, 87.11) --
	(438.26, 87.11) --
	cycle;

\path[fill=fillColor] (446.26, 91.97) --
	(448.90, 87.39) --
	(443.62, 87.39) --
	cycle;

\path[fill=fillColor] (451.61, 92.68) --
	(454.25, 88.10) --
	(448.97, 88.10) --
	cycle;

\path[fill=fillColor] (456.96, 94.21) --
	(459.61, 89.64) --
	(454.32, 89.64) --
	cycle;

\path[fill=fillColor] (462.32, 97.53) --
	(464.96, 92.95) --
	(459.67, 92.95) --
	cycle;

\path[fill=fillColor] (467.67,103.50) --
	(470.31, 98.92) --
	(465.03, 98.92) --
	cycle;

\path[fill=fillColor] (473.02,111.34) --
	(475.66,106.76) --
	(470.38,106.76) --
	cycle;

\path[fill=fillColor] (478.37,122.19) --
	(481.02,117.61) --
	(475.73,117.61) --
	cycle;

\path[fill=fillColor] (483.73,135.43) --
	(486.37,130.86) --
	(481.08,130.86) --
	cycle;

\path[fill=fillColor] (489.08,147.51) --
	(491.72,142.93) --
	(486.44,142.93) --
	cycle;

\path[fill=fillColor] (494.43,160.48) --
	(497.07,155.91) --
	(491.79,155.91) --
	cycle;
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (349.91, 57.56) -- (349.91, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (355.27, 57.56) -- (355.27, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (360.62, 57.56) -- (360.62, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (365.97, 57.56) -- (365.97, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (371.32, 57.56) -- (371.32, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (376.68, 57.56) -- (376.68, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (382.03, 57.56) -- (382.03, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (387.38, 57.56) -- (387.38, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (392.73, 57.56) -- (392.73, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (398.09, 57.56) -- (398.09, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (403.44, 57.56) -- (403.44, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (408.79, 57.56) -- (408.79, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (414.14, 57.56) -- (414.14, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (419.50, 57.56) -- (419.50, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (424.85, 57.56) -- (424.85, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (430.20, 57.56) -- (430.20, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (435.55, 57.56) -- (435.55, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (440.91, 57.56) -- (440.91, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (446.26, 57.56) -- (446.26, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (451.61, 57.56) -- (451.61, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (456.96, 57.56) -- (456.96, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (462.32, 57.56) -- (462.32, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (467.67, 57.56) -- (467.67, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (473.02, 57.56) -- (473.02, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (478.37, 57.56) -- (478.37, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (483.73, 57.56) -- (483.73, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (489.08, 57.56) -- (489.08, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (494.43, 57.56) -- (494.43, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (499.78, 57.56) -- (499.78, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31, 64.92) -- (354.00, 64.92);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31, 75.42) -- (354.00, 75.42);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31, 85.92) -- (354.00, 85.92);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31, 96.43) -- (354.00, 96.43);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31,106.93) -- (354.00,106.93);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31,117.43) -- (354.00,117.43);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31,127.94) -- (354.00,127.94);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31,138.44) -- (354.00,138.44);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31,148.95) -- (354.00,148.95);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31,159.45) -- (354.00,159.45);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31,169.95) -- (354.00,169.95);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31,180.46) -- (354.00,180.46);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31,190.96) -- (354.00,190.96);
\end{scope}
\begin{scope}
\path[clip] ( 33.14,198.75) rectangle (186.23,212.31);
\definecolor{fillColor}{gray}{0.85}

\path[fill=fillColor] ( 33.14,198.75) rectangle (186.23,212.31);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (109.69,203.05) {Bijections};
\end{scope}
\begin{scope}
\path[clip] (190.73,198.75) rectangle (343.81,212.31);
\definecolor{fillColor}{gray}{0.85}

\path[fill=fillColor] (190.73,198.75) rectangle (343.81,212.31);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (267.27,203.05) {Friends};
\end{scope}
\begin{scope}
\path[clip] (348.31,198.75) rectangle (501.39,212.31);
\definecolor{fillColor}{gray}{0.85}

\path[fill=fillColor] (348.31,198.75) rectangle (501.39,212.31);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (424.85,203.05) {Functions};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 66.87, 55.31) --
	( 66.87, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (104.33, 55.31) --
	(104.33, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (141.80, 55.31) --
	(141.80, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (179.27, 55.31) --
	(179.27, 57.56);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 66.87, 48.55) {$2^{5}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (104.33, 48.55) {$2^{12}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (141.80, 48.55) {$2^{19}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (179.27, 48.55) {$2^{26}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] (224.45, 55.31) --
	(224.45, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (261.91, 55.31) --
	(261.91, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (299.38, 55.31) --
	(299.38, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (336.85, 55.31) --
	(336.85, 57.56);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (224.45, 48.55) {$2^{5}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (261.91, 48.55) {$2^{12}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (299.38, 48.55) {$2^{19}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (336.85, 48.55) {$2^{26}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] (382.03, 55.31) --
	(382.03, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (419.50, 55.31) --
	(419.50, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (456.96, 55.31) --
	(456.96, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (494.43, 55.31) --
	(494.43, 57.56);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (382.03, 48.55) {$2^{5}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (419.50, 48.55) {$2^{12}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (456.96, 48.55) {$2^{19}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (494.43, 48.55) {$2^{26}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 29.09, 83.44) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 29.09,114.96) {16};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 29.09,146.47) {128};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 29.09,177.98) {1024};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 30.89, 85.92) --
	( 33.14, 85.92);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 30.89,117.43) --
	( 33.14,117.43);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 30.89,148.95) --
	( 33.14,148.95);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 30.89,180.46) --
	( 33.14,180.46);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (267.27, 38.70) {Domain size};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 10.70,128.16) {Total runtime (s)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (124.58,  4.50) rectangle (409.95, 27.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (133.58,  9.00) rectangle (148.04, 23.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (135.03, 16.23) -- (146.59, 16.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{fillColor}{RGB}{27,158,119}

\path[fill=fillColor] (140.81, 16.23) circle (  1.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (201.65,  9.00) rectangle (216.10, 23.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (203.10, 16.23) -- (214.66, 16.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{fillColor}{RGB}{217,95,2}

\path[fill=fillColor] (208.88, 19.28) --
	(211.52, 14.70) --
	(206.24, 14.70) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (281.98,  9.00) rectangle (296.43, 23.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] (283.42, 16.23) -- (294.99, 16.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{fillColor}{RGB}{117,112,179}

\path[fill=fillColor] (287.24, 14.26) --
	(291.17, 14.26) --
	(291.17, 18.19) --
	(287.24, 18.19) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (353.42,  9.00) rectangle (367.87, 23.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (354.86, 16.23) -- (366.43, 16.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (357.87, 16.23) -- (363.42, 16.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (360.65, 13.45) -- (360.65, 19.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at (152.54, 13.75) {\textsc{Crane2-BFS}};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at (220.60, 13.75) {\textsc{Crane2-Greedy}};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at (300.93, 13.75) {\textsc{FastWFOMC}};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at (372.37, 13.75) {\textsc{ForcLift}};
\end{scope}
\end{tikzpicture}

\end{document}