### 2024-08-09 doc/paper/plot.tex
First (second?) draft of the 'experiments' section

% Created by tikzDevice version 0.12.6 on 2024-08-07 14:16:35
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (505.89,216.81);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.5pt,line join=round,line cap=round,fill=fillColor] ( -0.00,  0.00) rectangle (505.89,216.81);
\end{scope}
\begin{scope}
\path[clip] ( 33.14, 57.56) rectangle (186.23,198.75);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] ( 33.14, 57.56) rectangle (186.23,198.75);
\definecolor{drawColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 33.14, 70.17) --
	(186.23, 70.17);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 33.14,101.68) --
	(186.23,101.68);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 33.14,133.19) --
	(186.23,133.19);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 33.14,164.70) --
	(186.23,164.70);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 33.14,196.21) --
	(186.23,196.21);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 42.89, 57.56) --
	( 42.89,198.75);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 81.85, 57.56) --
	( 81.85,198.75);

\path[draw=drawColor,line width= 0.2pt,line join=round] (120.82, 57.56) --
	(120.82,198.75);

\path[draw=drawColor,line width= 0.2pt,line join=round] (159.78, 57.56) --
	(159.78,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 33.14, 85.92) --
	(186.23, 85.92);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 33.14,117.43) --
	(186.23,117.43);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 33.14,148.95) --
	(186.23,148.95);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 33.14,180.46) --
	(186.23,180.46);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 62.37, 57.56) --
	( 62.37,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (101.34, 57.56) --
	(101.34,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (140.30, 57.56) --
	(140.30,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (179.27, 57.56) --
	(179.27,198.75);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.10,106.36) --
	( 45.67,106.36) --
	( 51.24,106.36) --
	( 56.80,106.36) --
	( 62.37,106.36) --
	( 67.94,106.37) --
	( 73.50,106.38) --
	( 79.07,106.42) --
	( 84.64,106.58) --
	( 90.20,107.23) --
	( 95.77,109.99) --
	(101.34,117.90);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 40.10,123.48) --
	( 45.67,112.22) --
	( 51.24,112.22) --
	( 56.80,112.42) --
	( 62.37,119.02) --
	( 67.94,162.96);
\definecolor{fillColor}{RGB}{27,158,119}

\path[fill=fillColor] ( 40.10,106.36) circle (  1.96);

\path[fill=fillColor] ( 45.67,106.36) circle (  1.96);

\path[fill=fillColor] ( 51.24,106.36) circle (  1.96);

\path[fill=fillColor] ( 56.80,106.36) circle (  1.96);

\path[fill=fillColor] ( 62.37,106.36) circle (  1.96);

\path[fill=fillColor] ( 67.94,106.37) circle (  1.96);

\path[fill=fillColor] ( 73.50,106.38) circle (  1.96);

\path[fill=fillColor] ( 79.07,106.42) circle (  1.96);

\path[fill=fillColor] ( 84.64,106.58) circle (  1.96);

\path[fill=fillColor] ( 90.20,107.23) circle (  1.96);

\path[fill=fillColor] ( 95.77,109.99) circle (  1.96);

\path[fill=fillColor] (101.34,117.90) circle (  1.96);
\definecolor{fillColor}{RGB}{117,112,179}

\path[fill=fillColor] ( 38.14,121.52) --
	( 42.06,121.52) --
	( 42.06,125.44) --
	( 38.14,125.44) --
	cycle;

\path[fill=fillColor] ( 43.71,110.26) --
	( 47.63,110.26) --
	( 47.63,114.18) --
	( 43.71,114.18) --
	cycle;

\path[fill=fillColor] ( 49.27,110.26) --
	( 53.20,110.26) --
	( 53.20,114.18) --
	( 49.27,114.18) --
	cycle;

\path[fill=fillColor] ( 54.84,110.46) --
	( 58.76,110.46) --
	( 58.76,114.38) --
	( 54.84,114.38) --
	cycle;

\path[fill=fillColor] ( 60.41,117.06) --
	( 64.33,117.06) --
	( 64.33,120.99) --
	( 60.41,120.99) --
	cycle;

\path[fill=fillColor] ( 65.97,161.00) --
	( 69.90,161.00) --
	( 69.90,164.92) --
	( 65.97,164.92) --
	cycle;
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 34.54, 57.56) -- ( 34.54, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.10, 57.56) -- ( 40.10, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 45.67, 57.56) -- ( 45.67, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 51.24, 57.56) -- ( 51.24, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 56.80, 57.56) -- ( 56.80, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 62.37, 57.56) -- ( 62.37, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 67.94, 57.56) -- ( 67.94, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 73.50, 57.56) -- ( 73.50, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 79.07, 57.56) -- ( 79.07, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 84.64, 57.56) -- ( 84.64, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 90.20, 57.56) -- ( 90.20, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 95.77, 57.56) -- ( 95.77, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (101.34, 57.56) -- (101.34, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (106.90, 57.56) -- (106.90, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (112.47, 57.56) -- (112.47, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (118.04, 57.56) -- (118.04, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (123.60, 57.56) -- (123.60, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (129.17, 57.56) -- (129.17, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (134.74, 57.56) -- (134.74, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (140.30, 57.56) -- (140.30, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (145.87, 57.56) -- (145.87, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (151.44, 57.56) -- (151.44, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (157.00, 57.56) -- (157.00, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (162.57, 57.56) -- (162.57, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (168.13, 57.56) -- (168.13, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (173.70, 57.56) -- (173.70, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (179.27, 57.56) -- (179.27, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (184.83, 57.56) -- (184.83, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14, 64.92) -- ( 38.84, 64.92);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14, 75.42) -- ( 38.84, 75.42);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14, 85.92) -- ( 38.84, 85.92);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14, 96.43) -- ( 38.84, 96.43);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,106.93) -- ( 38.84,106.93);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,117.43) -- ( 38.84,117.43);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,127.94) -- ( 38.84,127.94);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,138.44) -- ( 38.84,138.44);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,148.95) -- ( 38.84,148.95);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,159.45) -- ( 38.84,159.45);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,169.95) -- ( 38.84,169.95);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,180.46) -- ( 38.84,180.46);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,190.96) -- ( 38.84,190.96);
\end{scope}
\begin{scope}
\path[clip] (190.73, 57.56) rectangle (343.81,198.75);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (190.73, 57.56) rectangle (343.81,198.75);
\definecolor{drawColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.2pt,line join=round] (190.73, 70.17) --
	(343.81, 70.17);

\path[draw=drawColor,line width= 0.2pt,line join=round] (190.73,101.68) --
	(343.81,101.68);

\path[draw=drawColor,line width= 0.2pt,line join=round] (190.73,133.19) --
	(343.81,133.19);

\path[draw=drawColor,line width= 0.2pt,line join=round] (190.73,164.70) --
	(343.81,164.70);

\path[draw=drawColor,line width= 0.2pt,line join=round] (190.73,196.21) --
	(343.81,196.21);

\path[draw=drawColor,line width= 0.2pt,line join=round] (200.47, 57.56) --
	(200.47,198.75);

\path[draw=drawColor,line width= 0.2pt,line join=round] (239.43, 57.56) --
	(239.43,198.75);

\path[draw=drawColor,line width= 0.2pt,line join=round] (278.40, 57.56) --
	(278.40,198.75);

\path[draw=drawColor,line width= 0.2pt,line join=round] (317.37, 57.56) --
	(317.37,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (190.73, 85.92) --
	(343.81, 85.92);

\path[draw=drawColor,line width= 0.5pt,line join=round] (190.73,117.43) --
	(343.81,117.43);

\path[draw=drawColor,line width= 0.5pt,line join=round] (190.73,148.95) --
	(343.81,148.95);

\path[draw=drawColor,line width= 0.5pt,line join=round] (190.73,180.46) --
	(343.81,180.46);

\path[draw=drawColor,line width= 0.5pt,line join=round] (219.95, 57.56) --
	(219.95,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (258.92, 57.56) --
	(258.92,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (297.88, 57.56) --
	(297.88,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (336.85, 57.56) --
	(336.85,198.75);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (197.68, 87.62) --
	(203.25, 87.61) --
	(208.82, 87.61) --
	(214.38, 87.61) --
	(219.95, 87.61) --
	(225.52, 87.61) --
	(231.08, 87.62) --
	(236.65, 87.69) --
	(242.22, 88.58) --
	(247.78, 97.29) --
	(253.35,123.32) --
	(258.92,157.55) --
	(264.48,192.33);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (197.68, 87.37) --
	(203.25, 87.36) --
	(208.82, 87.36) --
	(214.38, 87.36) --
	(219.95, 87.36) --
	(225.52, 87.36) --
	(231.08, 87.37) --
	(236.65, 87.44) --
	(242.22, 88.34) --
	(247.78, 97.16) --
	(253.35,123.35) --
	(258.92,157.54) --
	(264.48,192.33);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] (197.68,113.24) --
	(203.25,111.85) --
	(208.82,111.59) --
	(214.38,111.66) --
	(219.95,112.12) --
	(225.52,112.27) --
	(231.08,112.50) --
	(236.65,116.23) --
	(242.22,130.94) --
	(247.78,168.51);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (197.68, 68.42) --
	(203.25, 63.98) --
	(208.82, 63.98) --
	(214.38, 63.98);
\definecolor{fillColor}{RGB}{117,112,179}

\path[fill=fillColor] (195.72,111.28) --
	(199.65,111.28) --
	(199.65,115.20) --
	(195.72,115.20) --
	cycle;

\path[fill=fillColor] (201.29,109.89) --
	(205.21,109.89) --
	(205.21,113.82) --
	(201.29,113.82) --
	cycle;

\path[fill=fillColor] (206.86,109.63) --
	(210.78,109.63) --
	(210.78,113.55) --
	(206.86,113.55) --
	cycle;

\path[fill=fillColor] (212.42,109.70) --
	(216.35,109.70) --
	(216.35,113.62) --
	(212.42,113.62) --
	cycle;

\path[fill=fillColor] (217.99,110.16) --
	(221.91,110.16) --
	(221.91,114.09) --
	(217.99,114.09) --
	cycle;

\path[fill=fillColor] (223.56,110.31) --
	(227.48,110.31) --
	(227.48,114.23) --
	(223.56,114.23) --
	cycle;

\path[fill=fillColor] (229.12,110.53) --
	(233.05,110.53) --
	(233.05,114.46) --
	(229.12,114.46) --
	cycle;

\path[fill=fillColor] (234.69,114.27) --
	(238.61,114.27) --
	(238.61,118.20) --
	(234.69,118.20) --
	cycle;

\path[fill=fillColor] (240.26,128.97) --
	(244.18,128.97) --
	(244.18,132.90) --
	(240.26,132.90) --
	cycle;

\path[fill=fillColor] (245.82,166.55) --
	(249.75,166.55) --
	(249.75,170.47) --
	(245.82,170.47) --
	cycle;
\definecolor{fillColor}{RGB}{217,95,2}

\path[fill=fillColor] (197.68, 90.42) --
	(200.33, 85.84) --
	(195.04, 85.84) --
	cycle;

\path[fill=fillColor] (203.25, 90.41) --
	(205.89, 85.84) --
	(200.61, 85.84) --
	cycle;

\path[fill=fillColor] (208.82, 90.41) --
	(211.46, 85.84) --
	(206.18, 85.84) --
	cycle;

\path[fill=fillColor] (214.38, 90.41) --
	(217.03, 85.84) --
	(211.74, 85.84) --
	cycle;

\path[fill=fillColor] (219.95, 90.41) --
	(222.59, 85.84) --
	(217.31, 85.84) --
	cycle;

\path[fill=fillColor] (225.52, 90.41) --
	(228.16, 85.84) --
	(222.88, 85.84) --
	cycle;

\path[fill=fillColor] (231.08, 90.42) --
	(233.73, 85.84) --
	(228.44, 85.84) --
	cycle;

\path[fill=fillColor] (236.65, 90.49) --
	(239.29, 85.92) --
	(234.01, 85.92) --
	cycle;

\path[fill=fillColor] (242.22, 91.39) --
	(244.86, 86.82) --
	(239.57, 86.82) --
	cycle;

\path[fill=fillColor] (247.78,100.21) --
	(250.43, 95.64) --
	(245.14, 95.64) --
	cycle;

\path[fill=fillColor] (253.35,126.40) --
	(255.99,121.82) --
	(250.71,121.82) --
	cycle;

\path[fill=fillColor] (258.92,160.59) --
	(261.56,156.02) --
	(256.27,156.02) --
	cycle;

\path[fill=fillColor] (264.48,195.39) --
	(267.13,190.81) --
	(261.84,190.81) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (194.91, 68.42) -- (200.46, 68.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (197.68, 65.64) -- (197.68, 71.19);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (200.48, 63.98) -- (206.03, 63.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (203.25, 61.20) -- (203.25, 66.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (206.04, 63.98) -- (211.59, 63.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (208.82, 61.20) -- (208.82, 66.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (211.61, 63.98) -- (217.16, 63.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (214.38, 61.20) -- (214.38, 66.75);
\definecolor{fillColor}{RGB}{27,158,119}

\path[fill=fillColor] (197.68, 87.62) circle (  1.96);

\path[fill=fillColor] (203.25, 87.61) circle (  1.96);

\path[fill=fillColor] (208.82, 87.61) circle (  1.96);

\path[fill=fillColor] (214.38, 87.61) circle (  1.96);

\path[fill=fillColor] (219.95, 87.61) circle (  1.96);

\path[fill=fillColor] (225.52, 87.61) circle (  1.96);

\path[fill=fillColor] (231.08, 87.62) circle (  1.96);

\path[fill=fillColor] (236.65, 87.69) circle (  1.96);

\path[fill=fillColor] (242.22, 88.58) circle (  1.96);

\path[fill=fillColor] (247.78, 97.29) circle (  1.96);

\path[fill=fillColor] (253.35,123.32) circle (  1.96);

\path[fill=fillColor] (258.92,157.55) circle (  1.96);

\path[fill=fillColor] (264.48,192.33) circle (  1.96);
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (192.12, 57.56) -- (192.12, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (197.68, 57.56) -- (197.68, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (203.25, 57.56) -- (203.25, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (208.82, 57.56) -- (208.82, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (214.38, 57.56) -- (214.38, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (219.95, 57.56) -- (219.95, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (225.52, 57.56) -- (225.52, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (231.08, 57.56) -- (231.08, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (236.65, 57.56) -- (236.65, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (242.22, 57.56) -- (242.22, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (247.78, 57.56) -- (247.78, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (253.35, 57.56) -- (253.35, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (258.92, 57.56) -- (258.92, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (264.48, 57.56) -- (264.48, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (270.05, 57.56) -- (270.05, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (275.62, 57.56) -- (275.62, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (281.18, 57.56) -- (281.18, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (286.75, 57.56) -- (286.75, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (292.32, 57.56) -- (292.32, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.88, 57.56) -- (297.88, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (303.45, 57.56) -- (303.45, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (309.02, 57.56) -- (309.02, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (314.58, 57.56) -- (314.58, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (320.15, 57.56) -- (320.15, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (325.72, 57.56) -- (325.72, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (331.28, 57.56) -- (331.28, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (336.85, 57.56) -- (336.85, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (342.42, 57.56) -- (342.42, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73, 64.92) -- (196.42, 64.92);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73, 75.42) -- (196.42, 75.42);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73, 85.92) -- (196.42, 85.92);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73, 96.43) -- (196.42, 96.43);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73,106.93) -- (196.42,106.93);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73,117.43) -- (196.42,117.43);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73,127.94) -- (196.42,127.94);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73,138.44) -- (196.42,138.44);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73,148.95) -- (196.42,148.95);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73,159.45) -- (196.42,159.45);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73,169.95) -- (196.42,169.95);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73,180.46) -- (196.42,180.46);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73,190.96) -- (196.42,190.96);
\end{scope}
\begin{scope}
\path[clip] (348.31, 57.56) rectangle (501.39,198.75);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (348.31, 57.56) rectangle (501.39,198.75);
\definecolor{drawColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.2pt,line join=round] (348.31, 70.17) --
	(501.39, 70.17);

\path[draw=drawColor,line width= 0.2pt,line join=round] (348.31,101.68) --
	(501.39,101.68);

\path[draw=drawColor,line width= 0.2pt,line join=round] (348.31,133.19) --
	(501.39,133.19);

\path[draw=drawColor,line width= 0.2pt,line join=round] (348.31,164.70) --
	(501.39,164.70);

\path[draw=drawColor,line width= 0.2pt,line join=round] (348.31,196.21) --
	(501.39,196.21);

\path[draw=drawColor,line width= 0.2pt,line join=round] (358.05, 57.56) --
	(358.05,198.75);

\path[draw=drawColor,line width= 0.2pt,line join=round] (397.02, 57.56) --
	(397.02,198.75);

\path[draw=drawColor,line width= 0.2pt,line join=round] (435.98, 57.56) --
	(435.98,198.75);

\path[draw=drawColor,line width= 0.2pt,line join=round] (474.95, 57.56) --
	(474.95,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (348.31, 85.92) --
	(501.39, 85.92);

\path[draw=drawColor,line width= 0.5pt,line join=round] (348.31,117.43) --
	(501.39,117.43);

\path[draw=drawColor,line width= 0.5pt,line join=round] (348.31,148.95) --
	(501.39,148.95);

\path[draw=drawColor,line width= 0.5pt,line join=round] (348.31,180.46) --
	(501.39,180.46);

\path[draw=drawColor,line width= 0.5pt,line join=round] (377.53, 57.56) --
	(377.53,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (416.50, 57.56) --
	(416.50,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (455.47, 57.56) --
	(455.47,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (494.43, 57.56) --
	(494.43,198.75);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (355.27, 89.52) --
	(360.83, 89.52) --
	(366.40, 89.52) --
	(371.97, 89.52) --
	(377.53, 89.52) --
	(383.10, 89.52) --
	(388.67, 89.52) --
	(394.23, 89.52) --
	(399.80, 89.52) --
	(405.37, 89.52) --
	(410.93, 89.52) --
	(416.50, 89.52) --
	(422.07, 89.53) --
	(427.63, 89.55) --
	(433.20, 89.61) --
	(438.77, 89.71) --
	(444.33, 89.97) --
	(449.90, 90.65) --
	(455.47, 92.11) --
	(461.03, 95.47) --
	(466.60,101.07) --
	(472.17,108.80) --
	(477.73,119.46) --
	(483.30,132.70) --
	(488.87,144.57) --
	(494.43,157.31);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (355.27, 88.43) --
	(360.83, 88.43) --
	(366.40, 88.43) --
	(371.97, 88.43) --
	(377.53, 88.43) --
	(383.10, 88.43) --
	(388.67, 88.43) --
	(394.23, 88.43) --
	(399.80, 88.43) --
	(405.37, 88.43) --
	(410.93, 88.43) --
	(416.50, 88.43) --
	(422.07, 88.44) --
	(427.63, 88.46) --
	(433.20, 88.51) --
	(438.77, 88.64) --
	(444.33, 88.92) --
	(449.90, 89.63) --
	(455.47, 91.16) --
	(461.03, 94.48) --
	(466.60,100.45) --
	(472.17,108.29) --
	(477.73,119.14) --
	(483.30,132.38) --
	(488.87,144.46) --
	(494.43,157.43);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] (355.27,114.65) --
	(360.83,111.15) --
	(366.40,111.49) --
	(371.97,111.10) --
	(377.53,115.68) --
	(383.10,162.17);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (355.27, 66.36) --
	(360.83, 64.61) --
	(366.40, 64.61) --
	(371.97, 64.61) --
	(377.53, 64.61) --
	(383.10, 64.92) --
	(388.67, 64.92);
\definecolor{fillColor}{RGB}{27,158,119}

\path[fill=fillColor] (355.27, 89.52) circle (  1.96);

\path[fill=fillColor] (360.83, 89.52) circle (  1.96);

\path[fill=fillColor] (366.40, 89.52) circle (  1.96);

\path[fill=fillColor] (371.97, 89.52) circle (  1.96);

\path[fill=fillColor] (377.53, 89.52) circle (  1.96);

\path[fill=fillColor] (383.10, 89.52) circle (  1.96);

\path[fill=fillColor] (388.67, 89.52) circle (  1.96);

\path[fill=fillColor] (394.23, 89.52) circle (  1.96);

\path[fill=fillColor] (399.80, 89.52) circle (  1.96);

\path[fill=fillColor] (405.37, 89.52) circle (  1.96);

\path[fill=fillColor] (410.93, 89.52) circle (  1.96);

\path[fill=fillColor] (416.50, 89.52) circle (  1.96);

\path[fill=fillColor] (422.07, 89.53) circle (  1.96);

\path[fill=fillColor] (427.63, 89.55) circle (  1.96);

\path[fill=fillColor] (433.20, 89.61) circle (  1.96);

\path[fill=fillColor] (438.77, 89.71) circle (  1.96);

\path[fill=fillColor] (444.33, 89.97) circle (  1.96);

\path[fill=fillColor] (449.90, 90.65) circle (  1.96);

\path[fill=fillColor] (455.47, 92.11) circle (  1.96);

\path[fill=fillColor] (461.03, 95.47) circle (  1.96);

\path[fill=fillColor] (466.60,101.07) circle (  1.96);

\path[fill=fillColor] (472.17,108.80) circle (  1.96);

\path[fill=fillColor] (477.73,119.46) circle (  1.96);

\path[fill=fillColor] (483.30,132.70) circle (  1.96);

\path[fill=fillColor] (488.87,144.57) circle (  1.96);

\path[fill=fillColor] (494.43,157.31) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (352.49, 66.36) -- (358.04, 66.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (355.27, 63.59) -- (355.27, 69.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (358.06, 64.61) -- (363.61, 64.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (360.83, 61.83) -- (360.83, 67.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (363.62, 64.61) -- (369.17, 64.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (366.40, 61.83) -- (366.40, 67.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (369.19, 64.61) -- (374.74, 64.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (371.97, 61.83) -- (371.97, 67.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (374.76, 64.61) -- (380.31, 64.61);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (377.53, 61.83) -- (377.53, 67.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (380.32, 64.92) -- (385.87, 64.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (383.10, 62.14) -- (383.10, 67.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (385.89, 64.92) -- (391.44, 64.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (388.67, 62.14) -- (388.67, 67.69);
\definecolor{fillColor}{RGB}{117,112,179}

\path[fill=fillColor] (353.30,112.68) --
	(357.23,112.68) --
	(357.23,116.61) --
	(353.30,116.61) --
	cycle;

\path[fill=fillColor] (358.87,109.19) --
	(362.80,109.19) --
	(362.80,113.11) --
	(358.87,113.11) --
	cycle;

\path[fill=fillColor] (364.44,109.53) --
	(368.36,109.53) --
	(368.36,113.45) --
	(364.44,113.45) --
	cycle;

\path[fill=fillColor] (370.00,109.13) --
	(373.93,109.13) --
	(373.93,113.06) --
	(370.00,113.06) --
	cycle;

\path[fill=fillColor] (375.57,113.72) --
	(379.50,113.72) --
	(379.50,117.64) --
	(375.57,117.64) --
	cycle;

\path[fill=fillColor] (381.14,160.21) --
	(385.06,160.21) --
	(385.06,164.13) --
	(381.14,164.13) --
	cycle;
\definecolor{fillColor}{RGB}{217,95,2}

\path[fill=fillColor] (355.27, 91.48) --
	(357.91, 86.90) --
	(352.62, 86.90) --
	cycle;

\path[fill=fillColor] (360.83, 91.48) --
	(363.48, 86.90) --
	(358.19, 86.90) --
	cycle;

\path[fill=fillColor] (366.40, 91.48) --
	(369.04, 86.90) --
	(363.76, 86.90) --
	cycle;

\path[fill=fillColor] (371.97, 91.48) --
	(374.61, 86.90) --
	(369.32, 86.90) --
	cycle;

\path[fill=fillColor] (377.53, 91.48) --
	(380.18, 86.90) --
	(374.89, 86.90) --
	cycle;

\path[fill=fillColor] (383.10, 91.48) --
	(385.74, 86.90) --
	(380.46, 86.90) --
	cycle;

\path[fill=fillColor] (388.67, 91.48) --
	(391.31, 86.91) --
	(386.02, 86.91) --
	cycle;

\path[fill=fillColor] (394.23, 91.48) --
	(396.88, 86.90) --
	(391.59, 86.90) --
	cycle;

\path[fill=fillColor] (399.80, 91.48) --
	(402.44, 86.90) --
	(397.16, 86.90) --
	cycle;

\path[fill=fillColor] (405.37, 91.48) --
	(408.01, 86.91) --
	(402.72, 86.91) --
	cycle;

\path[fill=fillColor] (410.93, 91.48) --
	(413.58, 86.91) --
	(408.29, 86.91) --
	cycle;

\path[fill=fillColor] (416.50, 91.48) --
	(419.14, 86.91) --
	(413.86, 86.91) --
	cycle;

\path[fill=fillColor] (422.07, 91.50) --
	(424.71, 86.92) --
	(419.42, 86.92) --
	cycle;

\path[fill=fillColor] (427.63, 91.51) --
	(430.27, 86.93) --
	(424.99, 86.93) --
	cycle;

\path[fill=fillColor] (433.20, 91.57) --
	(435.84, 86.99) --
	(430.56, 86.99) --
	cycle;

\path[fill=fillColor] (438.77, 91.69) --
	(441.41, 87.11) --
	(436.12, 87.11) --
	cycle;

\path[fill=fillColor] (444.33, 91.97) --
	(446.97, 87.39) --
	(441.69, 87.39) --
	cycle;

\path[fill=fillColor] (449.90, 92.68) --
	(452.54, 88.10) --
	(447.26, 88.10) --
	cycle;

\path[fill=fillColor] (455.47, 94.21) --
	(458.11, 89.64) --
	(452.82, 89.64) --
	cycle;

\path[fill=fillColor] (461.03, 97.53) --
	(463.67, 92.95) --
	(458.39, 92.95) --
	cycle;

\path[fill=fillColor] (466.60,103.50) --
	(469.24, 98.92) --
	(463.96, 98.92) --
	cycle;

\path[fill=fillColor] (472.17,111.34) --
	(474.81,106.76) --
	(469.52,106.76) --
	cycle;

\path[fill=fillColor] (477.73,122.19) --
	(480.37,117.61) --
	(475.09,117.61) --
	cycle;

\path[fill=fillColor] (483.30,135.43) --
	(485.94,130.86) --
	(480.66,130.86) --
	cycle;

\path[fill=fillColor] (488.87,147.51) --
	(491.51,142.93) --
	(486.22,142.93) --
	cycle;

\path[fill=fillColor] (494.43,160.48) --
	(497.07,155.91) --
	(491.79,155.91) --
	cycle;
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (349.70, 57.56) -- (349.70, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (355.27, 57.56) -- (355.27, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (360.83, 57.56) -- (360.83, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (366.40, 57.56) -- (366.40, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (371.97, 57.56) -- (371.97, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (377.53, 57.56) -- (377.53, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (383.10, 57.56) -- (383.10, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (388.67, 57.56) -- (388.67, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (394.23, 57.56) -- (394.23, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (399.80, 57.56) -- (399.80, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (405.37, 57.56) -- (405.37, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (410.93, 57.56) -- (410.93, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (416.50, 57.56) -- (416.50, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (422.07, 57.56) -- (422.07, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (427.63, 57.56) -- (427.63, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (433.20, 57.56) -- (433.20, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (438.77, 57.56) -- (438.77, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (444.33, 57.56) -- (444.33, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (449.90, 57.56) -- (449.90, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (455.47, 57.56) -- (455.47, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (461.03, 57.56) -- (461.03, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (466.60, 57.56) -- (466.60, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (472.17, 57.56) -- (472.17, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (477.73, 57.56) -- (477.73, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (483.30, 57.56) -- (483.30, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (488.87, 57.56) -- (488.87, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (494.43, 57.56) -- (494.43, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (500.00, 57.56) -- (500.00, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31, 64.92) -- (354.00, 64.92);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31, 75.42) -- (354.00, 75.42);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31, 85.92) -- (354.00, 85.92);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31, 96.43) -- (354.00, 96.43);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31,106.93) -- (354.00,106.93);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31,117.43) -- (354.00,117.43);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31,127.94) -- (354.00,127.94);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31,138.44) -- (354.00,138.44);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31,148.95) -- (354.00,148.95);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31,159.45) -- (354.00,159.45);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31,169.95) -- (354.00,169.95);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31,180.46) -- (354.00,180.46);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31,190.96) -- (354.00,190.96);
\end{scope}
\begin{scope}
\path[clip] ( 33.14,198.75) rectangle (186.23,212.31);
\definecolor{fillColor}{gray}{0.85}

\path[fill=fillColor] ( 33.14,198.75) rectangle (186.23,212.31);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (109.69,203.05) {Bijections};
\end{scope}
\begin{scope}
\path[clip] (190.73,198.75) rectangle (343.81,212.31);
\definecolor{fillColor}{gray}{0.85}

\path[fill=fillColor] (190.73,198.75) rectangle (343.81,212.31);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (267.27,203.05) {Friends};
\end{scope}
\begin{scope}
\path[clip] (348.31,198.75) rectangle (501.39,212.31);
\definecolor{fillColor}{gray}{0.85}

\path[fill=fillColor] (348.31,198.75) rectangle (501.39,212.31);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (424.85,203.05) {Functions};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 62.37, 55.31) --
	( 62.37, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (101.34, 55.31) --
	(101.34, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (140.30, 55.31) --
	(140.30, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (179.27, 55.31) --
	(179.27, 57.56);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 62.37, 48.55) {$2^{5}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (101.34, 48.55) {$2^{12}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (140.30, 48.55) {$2^{19}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (179.27, 48.55) {$2^{26}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] (219.95, 55.31) --
	(219.95, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (258.92, 55.31) --
	(258.92, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (297.88, 55.31) --
	(297.88, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (336.85, 55.31) --
	(336.85, 57.56);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (219.95, 48.55) {$2^{5}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (258.92, 48.55) {$2^{12}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (297.88, 48.55) {$2^{19}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (336.85, 48.55) {$2^{26}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] (377.53, 55.31) --
	(377.53, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (416.50, 55.31) --
	(416.50, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (455.47, 55.31) --
	(455.47, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (494.43, 55.31) --
	(494.43, 57.56);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (377.53, 48.55) {$2^{5}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (416.50, 48.55) {$2^{12}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (455.47, 48.55) {$2^{19}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (494.43, 48.55) {$2^{26}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 29.09, 83.44) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 29.09,114.96) {16};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 29.09,146.47) {128};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 29.09,177.98) {1024};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 30.89, 85.92) --
	( 33.14, 85.92);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 30.89,117.43) --
	( 33.14,117.43);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 30.89,148.95) --
	( 33.14,148.95);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 30.89,180.46) --
	( 33.14,180.46);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (267.27, 38.70) {Domain size};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 10.70,128.16) {Total runtime (s)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (124.58,  4.50) rectangle (409.95, 27.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (133.58,  9.00) rectangle (148.04, 23.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (135.03, 16.23) -- (146.59, 16.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{fillColor}{RGB}{27,158,119}

\path[fill=fillColor] (140.81, 16.23) circle (  1.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (201.65,  9.00) rectangle (216.10, 23.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (203.10, 16.23) -- (214.66, 16.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{fillColor}{RGB}{217,95,2}

\path[fill=fillColor] (208.88, 19.28) --
	(211.52, 14.70) --
	(206.24, 14.70) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (281.98,  9.00) rectangle (296.43, 23.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] (283.42, 16.23) -- (294.99, 16.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{fillColor}{RGB}{117,112,179}

\path[fill=fillColor] (287.24, 14.26) --
	(291.17, 14.26) --
	(291.17, 18.19) --
	(287.24, 18.19) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (353.42,  9.00) rectangle (367.87, 23.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (354.86, 16.23) -- (366.43, 16.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (357.87, 16.23) -- (363.42, 16.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (360.65, 13.45) -- (360.65, 19.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at (152.54, 13.75) {\textsc{Crane2-BFS}};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at (220.60, 13.75) {\textsc{Crane2-Greedy}};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at (300.93, 13.75) {\textsc{FastWFOMC}};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at (372.37, 13.75) {\textsc{ForcLift}};
\end{scope}
\end{tikzpicture}

\end{document}