### 2025-01-11 doc/old_paper/plot.tex
First draft of the SAT version

% Created by tikzDevice version 0.12.6 on 2024-08-15 12:12:55
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (505.89,216.81);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.5pt,line join=round,line cap=round,fill=fillColor] ( -0.00,  0.00) rectangle (505.89,216.81);
\end{scope}
\begin{scope}
\path[clip] ( 33.14, 57.56) rectangle (186.23,198.75);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] ( 33.14, 57.56) rectangle (186.23,198.75);
\definecolor{drawColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 33.14, 70.17) --
	(186.23, 70.17);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 33.14,101.68) --
	(186.23,101.68);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 33.14,133.19) --
	(186.23,133.19);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 33.14,164.70) --
	(186.23,164.70);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 33.14,196.21) --
	(186.23,196.21);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 42.89, 57.56) --
	( 42.89,198.75);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 81.85, 57.56) --
	( 81.85,198.75);

\path[draw=drawColor,line width= 0.2pt,line join=round] (120.82, 57.56) --
	(120.82,198.75);

\path[draw=drawColor,line width= 0.2pt,line join=round] (159.78, 57.56) --
	(159.78,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 33.14, 85.92) --
	(186.23, 85.92);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 33.14,117.43) --
	(186.23,117.43);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 33.14,148.95) --
	(186.23,148.95);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 33.14,180.46) --
	(186.23,180.46);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 62.37, 57.56) --
	( 62.37,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (101.34, 57.56) --
	(101.34,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (140.30, 57.56) --
	(140.30,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (179.27, 57.56) --
	(179.27,198.75);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 40.10,123.48) --
	( 45.67,112.22) --
	( 51.24,112.22) --
	( 56.80,112.42) --
	( 62.37,119.02) --
	( 67.94,162.96);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 40.10,106.36) --
	( 45.67,106.36) --
	( 51.24,106.36) --
	( 56.80,106.36) --
	( 62.37,106.36) --
	( 67.94,106.37) --
	( 73.50,106.38) --
	( 79.07,106.42) --
	( 84.64,106.58) --
	( 90.20,107.23) --
	( 95.77,109.99) --
	(101.34,117.90);
\definecolor{fillColor}{RGB}{117,112,179}

\path[fill=fillColor] ( 38.14,104.40) --
	( 42.06,104.40) --
	( 42.06,108.33) --
	( 38.14,108.33) --
	cycle;

\path[fill=fillColor] ( 43.71,104.40) --
	( 47.63,104.40) --
	( 47.63,108.33) --
	( 43.71,108.33) --
	cycle;

\path[fill=fillColor] ( 49.27,104.40) --
	( 53.20,104.40) --
	( 53.20,108.33) --
	( 49.27,108.33) --
	cycle;

\path[fill=fillColor] ( 54.84,104.40) --
	( 58.76,104.40) --
	( 58.76,108.33) --
	( 54.84,108.33) --
	cycle;

\path[fill=fillColor] ( 60.41,104.40) --
	( 64.33,104.40) --
	( 64.33,108.33) --
	( 60.41,108.33) --
	cycle;

\path[fill=fillColor] ( 65.97,104.40) --
	( 69.90,104.40) --
	( 69.90,108.33) --
	( 65.97,108.33) --
	cycle;

\path[fill=fillColor] ( 71.54,104.41) --
	( 75.46,104.41) --
	( 75.46,108.34) --
	( 71.54,108.34) --
	cycle;

\path[fill=fillColor] ( 77.11,104.45) --
	( 81.03,104.45) --
	( 81.03,108.38) --
	( 77.11,108.38) --
	cycle;

\path[fill=fillColor] ( 82.67,104.62) --
	( 86.60,104.62) --
	( 86.60,108.54) --
	( 82.67,108.54) --
	cycle;

\path[fill=fillColor] ( 88.24,105.27) --
	( 92.16,105.27) --
	( 92.16,109.19) --
	( 88.24,109.19) --
	cycle;

\path[fill=fillColor] ( 93.81,108.02) --
	( 97.73,108.02) --
	( 97.73,111.95) --
	( 93.81,111.95) --
	cycle;

\path[fill=fillColor] ( 99.37,115.94) --
	(103.30,115.94) --
	(103.30,119.87) --
	( 99.37,119.87) --
	cycle;
\definecolor{fillColor}{RGB}{27,158,119}

\path[fill=fillColor] ( 40.10,123.48) circle (  1.96);

\path[fill=fillColor] ( 45.67,112.22) circle (  1.96);

\path[fill=fillColor] ( 51.24,112.22) circle (  1.96);

\path[fill=fillColor] ( 56.80,112.42) circle (  1.96);

\path[fill=fillColor] ( 62.37,119.02) circle (  1.96);

\path[fill=fillColor] ( 67.94,162.96) circle (  1.96);
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 34.54, 57.56) -- ( 34.54, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 40.10, 57.56) -- ( 40.10, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 45.67, 57.56) -- ( 45.67, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 51.24, 57.56) -- ( 51.24, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 56.80, 57.56) -- ( 56.80, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 62.37, 57.56) -- ( 62.37, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 67.94, 57.56) -- ( 67.94, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 73.50, 57.56) -- ( 73.50, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 79.07, 57.56) -- ( 79.07, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 84.64, 57.56) -- ( 84.64, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 90.20, 57.56) -- ( 90.20, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 95.77, 57.56) -- ( 95.77, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (101.34, 57.56) -- (101.34, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (106.90, 57.56) -- (106.90, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (112.47, 57.56) -- (112.47, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (118.04, 57.56) -- (118.04, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (123.60, 57.56) -- (123.60, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (129.17, 57.56) -- (129.17, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (134.74, 57.56) -- (134.74, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (140.30, 57.56) -- (140.30, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (145.87, 57.56) -- (145.87, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (151.44, 57.56) -- (151.44, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (157.00, 57.56) -- (157.00, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (162.57, 57.56) -- (162.57, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (168.13, 57.56) -- (168.13, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (173.70, 57.56) -- (173.70, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (179.27, 57.56) -- (179.27, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (184.83, 57.56) -- (184.83, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14, 64.92) -- ( 38.84, 64.92);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14, 75.42) -- ( 38.84, 75.42);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14, 85.92) -- ( 38.84, 85.92);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14, 96.43) -- ( 38.84, 96.43);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,106.93) -- ( 38.84,106.93);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,117.43) -- ( 38.84,117.43);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,127.94) -- ( 38.84,127.94);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,138.44) -- ( 38.84,138.44);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,148.95) -- ( 38.84,148.95);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,159.45) -- ( 38.84,159.45);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,169.95) -- ( 38.84,169.95);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,180.46) -- ( 38.84,180.46);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,190.96) -- ( 38.84,190.96);
\end{scope}
\begin{scope}
\path[clip] (190.73, 57.56) rectangle (343.81,198.75);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (190.73, 57.56) rectangle (343.81,198.75);
\definecolor{drawColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.2pt,line join=round] (190.73, 70.17) --
	(343.81, 70.17);

\path[draw=drawColor,line width= 0.2pt,line join=round] (190.73,101.68) --
	(343.81,101.68);

\path[draw=drawColor,line width= 0.2pt,line join=round] (190.73,133.19) --
	(343.81,133.19);

\path[draw=drawColor,line width= 0.2pt,line join=round] (190.73,164.70) --
	(343.81,164.70);

\path[draw=drawColor,line width= 0.2pt,line join=round] (190.73,196.21) --
	(343.81,196.21);

\path[draw=drawColor,line width= 0.2pt,line join=round] (200.47, 57.56) --
	(200.47,198.75);

\path[draw=drawColor,line width= 0.2pt,line join=round] (239.43, 57.56) --
	(239.43,198.75);

\path[draw=drawColor,line width= 0.2pt,line join=round] (278.40, 57.56) --
	(278.40,198.75);

\path[draw=drawColor,line width= 0.2pt,line join=round] (317.37, 57.56) --
	(317.37,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (190.73, 85.92) --
	(343.81, 85.92);

\path[draw=drawColor,line width= 0.5pt,line join=round] (190.73,117.43) --
	(343.81,117.43);

\path[draw=drawColor,line width= 0.5pt,line join=round] (190.73,148.95) --
	(343.81,148.95);

\path[draw=drawColor,line width= 0.5pt,line join=round] (190.73,180.46) --
	(343.81,180.46);

\path[draw=drawColor,line width= 0.5pt,line join=round] (219.95, 57.56) --
	(219.95,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (258.92, 57.56) --
	(258.92,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (297.88, 57.56) --
	(297.88,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (336.85, 57.56) --
	(336.85,198.75);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (197.68,113.24) --
	(203.25,111.85) --
	(208.82,111.59) --
	(214.38,111.66) --
	(219.95,112.12) --
	(225.52,112.27) --
	(231.08,112.50) --
	(236.65,116.23) --
	(242.22,130.94) --
	(247.78,168.51);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (197.68, 68.42) --
	(203.25, 63.98) --
	(208.82, 63.98) --
	(214.38, 63.98);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] (197.68, 87.62) --
	(203.25, 87.61) --
	(208.82, 87.61) --
	(214.38, 87.61) --
	(219.95, 87.61) --
	(225.52, 87.61) --
	(231.08, 87.62) --
	(236.65, 87.69) --
	(242.22, 88.58) --
	(247.78, 97.29) --
	(253.35,123.32) --
	(258.92,157.55) --
	(264.48,192.33);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (197.68, 87.37) --
	(203.25, 87.36) --
	(208.82, 87.36) --
	(214.38, 87.36) --
	(219.95, 87.36) --
	(225.52, 87.36) --
	(231.08, 87.37) --
	(236.65, 87.44) --
	(242.22, 88.34) --
	(247.78, 97.16) --
	(253.35,123.35) --
	(258.92,157.54) --
	(264.48,192.33);
\definecolor{fillColor}{RGB}{27,158,119}

\path[fill=fillColor] (197.68,113.24) circle (  1.96);

\path[fill=fillColor] (203.25,111.85) circle (  1.96);

\path[fill=fillColor] (208.82,111.59) circle (  1.96);

\path[fill=fillColor] (214.38,111.66) circle (  1.96);

\path[fill=fillColor] (219.95,112.12) circle (  1.96);

\path[fill=fillColor] (225.52,112.27) circle (  1.96);

\path[fill=fillColor] (231.08,112.50) circle (  1.96);

\path[fill=fillColor] (236.65,116.23) circle (  1.96);

\path[fill=fillColor] (242.22,130.94) circle (  1.96);

\path[fill=fillColor] (247.78,168.51) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (194.91, 87.37) -- (200.46, 87.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (197.68, 84.59) -- (197.68, 90.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (200.48, 87.36) -- (206.03, 87.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (203.25, 84.59) -- (203.25, 90.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (206.04, 87.36) -- (211.59, 87.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (208.82, 84.59) -- (208.82, 90.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (211.61, 87.36) -- (217.16, 87.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (214.38, 84.59) -- (214.38, 90.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (217.18, 87.36) -- (222.73, 87.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (219.95, 84.59) -- (219.95, 90.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (222.74, 87.36) -- (228.29, 87.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (225.52, 84.59) -- (225.52, 90.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (228.31, 87.37) -- (233.86, 87.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (231.08, 84.59) -- (231.08, 90.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (233.88, 87.44) -- (239.43, 87.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (236.65, 84.67) -- (236.65, 90.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (239.44, 88.34) -- (244.99, 88.34);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (242.22, 85.57) -- (242.22, 91.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (245.01, 97.16) -- (250.56, 97.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (247.78, 94.39) -- (247.78, 99.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (250.58,123.35) -- (256.13,123.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (253.35,120.57) -- (253.35,126.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (256.14,157.54) -- (261.69,157.54);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (258.92,154.77) -- (258.92,160.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.71,192.33) -- (267.26,192.33);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (264.48,189.56) -- (264.48,195.11);
\definecolor{fillColor}{RGB}{217,95,2}

\path[fill=fillColor] (197.68, 71.47) --
	(200.33, 66.89) --
	(195.04, 66.89) --
	cycle;

\path[fill=fillColor] (203.25, 67.03) --
	(205.89, 62.45) --
	(200.61, 62.45) --
	cycle;

\path[fill=fillColor] (208.82, 67.03) --
	(211.46, 62.45) --
	(206.18, 62.45) --
	cycle;

\path[fill=fillColor] (214.38, 67.03) --
	(217.03, 62.45) --
	(211.74, 62.45) --
	cycle;
\definecolor{fillColor}{RGB}{117,112,179}

\path[fill=fillColor] (195.72, 85.66) --
	(199.65, 85.66) --
	(199.65, 89.58) --
	(195.72, 89.58) --
	cycle;

\path[fill=fillColor] (201.29, 85.65) --
	(205.21, 85.65) --
	(205.21, 89.58) --
	(201.29, 89.58) --
	cycle;

\path[fill=fillColor] (206.86, 85.65) --
	(210.78, 85.65) --
	(210.78, 89.58) --
	(206.86, 89.58) --
	cycle;

\path[fill=fillColor] (212.42, 85.65) --
	(216.35, 85.65) --
	(216.35, 89.58) --
	(212.42, 89.58) --
	cycle;

\path[fill=fillColor] (217.99, 85.65) --
	(221.91, 85.65) --
	(221.91, 89.58) --
	(217.99, 89.58) --
	cycle;

\path[fill=fillColor] (223.56, 85.65) --
	(227.48, 85.65) --
	(227.48, 89.58) --
	(223.56, 89.58) --
	cycle;

\path[fill=fillColor] (229.12, 85.66) --
	(233.05, 85.66) --
	(233.05, 89.58) --
	(229.12, 89.58) --
	cycle;

\path[fill=fillColor] (234.69, 85.73) --
	(238.61, 85.73) --
	(238.61, 89.66) --
	(234.69, 89.66) --
	cycle;

\path[fill=fillColor] (240.26, 86.62) --
	(244.18, 86.62) --
	(244.18, 90.54) --
	(240.26, 90.54) --
	cycle;

\path[fill=fillColor] (245.82, 95.32) --
	(249.75, 95.32) --
	(249.75, 99.25) --
	(245.82, 99.25) --
	cycle;

\path[fill=fillColor] (251.39,121.35) --
	(255.31,121.35) --
	(255.31,125.28) --
	(251.39,125.28) --
	cycle;

\path[fill=fillColor] (256.96,155.59) --
	(260.88,155.59) --
	(260.88,159.51) --
	(256.96,159.51) --
	cycle;

\path[fill=fillColor] (262.52,190.37) --
	(266.45,190.37) --
	(266.45,194.30) --
	(262.52,194.30) --
	cycle;
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (192.12, 57.56) -- (192.12, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (197.68, 57.56) -- (197.68, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (203.25, 57.56) -- (203.25, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (208.82, 57.56) -- (208.82, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (214.38, 57.56) -- (214.38, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (219.95, 57.56) -- (219.95, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (225.52, 57.56) -- (225.52, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (231.08, 57.56) -- (231.08, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (236.65, 57.56) -- (236.65, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (242.22, 57.56) -- (242.22, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (247.78, 57.56) -- (247.78, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (253.35, 57.56) -- (253.35, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (258.92, 57.56) -- (258.92, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (264.48, 57.56) -- (264.48, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (270.05, 57.56) -- (270.05, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (275.62, 57.56) -- (275.62, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (281.18, 57.56) -- (281.18, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (286.75, 57.56) -- (286.75, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (292.32, 57.56) -- (292.32, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (297.88, 57.56) -- (297.88, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (303.45, 57.56) -- (303.45, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (309.02, 57.56) -- (309.02, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (314.58, 57.56) -- (314.58, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (320.15, 57.56) -- (320.15, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (325.72, 57.56) -- (325.72, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (331.28, 57.56) -- (331.28, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (336.85, 57.56) -- (336.85, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (342.42, 57.56) -- (342.42, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73, 64.92) -- (196.42, 64.92);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73, 75.42) -- (196.42, 75.42);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73, 85.92) -- (196.42, 85.92);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73, 96.43) -- (196.42, 96.43);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73,106.93) -- (196.42,106.93);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73,117.43) -- (196.42,117.43);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73,127.94) -- (196.42,127.94);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73,138.44) -- (196.42,138.44);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73,148.95) -- (196.42,148.95);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73,159.45) -- (196.42,159.45);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73,169.95) -- (196.42,169.95);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73,180.46) -- (196.42,180.46);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.73,190.96) -- (196.42,190.96);
\end{scope}
\begin{scope}
\path[clip] (348.31, 57.56) rectangle (501.39,198.75);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (348.31, 57.56) rectangle (501.39,198.75);
\definecolor{drawColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.2pt,line join=round] (348.31, 70.17) --
	(501.39, 70.17);

\path[draw=drawColor,line width= 0.2pt,line join=round] (348.31,101.68) --
	(501.39,101.68);

\path[draw=drawColor,line width= 0.2pt,line join=round] (348.31,133.19) --
	(501.39,133.19);

\path[draw=drawColor,line width= 0.2pt,line join=round] (348.31,164.70) --
	(501.39,164.70);

\path[draw=drawColor,line width= 0.2pt,line join=round] (348.31,196.21) --
	(501.39,196.21);

\path[draw=drawColor,line width= 0.2pt,line join=round] (358.05, 57.56) --
	(358.05,198.75);

\path[draw=drawColor,line width= 0.2pt,line join=round] (397.02, 57.56) --
	(397.02,198.75);

\path[draw=drawColor,line width= 0.2pt,line join=round] (435.98, 57.56) --
	(435.98,198.75);

\path[draw=drawColor,line width= 0.2pt,line join=round] (474.95, 57.56) --
	(474.95,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (348.31, 85.92) --
	(501.39, 85.92);

\path[draw=drawColor,line width= 0.5pt,line join=round] (348.31,117.43) --
	(501.39,117.43);

\path[draw=drawColor,line width= 0.5pt,line join=round] (348.31,148.95) --
	(501.39,148.95);

\path[draw=drawColor,line width= 0.5pt,line join=round] (348.31,180.46) --
	(501.39,180.46);

\path[draw=drawColor,line width= 0.5pt,line join=round] (377.53, 57.56) --
	(377.53,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (416.50, 57.56) --
	(416.50,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (455.47, 57.56) --
	(455.47,198.75);

\path[draw=drawColor,line width= 0.5pt,line join=round] (494.43, 57.56) --
	(494.43,198.75);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (355.27,114.65) --
	(360.83,111.15) --
	(366.40,111.49) --
	(371.97,111.10) --
	(377.53,115.68) --
	(383.10,162.17);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (355.27, 66.36) --
	(360.83, 64.61) --
	(366.40, 64.61) --
	(371.97, 64.61) --
	(377.53, 64.61) --
	(383.10, 64.92) --
	(388.67, 64.92);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] (355.27, 89.52) --
	(360.83, 89.52) --
	(366.40, 89.52) --
	(371.97, 89.52) --
	(377.53, 89.52) --
	(383.10, 89.52) --
	(388.67, 89.52) --
	(394.23, 89.52) --
	(399.80, 89.52) --
	(405.37, 89.52) --
	(410.93, 89.52) --
	(416.50, 89.52) --
	(422.07, 89.53) --
	(427.63, 89.55) --
	(433.20, 89.61) --
	(438.77, 89.71) --
	(444.33, 89.97) --
	(449.90, 90.65) --
	(455.47, 92.11) --
	(461.03, 95.47) --
	(466.60,101.07) --
	(472.17,108.80) --
	(477.73,119.46) --
	(483.30,132.70) --
	(488.87,144.57) --
	(494.43,157.31);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (355.27, 88.43) --
	(360.83, 88.43) --
	(366.40, 88.43) --
	(371.97, 88.43) --
	(377.53, 88.43) --
	(383.10, 88.43) --
	(388.67, 88.43) --
	(394.23, 88.43) --
	(399.80, 88.43) --
	(405.37, 88.43) --
	(410.93, 88.43) --
	(416.50, 88.43) --
	(422.07, 88.44) --
	(427.63, 88.46) --
	(433.20, 88.51) --
	(438.77, 88.64) --
	(444.33, 88.92) --
	(449.90, 89.63) --
	(455.47, 91.16) --
	(461.03, 94.48) --
	(466.60,100.45) --
	(472.17,108.29) --
	(477.73,119.14) --
	(483.30,132.38) --
	(488.87,144.46) --
	(494.43,157.43);
\definecolor{fillColor}{RGB}{117,112,179}

\path[fill=fillColor] (353.30, 87.55) --
	(357.23, 87.55) --
	(357.23, 91.48) --
	(353.30, 91.48) --
	cycle;

\path[fill=fillColor] (358.87, 87.55) --
	(362.80, 87.55) --
	(362.80, 91.48) --
	(358.87, 91.48) --
	cycle;

\path[fill=fillColor] (364.44, 87.55) --
	(368.36, 87.55) --
	(368.36, 91.48) --
	(364.44, 91.48) --
	cycle;

\path[fill=fillColor] (370.00, 87.55) --
	(373.93, 87.55) --
	(373.93, 91.48) --
	(370.00, 91.48) --
	cycle;

\path[fill=fillColor] (375.57, 87.55) --
	(379.50, 87.55) --
	(379.50, 91.48) --
	(375.57, 91.48) --
	cycle;

\path[fill=fillColor] (381.14, 87.56) --
	(385.06, 87.56) --
	(385.06, 91.48) --
	(381.14, 91.48) --
	cycle;

\path[fill=fillColor] (386.70, 87.55) --
	(390.63, 87.55) --
	(390.63, 91.48) --
	(386.70, 91.48) --
	cycle;

\path[fill=fillColor] (392.27, 87.55) --
	(396.19, 87.55) --
	(396.19, 91.48) --
	(392.27, 91.48) --
	cycle;

\path[fill=fillColor] (397.84, 87.55) --
	(401.76, 87.55) --
	(401.76, 91.48) --
	(397.84, 91.48) --
	cycle;

\path[fill=fillColor] (403.40, 87.56) --
	(407.33, 87.56) --
	(407.33, 91.48) --
	(403.40, 91.48) --
	cycle;

\path[fill=fillColor] (408.97, 87.56) --
	(412.89, 87.56) --
	(412.89, 91.48) --
	(408.97, 91.48) --
	cycle;

\path[fill=fillColor] (414.54, 87.56) --
	(418.46, 87.56) --
	(418.46, 91.48) --
	(414.54, 91.48) --
	cycle;

\path[fill=fillColor] (420.10, 87.57) --
	(424.03, 87.57) --
	(424.03, 91.50) --
	(420.10, 91.50) --
	cycle;

\path[fill=fillColor] (425.67, 87.58) --
	(429.59, 87.58) --
	(429.59, 91.51) --
	(425.67, 91.51) --
	cycle;

\path[fill=fillColor] (431.24, 87.64) --
	(435.16, 87.64) --
	(435.16, 91.57) --
	(431.24, 91.57) --
	cycle;

\path[fill=fillColor] (436.80, 87.75) --
	(440.73, 87.75) --
	(440.73, 91.67) --
	(436.80, 91.67) --
	cycle;

\path[fill=fillColor] (442.37, 88.01) --
	(446.29, 88.01) --
	(446.29, 91.94) --
	(442.37, 91.94) --
	cycle;

\path[fill=fillColor] (447.94, 88.69) --
	(451.86, 88.69) --
	(451.86, 92.61) --
	(447.94, 92.61) --
	cycle;

\path[fill=fillColor] (453.50, 90.15) --
	(457.43, 90.15) --
	(457.43, 94.07) --
	(453.50, 94.07) --
	cycle;

\path[fill=fillColor] (459.07, 93.51) --
	(462.99, 93.51) --
	(462.99, 97.43) --
	(459.07, 97.43) --
	cycle;

\path[fill=fillColor] (464.64, 99.11) --
	(468.56, 99.11) --
	(468.56,103.03) --
	(464.64,103.03) --
	cycle;

\path[fill=fillColor] (470.20,106.84) --
	(474.13,106.84) --
	(474.13,110.76) --
	(470.20,110.76) --
	cycle;

\path[fill=fillColor] (475.77,117.50) --
	(479.69,117.50) --
	(479.69,121.43) --
	(475.77,121.43) --
	cycle;

\path[fill=fillColor] (481.34,130.74) --
	(485.26,130.74) --
	(485.26,134.66) --
	(481.34,134.66) --
	cycle;

\path[fill=fillColor] (486.90,142.61) --
	(490.83,142.61) --
	(490.83,146.54) --
	(486.90,146.54) --
	cycle;

\path[fill=fillColor] (492.47,155.35) --
	(496.39,155.35) --
	(496.39,159.27) --
	(492.47,159.27) --
	cycle;
\definecolor{fillColor}{RGB}{217,95,2}

\path[fill=fillColor] (355.27, 69.41) --
	(357.91, 64.83) --
	(352.62, 64.83) --
	cycle;

\path[fill=fillColor] (360.83, 67.66) --
	(363.48, 63.08) --
	(358.19, 63.08) --
	cycle;

\path[fill=fillColor] (366.40, 67.66) --
	(369.04, 63.08) --
	(363.76, 63.08) --
	cycle;

\path[fill=fillColor] (371.97, 67.66) --
	(374.61, 63.08) --
	(369.32, 63.08) --
	cycle;

\path[fill=fillColor] (377.53, 67.66) --
	(380.18, 63.08) --
	(374.89, 63.08) --
	cycle;

\path[fill=fillColor] (383.10, 67.97) --
	(385.74, 63.39) --
	(380.46, 63.39) --
	cycle;

\path[fill=fillColor] (388.67, 67.97) --
	(391.31, 63.39) --
	(386.02, 63.39) --
	cycle;
\definecolor{fillColor}{RGB}{27,158,119}

\path[fill=fillColor] (355.27,114.65) circle (  1.96);

\path[fill=fillColor] (360.83,111.15) circle (  1.96);

\path[fill=fillColor] (366.40,111.49) circle (  1.96);

\path[fill=fillColor] (371.97,111.10) circle (  1.96);

\path[fill=fillColor] (377.53,115.68) circle (  1.96);

\path[fill=fillColor] (383.10,162.17) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (352.49, 88.43) -- (358.04, 88.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (355.27, 85.65) -- (355.27, 91.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (358.06, 88.43) -- (363.61, 88.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (360.83, 85.65) -- (360.83, 91.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (363.62, 88.43) -- (369.17, 88.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (366.40, 85.65) -- (366.40, 91.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (369.19, 88.43) -- (374.74, 88.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (371.97, 85.65) -- (371.97, 91.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (374.76, 88.43) -- (380.31, 88.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (377.53, 85.65) -- (377.53, 91.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (380.32, 88.43) -- (385.87, 88.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (383.10, 85.65) -- (383.10, 91.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (385.89, 88.43) -- (391.44, 88.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (388.67, 85.66) -- (388.67, 91.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (391.46, 88.43) -- (397.01, 88.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (394.23, 85.65) -- (394.23, 91.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (397.02, 88.43) -- (402.57, 88.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (399.80, 85.65) -- (399.80, 91.20);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (402.59, 88.43) -- (408.14, 88.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (405.37, 85.66) -- (405.37, 91.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (408.16, 88.43) -- (413.71, 88.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (410.93, 85.66) -- (410.93, 91.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (413.72, 88.43) -- (419.27, 88.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (416.50, 85.66) -- (416.50, 91.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (419.29, 88.44) -- (424.84, 88.44);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (422.07, 85.67) -- (422.07, 91.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (424.86, 88.46) -- (430.41, 88.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (427.63, 85.68) -- (427.63, 91.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (430.42, 88.51) -- (435.97, 88.51);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (433.20, 85.74) -- (433.20, 91.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (435.99, 88.64) -- (441.54, 88.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (438.77, 85.86) -- (438.77, 91.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (441.56, 88.92) -- (447.11, 88.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (444.33, 86.14) -- (444.33, 91.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (447.12, 89.63) -- (452.67, 89.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (449.90, 86.85) -- (449.90, 92.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (452.69, 91.16) -- (458.24, 91.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (455.47, 88.39) -- (455.47, 93.94);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (458.26, 94.48) -- (463.81, 94.48);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (461.03, 91.70) -- (461.03, 97.25);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (463.82,100.45) -- (469.37,100.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (466.60, 97.67) -- (466.60,103.22);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (469.39,108.29) -- (474.94,108.29);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (472.17,105.51) -- (472.17,111.06);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (474.96,119.14) -- (480.51,119.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (477.73,116.36) -- (477.73,121.91);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (480.52,132.38) -- (486.07,132.38);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (483.30,129.61) -- (483.30,135.16);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (486.09,144.46) -- (491.64,144.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (488.87,141.68) -- (488.87,147.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (491.66,157.43) -- (497.21,157.43);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (494.43,154.66) -- (494.43,160.21);
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (349.70, 57.56) -- (349.70, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (355.27, 57.56) -- (355.27, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (360.83, 57.56) -- (360.83, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (366.40, 57.56) -- (366.40, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (371.97, 57.56) -- (371.97, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (377.53, 57.56) -- (377.53, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (383.10, 57.56) -- (383.10, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (388.67, 57.56) -- (388.67, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (394.23, 57.56) -- (394.23, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (399.80, 57.56) -- (399.80, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (405.37, 57.56) -- (405.37, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (410.93, 57.56) -- (410.93, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (416.50, 57.56) -- (416.50, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (422.07, 57.56) -- (422.07, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (427.63, 57.56) -- (427.63, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (433.20, 57.56) -- (433.20, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (438.77, 57.56) -- (438.77, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (444.33, 57.56) -- (444.33, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (449.90, 57.56) -- (449.90, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (455.47, 57.56) -- (455.47, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (461.03, 57.56) -- (461.03, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (466.60, 57.56) -- (466.60, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (472.17, 57.56) -- (472.17, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (477.73, 57.56) -- (477.73, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (483.30, 57.56) -- (483.30, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (488.87, 57.56) -- (488.87, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (494.43, 57.56) -- (494.43, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (500.00, 57.56) -- (500.00, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31, 64.92) -- (354.00, 64.92);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31, 75.42) -- (354.00, 75.42);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31, 85.92) -- (354.00, 85.92);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31, 96.43) -- (354.00, 96.43);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31,106.93) -- (354.00,106.93);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31,117.43) -- (354.00,117.43);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31,127.94) -- (354.00,127.94);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31,138.44) -- (354.00,138.44);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31,148.95) -- (354.00,148.95);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31,159.45) -- (354.00,159.45);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31,169.95) -- (354.00,169.95);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31,180.46) -- (354.00,180.46);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.31,190.96) -- (354.00,190.96);
\end{scope}
\begin{scope}
\path[clip] ( 33.14,198.75) rectangle (186.23,212.31);
\definecolor{fillColor}{gray}{0.85}

\path[fill=fillColor] ( 33.14,198.75) rectangle (186.23,212.31);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (109.69,203.05) {Bijections};
\end{scope}
\begin{scope}
\path[clip] (190.73,198.75) rectangle (343.81,212.31);
\definecolor{fillColor}{gray}{0.85}

\path[fill=fillColor] (190.73,198.75) rectangle (343.81,212.31);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (267.27,203.05) {Friends};
\end{scope}
\begin{scope}
\path[clip] (348.31,198.75) rectangle (501.39,212.31);
\definecolor{fillColor}{gray}{0.85}

\path[fill=fillColor] (348.31,198.75) rectangle (501.39,212.31);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (424.85,203.05) {Functions};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 62.37, 55.31) --
	( 62.37, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (101.34, 55.31) --
	(101.34, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (140.30, 55.31) --
	(140.30, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (179.27, 55.31) --
	(179.27, 57.56);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 62.37, 48.55) {$2^{5}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (101.34, 48.55) {$2^{12}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (140.30, 48.55) {$2^{19}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (179.27, 48.55) {$2^{26}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] (219.95, 55.31) --
	(219.95, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (258.92, 55.31) --
	(258.92, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (297.88, 55.31) --
	(297.88, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (336.85, 55.31) --
	(336.85, 57.56);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (219.95, 48.55) {$2^{5}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (258.92, 48.55) {$2^{12}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (297.88, 48.55) {$2^{19}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (336.85, 48.55) {$2^{26}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] (377.53, 55.31) --
	(377.53, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (416.50, 55.31) --
	(416.50, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (455.47, 55.31) --
	(455.47, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (494.43, 55.31) --
	(494.43, 57.56);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (377.53, 48.55) {$2^{5}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (416.50, 48.55) {$2^{12}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (455.47, 48.55) {$2^{19}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (494.43, 48.55) {$2^{26}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 29.09, 83.44) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 29.09,114.96) {16};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 29.09,146.47) {128};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 29.09,177.98) {1024};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 30.89, 85.92) --
	( 33.14, 85.92);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 30.89,117.43) --
	( 33.14,117.43);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 30.89,148.95) --
	( 33.14,148.95);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 30.89,180.46) --
	( 33.14,180.46);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (267.27, 38.70) {Domain size};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 10.70,128.16) {Total runtime (s)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] (124.07,  4.50) rectangle (410.47, 27.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (133.07,  9.00) rectangle (147.52, 23.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (134.51, 16.23) -- (146.08, 16.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{fillColor}{RGB}{27,158,119}

\path[fill=fillColor] (140.29, 16.23) circle (  1.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (204.51,  9.00) rectangle (218.96, 23.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (205.95, 16.23) -- (217.52, 16.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{fillColor}{RGB}{217,95,2}

\path[fill=fillColor] (211.73, 19.28) --
	(214.38, 14.70) --
	(209.09, 14.70) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (261.04,  9.00) rectangle (275.49, 23.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] (262.49, 16.23) -- (274.05, 16.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{fillColor}{RGB}{117,112,179}

\path[fill=fillColor] (266.30, 14.26) --
	(270.23, 14.26) --
	(270.23, 18.19) --
	(266.30, 18.19) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (329.62,  9.00) rectangle (344.08, 23.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (331.07, 16.23) -- (342.63, 16.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (334.08, 16.23) -- (339.62, 16.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (336.85, 13.45) -- (336.85, 19.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at (152.02, 13.75) {\textsc{FastWFOMC}};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at (223.46, 13.75) {\textsc{ForcLift}};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at (279.99, 13.75) {\textsc{Gantry-BFS}};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (505.89,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at (348.58, 13.75) {\textsc{Gantry-Greedy}};
\end{scope}
\end{tikzpicture}

\end{document}