### 2025-01-24 doc/paper/plot.tex
Resize the figure

% Created by tikzDevice version 0.12.6 on 2025-01-24 14:00:20
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (361.35,144.54);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.5pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (361.35,144.54);
\end{scope}
\begin{scope}
\path[clip] ( 33.14, 57.56) rectangle (138.05,126.48);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] ( 33.14, 57.56) rectangle (138.05,126.48);
\definecolor{drawColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 33.14, 63.71) --
	(138.05, 63.71);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 33.14, 79.10) --
	(138.05, 79.10);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 33.14, 94.48) --
	(138.05, 94.48);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 33.14,109.86) --
	(138.05,109.86);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 33.14,125.24) --
	(138.05,125.24);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 39.82, 57.56) --
	( 39.82,126.48);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 66.52, 57.56) --
	( 66.52,126.48);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 93.22, 57.56) --
	( 93.22,126.48);

\path[draw=drawColor,line width= 0.2pt,line join=round] (119.93, 57.56) --
	(119.93,126.48);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 33.14, 71.41) --
	(138.05, 71.41);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 33.14, 86.79) --
	(138.05, 86.79);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 33.14,102.17) --
	(138.05,102.17);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 33.14,117.55) --
	(138.05,117.55);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 53.17, 57.56) --
	( 53.17,126.48);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 79.87, 57.56) --
	( 79.87,126.48);

\path[draw=drawColor,line width= 0.5pt,line join=round] (106.58, 57.56) --
	(106.58,126.48);

\path[draw=drawColor,line width= 0.5pt,line join=round] (133.28, 57.56) --
	(133.28,126.48);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.91, 89.74) --
	( 41.73, 84.24) --
	( 45.54, 84.24) --
	( 49.36, 84.34) --
	( 53.17, 87.56) --
	( 56.99,109.01);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 37.91, 81.38) --
	( 41.73, 81.38) --
	( 45.54, 81.38) --
	( 49.36, 81.38) --
	( 53.17, 81.38) --
	( 56.99, 81.38) --
	( 60.80, 81.39) --
	( 64.62, 81.41) --
	( 68.43, 81.49) --
	( 72.24, 81.81) --
	( 76.06, 83.15) --
	( 79.87, 87.02);
\definecolor{fillColor}{RGB}{117,112,179}

\path[fill=fillColor] ( 35.95, 79.42) --
	( 39.87, 79.42) --
	( 39.87, 83.35) --
	( 35.95, 83.35) --
	cycle;

\path[fill=fillColor] ( 39.77, 79.42) --
	( 43.69, 79.42) --
	( 43.69, 83.35) --
	( 39.77, 83.35) --
	cycle;

\path[fill=fillColor] ( 43.58, 79.42) --
	( 47.50, 79.42) --
	( 47.50, 83.35) --
	( 43.58, 83.35) --
	cycle;

\path[fill=fillColor] ( 47.39, 79.42) --
	( 51.32, 79.42) --
	( 51.32, 83.35) --
	( 47.39, 83.35) --
	cycle;

\path[fill=fillColor] ( 51.21, 79.42) --
	( 55.13, 79.42) --
	( 55.13, 83.35) --
	( 51.21, 83.35) --
	cycle;

\path[fill=fillColor] ( 55.02, 79.42) --
	( 58.95, 79.42) --
	( 58.95, 83.35) --
	( 55.02, 83.35) --
	cycle;

\path[fill=fillColor] ( 58.84, 79.43) --
	( 62.76, 79.43) --
	( 62.76, 83.35) --
	( 58.84, 83.35) --
	cycle;

\path[fill=fillColor] ( 62.65, 79.45) --
	( 66.58, 79.45) --
	( 66.58, 83.37) --
	( 62.65, 83.37) --
	cycle;

\path[fill=fillColor] ( 66.47, 79.53) --
	( 70.39, 79.53) --
	( 70.39, 83.45) --
	( 66.47, 83.45) --
	cycle;

\path[fill=fillColor] ( 70.28, 79.84) --
	( 74.21, 79.84) --
	( 74.21, 83.77) --
	( 70.28, 83.77) --
	cycle;

\path[fill=fillColor] ( 74.10, 81.19) --
	( 78.02, 81.19) --
	( 78.02, 85.11) --
	( 74.10, 85.11) --
	cycle;

\path[fill=fillColor] ( 77.91, 85.05) --
	( 81.84, 85.05) --
	( 81.84, 88.98) --
	( 77.91, 88.98) --
	cycle;
\definecolor{fillColor}{RGB}{27,158,119}

\path[fill=fillColor] ( 37.91, 89.74) circle (  1.96);

\path[fill=fillColor] ( 41.73, 84.24) circle (  1.96);

\path[fill=fillColor] ( 45.54, 84.24) circle (  1.96);

\path[fill=fillColor] ( 49.36, 84.34) circle (  1.96);

\path[fill=fillColor] ( 53.17, 87.56) circle (  1.96);

\path[fill=fillColor] ( 56.99,109.01) circle (  1.96);
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 34.10, 57.56) -- ( 34.10, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 37.91, 57.56) -- ( 37.91, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 41.73, 57.56) -- ( 41.73, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 45.54, 57.56) -- ( 45.54, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 49.36, 57.56) -- ( 49.36, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 53.17, 57.56) -- ( 53.17, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 56.99, 57.56) -- ( 56.99, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 60.80, 57.56) -- ( 60.80, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 64.62, 57.56) -- ( 64.62, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 68.43, 57.56) -- ( 68.43, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 72.24, 57.56) -- ( 72.24, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 76.06, 57.56) -- ( 76.06, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 79.87, 57.56) -- ( 79.87, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 83.69, 57.56) -- ( 83.69, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 87.50, 57.56) -- ( 87.50, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 91.32, 57.56) -- ( 91.32, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 95.13, 57.56) -- ( 95.13, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 98.95, 57.56) -- ( 98.95, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (102.76, 57.56) -- (102.76, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (106.58, 57.56) -- (106.58, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (110.39, 57.56) -- (110.39, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (114.21, 57.56) -- (114.21, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (118.02, 57.56) -- (118.02, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (121.83, 57.56) -- (121.83, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (125.65, 57.56) -- (125.65, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (129.46, 57.56) -- (129.46, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (133.28, 57.56) -- (133.28, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (137.09, 57.56) -- (137.09, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14, 61.15) -- ( 38.84, 61.15);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14, 66.28) -- ( 38.84, 66.28);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14, 71.41) -- ( 38.84, 71.41);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14, 76.53) -- ( 38.84, 76.53);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14, 81.66) -- ( 38.84, 81.66);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14, 86.79) -- ( 38.84, 86.79);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14, 91.91) -- ( 38.84, 91.91);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14, 97.04) -- ( 38.84, 97.04);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,102.17) -- ( 38.84,102.17);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,107.30) -- ( 38.84,107.30);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,112.42) -- ( 38.84,112.42);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,117.55) -- ( 38.84,117.55);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,122.68) -- ( 38.84,122.68);
\end{scope}
\begin{scope}
\path[clip] (142.55, 57.56) rectangle (247.45,126.48);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (142.55, 57.56) rectangle (247.45,126.48);
\definecolor{drawColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.2pt,line join=round] (142.55, 63.71) --
	(247.45, 63.71);

\path[draw=drawColor,line width= 0.2pt,line join=round] (142.55, 79.10) --
	(247.45, 79.10);

\path[draw=drawColor,line width= 0.2pt,line join=round] (142.55, 94.48) --
	(247.45, 94.48);

\path[draw=drawColor,line width= 0.2pt,line join=round] (142.55,109.86) --
	(247.45,109.86);

\path[draw=drawColor,line width= 0.2pt,line join=round] (142.55,125.24) --
	(247.45,125.24);

\path[draw=drawColor,line width= 0.2pt,line join=round] (149.22, 57.56) --
	(149.22,126.48);

\path[draw=drawColor,line width= 0.2pt,line join=round] (175.92, 57.56) --
	(175.92,126.48);

\path[draw=drawColor,line width= 0.2pt,line join=round] (202.63, 57.56) --
	(202.63,126.48);

\path[draw=drawColor,line width= 0.2pt,line join=round] (229.33, 57.56) --
	(229.33,126.48);

\path[draw=drawColor,line width= 0.5pt,line join=round] (142.55, 71.41) --
	(247.45, 71.41);

\path[draw=drawColor,line width= 0.5pt,line join=round] (142.55, 86.79) --
	(247.45, 86.79);

\path[draw=drawColor,line width= 0.5pt,line join=round] (142.55,102.17) --
	(247.45,102.17);

\path[draw=drawColor,line width= 0.5pt,line join=round] (142.55,117.55) --
	(247.45,117.55);

\path[draw=drawColor,line width= 0.5pt,line join=round] (162.57, 57.56) --
	(162.57,126.48);

\path[draw=drawColor,line width= 0.5pt,line join=round] (189.28, 57.56) --
	(189.28,126.48);

\path[draw=drawColor,line width= 0.5pt,line join=round] (215.98, 57.56) --
	(215.98,126.48);

\path[draw=drawColor,line width= 0.5pt,line join=round] (242.68, 57.56) --
	(242.68,126.48);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (147.31, 84.74) --
	(151.13, 84.06) --
	(154.94, 83.93) --
	(158.76, 83.97) --
	(162.57, 84.20) --
	(166.39, 84.27) --
	(170.20, 84.38) --
	(174.02, 86.20) --
	(177.83, 93.38) --
	(181.65,111.72);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (147.31, 62.86) --
	(151.13, 60.69) --
	(154.94, 60.69) --
	(158.76, 60.69);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] (147.31, 72.23) --
	(151.13, 72.23) --
	(154.94, 72.23) --
	(158.76, 72.23) --
	(162.57, 72.23) --
	(166.39, 72.23) --
	(170.20, 72.23) --
	(174.02, 72.27) --
	(177.83, 72.70) --
	(181.65, 76.95) --
	(185.46, 89.66) --
	(189.28,106.37) --
	(193.09,123.35);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (147.31, 72.11) --
	(151.13, 72.11) --
	(154.94, 72.11) --
	(158.76, 72.11) --
	(162.57, 72.11) --
	(166.39, 72.11) --
	(170.20, 72.11) --
	(174.02, 72.15) --
	(177.83, 72.59) --
	(181.65, 76.89) --
	(185.46, 89.67) --
	(189.28,106.37) --
	(193.09,123.35);
\definecolor{fillColor}{RGB}{27,158,119}

\path[fill=fillColor] (147.31, 84.74) circle (  1.96);

\path[fill=fillColor] (151.13, 84.06) circle (  1.96);

\path[fill=fillColor] (154.94, 83.93) circle (  1.96);

\path[fill=fillColor] (158.76, 83.97) circle (  1.96);

\path[fill=fillColor] (162.57, 84.20) circle (  1.96);

\path[fill=fillColor] (166.39, 84.27) circle (  1.96);

\path[fill=fillColor] (170.20, 84.38) circle (  1.96);

\path[fill=fillColor] (174.02, 86.20) circle (  1.96);

\path[fill=fillColor] (177.83, 93.38) circle (  1.96);

\path[fill=fillColor] (181.65,111.72) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (144.54, 72.11) -- (150.09, 72.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (147.31, 69.34) -- (147.31, 74.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (148.35, 72.11) -- (153.90, 72.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.13, 69.33) -- (151.13, 74.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (152.17, 72.11) -- (157.72, 72.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (154.94, 69.33) -- (154.94, 74.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (155.98, 72.11) -- (161.53, 72.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (158.76, 69.33) -- (158.76, 74.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (159.80, 72.11) -- (165.35, 72.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (162.57, 69.33) -- (162.57, 74.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (163.61, 72.11) -- (169.16, 72.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (166.39, 69.33) -- (166.39, 74.88);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (167.43, 72.11) -- (172.98, 72.11);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (170.20, 69.34) -- (170.20, 74.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (171.24, 72.15) -- (176.79, 72.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (174.02, 69.37) -- (174.02, 74.92);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (175.06, 72.59) -- (180.61, 72.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (177.83, 69.81) -- (177.83, 75.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (178.87, 76.89) -- (184.42, 76.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (181.65, 74.12) -- (181.65, 79.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (182.69, 89.67) -- (188.24, 89.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (185.46, 86.90) -- (185.46, 92.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (186.50,106.37) -- (192.05,106.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (189.28,103.59) -- (189.28,109.14);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (190.32,123.35) -- (195.86,123.35);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (193.09,120.57) -- (193.09,126.12);
\definecolor{fillColor}{RGB}{217,95,2}

\path[fill=fillColor] (147.31, 65.91) --
	(149.96, 61.33) --
	(144.67, 61.33) --
	cycle;

\path[fill=fillColor] (151.13, 63.74) --
	(153.77, 59.17) --
	(148.49, 59.17) --
	cycle;

\path[fill=fillColor] (154.94, 63.74) --
	(157.59, 59.17) --
	(152.30, 59.17) --
	cycle;

\path[fill=fillColor] (158.76, 63.74) --
	(161.40, 59.17) --
	(156.12, 59.17) --
	cycle;
\definecolor{fillColor}{RGB}{117,112,179}

\path[fill=fillColor] (145.35, 70.27) --
	(149.28, 70.27) --
	(149.28, 74.20) --
	(145.35, 74.20) --
	cycle;

\path[fill=fillColor] (149.17, 70.27) --
	(153.09, 70.27) --
	(153.09, 74.19) --
	(149.17, 74.19) --
	cycle;

\path[fill=fillColor] (152.98, 70.27) --
	(156.91, 70.27) --
	(156.91, 74.19) --
	(152.98, 74.19) --
	cycle;

\path[fill=fillColor] (156.80, 70.27) --
	(160.72, 70.27) --
	(160.72, 74.19) --
	(156.80, 74.19) --
	cycle;

\path[fill=fillColor] (160.61, 70.27) --
	(164.54, 70.27) --
	(164.54, 74.19) --
	(160.61, 74.19) --
	cycle;

\path[fill=fillColor] (164.43, 70.27) --
	(168.35, 70.27) --
	(168.35, 74.19) --
	(164.43, 74.19) --
	cycle;

\path[fill=fillColor] (168.24, 70.27) --
	(172.16, 70.27) --
	(172.16, 74.20) --
	(168.24, 74.20) --
	cycle;

\path[fill=fillColor] (172.05, 70.31) --
	(175.98, 70.31) --
	(175.98, 74.23) --
	(172.05, 74.23) --
	cycle;

\path[fill=fillColor] (175.87, 70.74) --
	(179.79, 70.74) --
	(179.79, 74.66) --
	(175.87, 74.66) --
	cycle;

\path[fill=fillColor] (179.68, 74.99) --
	(183.61, 74.99) --
	(183.61, 78.91) --
	(179.68, 78.91) --
	cycle;

\path[fill=fillColor] (183.50, 87.70) --
	(187.42, 87.70) --
	(187.42, 91.62) --
	(183.50, 91.62) --
	cycle;

\path[fill=fillColor] (187.31,104.41) --
	(191.24,104.41) --
	(191.24,108.33) --
	(187.31,108.33) --
	cycle;

\path[fill=fillColor] (191.13,121.39) --
	(195.05,121.39) --
	(195.05,125.31) --
	(191.13,125.31) --
	cycle;
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (143.50, 57.56) -- (143.50, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (147.31, 57.56) -- (147.31, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (151.13, 57.56) -- (151.13, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (154.94, 57.56) -- (154.94, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (158.76, 57.56) -- (158.76, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (162.57, 57.56) -- (162.57, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (166.39, 57.56) -- (166.39, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (170.20, 57.56) -- (170.20, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (174.02, 57.56) -- (174.02, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (177.83, 57.56) -- (177.83, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (181.65, 57.56) -- (181.65, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (185.46, 57.56) -- (185.46, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (189.28, 57.56) -- (189.28, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (193.09, 57.56) -- (193.09, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (196.90, 57.56) -- (196.90, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (200.72, 57.56) -- (200.72, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (204.53, 57.56) -- (204.53, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (208.35, 57.56) -- (208.35, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (212.16, 57.56) -- (212.16, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (215.98, 57.56) -- (215.98, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (219.79, 57.56) -- (219.79, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (223.61, 57.56) -- (223.61, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (227.42, 57.56) -- (227.42, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (231.24, 57.56) -- (231.24, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (235.05, 57.56) -- (235.05, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (238.87, 57.56) -- (238.87, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (242.68, 57.56) -- (242.68, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (246.49, 57.56) -- (246.49, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (142.55, 61.15) -- (148.24, 61.15);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (142.55, 66.28) -- (148.24, 66.28);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (142.55, 71.41) -- (148.24, 71.41);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (142.55, 76.53) -- (148.24, 76.53);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (142.55, 81.66) -- (148.24, 81.66);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (142.55, 86.79) -- (148.24, 86.79);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (142.55, 91.91) -- (148.24, 91.91);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (142.55, 97.04) -- (148.24, 97.04);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (142.55,102.17) -- (148.24,102.17);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (142.55,107.30) -- (148.24,107.30);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (142.55,112.42) -- (148.24,112.42);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (142.55,117.55) -- (148.24,117.55);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (142.55,122.68) -- (148.24,122.68);
\end{scope}
\begin{scope}
\path[clip] (251.95, 57.56) rectangle (356.85,126.48);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (251.95, 57.56) rectangle (356.85,126.48);
\definecolor{drawColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.2pt,line join=round] (251.95, 63.71) --
	(356.85, 63.71);

\path[draw=drawColor,line width= 0.2pt,line join=round] (251.95, 79.10) --
	(356.85, 79.10);

\path[draw=drawColor,line width= 0.2pt,line join=round] (251.95, 94.48) --
	(356.85, 94.48);

\path[draw=drawColor,line width= 0.2pt,line join=round] (251.95,109.86) --
	(356.85,109.86);

\path[draw=drawColor,line width= 0.2pt,line join=round] (251.95,125.24) --
	(356.85,125.24);

\path[draw=drawColor,line width= 0.2pt,line join=round] (258.62, 57.56) --
	(258.62,126.48);

\path[draw=drawColor,line width= 0.2pt,line join=round] (285.33, 57.56) --
	(285.33,126.48);

\path[draw=drawColor,line width= 0.2pt,line join=round] (312.03, 57.56) --
	(312.03,126.48);

\path[draw=drawColor,line width= 0.2pt,line join=round] (338.73, 57.56) --
	(338.73,126.48);

\path[draw=drawColor,line width= 0.5pt,line join=round] (251.95, 71.41) --
	(356.85, 71.41);

\path[draw=drawColor,line width= 0.5pt,line join=round] (251.95, 86.79) --
	(356.85, 86.79);

\path[draw=drawColor,line width= 0.5pt,line join=round] (251.95,102.17) --
	(356.85,102.17);

\path[draw=drawColor,line width= 0.5pt,line join=round] (251.95,117.55) --
	(356.85,117.55);

\path[draw=drawColor,line width= 0.5pt,line join=round] (271.97, 57.56) --
	(271.97,126.48);

\path[draw=drawColor,line width= 0.5pt,line join=round] (298.68, 57.56) --
	(298.68,126.48);

\path[draw=drawColor,line width= 0.5pt,line join=round] (325.38, 57.56) --
	(325.38,126.48);

\path[draw=drawColor,line width= 0.5pt,line join=round] (352.08, 57.56) --
	(352.08,126.48);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (256.72, 85.43) --
	(260.53, 83.72) --
	(264.35, 83.89) --
	(268.16, 83.69) --
	(271.97, 85.93) --
	(275.79,108.62);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (256.72, 61.86) --
	(260.53, 61.00) --
	(264.35, 61.00) --
	(268.16, 61.00) --
	(271.97, 61.00) --
	(275.79, 61.15) --
	(279.60, 61.15);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] (256.72, 73.16) --
	(260.53, 73.16) --
	(264.35, 73.16) --
	(268.16, 73.16) --
	(271.97, 73.16) --
	(275.79, 73.16) --
	(279.60, 73.16) --
	(283.42, 73.16) --
	(287.23, 73.16) --
	(291.05, 73.16) --
	(294.86, 73.16) --
	(298.68, 73.16) --
	(302.49, 73.17) --
	(306.31, 73.17) --
	(310.12, 73.20) --
	(313.94, 73.25) --
	(317.75, 73.38) --
	(321.56, 73.71) --
	(325.38, 74.42) --
	(329.19, 76.07) --
	(333.01, 78.80) --
	(336.82, 82.57) --
	(340.64, 87.78) --
	(344.45, 94.24) --
	(348.27,100.04) --
	(352.08,106.25);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (256.72, 72.63) --
	(260.53, 72.63) --
	(264.35, 72.63) --
	(268.16, 72.63) --
	(271.97, 72.63) --
	(275.79, 72.63) --
	(279.60, 72.63) --
	(283.42, 72.63) --
	(287.23, 72.63) --
	(291.05, 72.63) --
	(294.86, 72.63) --
	(298.68, 72.63) --
	(302.49, 72.64) --
	(306.31, 72.64) --
	(310.12, 72.67) --
	(313.94, 72.73) --
	(317.75, 72.87) --
	(321.56, 73.21) --
	(325.38, 73.96) --
	(329.19, 75.58) --
	(333.01, 78.49) --
	(336.82, 82.32) --
	(340.64, 87.62) --
	(344.45, 94.08) --
	(348.27, 99.98) --
	(352.08,106.31);
\definecolor{fillColor}{RGB}{117,112,179}

\path[fill=fillColor] (254.75, 71.20) --
	(258.68, 71.20) --
	(258.68, 75.12) --
	(254.75, 75.12) --
	cycle;

\path[fill=fillColor] (258.57, 71.20) --
	(262.49, 71.20) --
	(262.49, 75.12) --
	(258.57, 75.12) --
	cycle;

\path[fill=fillColor] (262.38, 71.20) --
	(266.31, 71.20) --
	(266.31, 75.12) --
	(262.38, 75.12) --
	cycle;

\path[fill=fillColor] (266.20, 71.20) --
	(270.12, 71.20) --
	(270.12, 75.12) --
	(266.20, 75.12) --
	cycle;

\path[fill=fillColor] (270.01, 71.20) --
	(273.94, 71.20) --
	(273.94, 75.12) --
	(270.01, 75.12) --
	cycle;

\path[fill=fillColor] (273.83, 71.20) --
	(277.75, 71.20) --
	(277.75, 75.12) --
	(273.83, 75.12) --
	cycle;

\path[fill=fillColor] (277.64, 71.20) --
	(281.57, 71.20) --
	(281.57, 75.12) --
	(277.64, 75.12) --
	cycle;

\path[fill=fillColor] (281.46, 71.20) --
	(285.38, 71.20) --
	(285.38, 75.12) --
	(281.46, 75.12) --
	cycle;

\path[fill=fillColor] (285.27, 71.20) --
	(289.20, 71.20) --
	(289.20, 75.12) --
	(285.27, 75.12) --
	cycle;

\path[fill=fillColor] (289.09, 71.20) --
	(293.01, 71.20) --
	(293.01, 75.12) --
	(289.09, 75.12) --
	cycle;

\path[fill=fillColor] (292.90, 71.20) --
	(296.82, 71.20) --
	(296.82, 75.12) --
	(292.90, 75.12) --
	cycle;

\path[fill=fillColor] (296.72, 71.20) --
	(300.64, 71.20) --
	(300.64, 75.12) --
	(296.72, 75.12) --
	cycle;

\path[fill=fillColor] (300.53, 71.21) --
	(304.45, 71.21) --
	(304.45, 75.13) --
	(300.53, 75.13) --
	cycle;

\path[fill=fillColor] (304.34, 71.21) --
	(308.27, 71.21) --
	(308.27, 75.14) --
	(304.34, 75.14) --
	cycle;

\path[fill=fillColor] (308.16, 71.24) --
	(312.08, 71.24) --
	(312.08, 75.16) --
	(308.16, 75.16) --
	cycle;

\path[fill=fillColor] (311.97, 71.29) --
	(315.90, 71.29) --
	(315.90, 75.22) --
	(311.97, 75.22) --
	cycle;

\path[fill=fillColor] (315.79, 71.42) --
	(319.71, 71.42) --
	(319.71, 75.35) --
	(315.79, 75.35) --
	cycle;

\path[fill=fillColor] (319.60, 71.75) --
	(323.53, 71.75) --
	(323.53, 75.67) --
	(319.60, 75.67) --
	cycle;

\path[fill=fillColor] (323.42, 72.46) --
	(327.34, 72.46) --
	(327.34, 76.39) --
	(323.42, 76.39) --
	cycle;

\path[fill=fillColor] (327.23, 74.10) --
	(331.16, 74.10) --
	(331.16, 78.03) --
	(327.23, 78.03) --
	cycle;

\path[fill=fillColor] (331.05, 76.84) --
	(334.97, 76.84) --
	(334.97, 80.76) --
	(331.05, 80.76) --
	cycle;

\path[fill=fillColor] (334.86, 80.61) --
	(338.79, 80.61) --
	(338.79, 84.53) --
	(334.86, 84.53) --
	cycle;

\path[fill=fillColor] (338.68, 85.82) --
	(342.60, 85.82) --
	(342.60, 89.74) --
	(338.68, 89.74) --
	cycle;

\path[fill=fillColor] (342.49, 92.28) --
	(346.41, 92.28) --
	(346.41, 96.20) --
	(342.49, 96.20) --
	cycle;

\path[fill=fillColor] (346.30, 98.07) --
	(350.23, 98.07) --
	(350.23,102.00) --
	(346.30,102.00) --
	cycle;

\path[fill=fillColor] (350.12,104.29) --
	(354.04,104.29) --
	(354.04,108.21) --
	(350.12,108.21) --
	cycle;
\definecolor{fillColor}{RGB}{217,95,2}

\path[fill=fillColor] (256.72, 64.91) --
	(259.36, 60.33) --
	(254.07, 60.33) --
	cycle;

\path[fill=fillColor] (260.53, 64.05) --
	(263.17, 59.48) --
	(257.89, 59.48) --
	cycle;

\path[fill=fillColor] (264.35, 64.05) --
	(266.99, 59.48) --
	(261.70, 59.48) --
	cycle;

\path[fill=fillColor] (268.16, 64.05) --
	(270.80, 59.48) --
	(265.52, 59.48) --
	cycle;

\path[fill=fillColor] (271.97, 64.05) --
	(274.62, 59.48) --
	(269.33, 59.48) --
	cycle;

\path[fill=fillColor] (275.79, 64.20) --
	(278.43, 59.63) --
	(273.15, 59.63) --
	cycle;

\path[fill=fillColor] (279.60, 64.20) --
	(282.25, 59.63) --
	(276.96, 59.63) --
	cycle;
\definecolor{fillColor}{RGB}{27,158,119}

\path[fill=fillColor] (256.72, 85.43) circle (  1.96);

\path[fill=fillColor] (260.53, 83.72) circle (  1.96);

\path[fill=fillColor] (264.35, 83.89) circle (  1.96);

\path[fill=fillColor] (268.16, 83.69) circle (  1.96);

\path[fill=fillColor] (271.97, 85.93) circle (  1.96);

\path[fill=fillColor] (275.79,108.62) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (253.94, 72.63) -- (259.49, 72.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (256.72, 69.85) -- (256.72, 75.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (257.76, 72.63) -- (263.31, 72.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (260.53, 69.85) -- (260.53, 75.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.57, 72.63) -- (267.12, 72.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (264.35, 69.85) -- (264.35, 75.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (265.39, 72.63) -- (270.94, 72.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (268.16, 69.85) -- (268.16, 75.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (269.20, 72.63) -- (274.75, 72.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (271.97, 69.85) -- (271.97, 75.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (273.01, 72.63) -- (278.56, 72.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (275.79, 69.85) -- (275.79, 75.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (276.83, 72.63) -- (282.38, 72.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (279.60, 69.86) -- (279.60, 75.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (280.64, 72.63) -- (286.19, 72.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (283.42, 69.85) -- (283.42, 75.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (284.46, 72.63) -- (290.01, 72.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (287.23, 69.85) -- (287.23, 75.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (288.27, 72.63) -- (293.82, 72.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (291.05, 69.86) -- (291.05, 75.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (292.09, 72.63) -- (297.64, 72.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (294.86, 69.86) -- (294.86, 75.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (295.90, 72.63) -- (301.45, 72.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (298.68, 69.86) -- (298.68, 75.40);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (299.72, 72.64) -- (305.27, 72.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (302.49, 69.86) -- (302.49, 75.41);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (303.53, 72.64) -- (309.08, 72.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (306.31, 69.87) -- (306.31, 75.42);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (307.35, 72.67) -- (312.90, 72.67);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (310.12, 69.90) -- (310.12, 75.45);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (311.16, 72.73) -- (316.71, 72.73);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (313.94, 69.95) -- (313.94, 75.50);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (314.98, 72.87) -- (320.53, 72.87);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (317.75, 70.09) -- (317.75, 75.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (318.79, 73.21) -- (324.34, 73.21);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (321.56, 70.44) -- (321.56, 75.99);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (322.60, 73.96) -- (328.15, 73.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (325.38, 71.19) -- (325.38, 76.74);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (326.42, 75.58) -- (331.97, 75.58);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (329.19, 72.81) -- (329.19, 78.36);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (330.23, 78.49) -- (335.78, 78.49);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (333.01, 75.72) -- (333.01, 81.27);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (334.05, 82.32) -- (339.60, 82.32);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (336.82, 79.55) -- (336.82, 85.10);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (337.86, 87.62) -- (343.41, 87.62);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (340.64, 84.84) -- (340.64, 90.39);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (341.68, 94.08) -- (347.23, 94.08);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (344.45, 91.31) -- (344.45, 96.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (345.49, 99.98) -- (351.04, 99.98);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (348.27, 97.20) -- (348.27,102.75);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (349.31,106.31) -- (354.86,106.31);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (352.08,103.54) -- (352.08,109.09);
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (252.90, 57.56) -- (252.90, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (256.72, 57.56) -- (256.72, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (260.53, 57.56) -- (260.53, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (264.35, 57.56) -- (264.35, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (268.16, 57.56) -- (268.16, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (271.97, 57.56) -- (271.97, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (275.79, 57.56) -- (275.79, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (279.60, 57.56) -- (279.60, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (283.42, 57.56) -- (283.42, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (287.23, 57.56) -- (287.23, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (291.05, 57.56) -- (291.05, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (294.86, 57.56) -- (294.86, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (298.68, 57.56) -- (298.68, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.49, 57.56) -- (302.49, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (306.31, 57.56) -- (306.31, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (310.12, 57.56) -- (310.12, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (313.94, 57.56) -- (313.94, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (317.75, 57.56) -- (317.75, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (321.56, 57.56) -- (321.56, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (325.38, 57.56) -- (325.38, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (329.19, 57.56) -- (329.19, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (333.01, 57.56) -- (333.01, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (336.82, 57.56) -- (336.82, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (340.64, 57.56) -- (340.64, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (344.45, 57.56) -- (344.45, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.27, 57.56) -- (348.27, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (352.08, 57.56) -- (352.08, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (355.90, 57.56) -- (355.90, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.95, 61.15) -- (257.64, 61.15);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.95, 66.28) -- (257.64, 66.28);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.95, 71.41) -- (257.64, 71.41);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.95, 76.53) -- (257.64, 76.53);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.95, 81.66) -- (257.64, 81.66);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.95, 86.79) -- (257.64, 86.79);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.95, 91.91) -- (257.64, 91.91);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.95, 97.04) -- (257.64, 97.04);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.95,102.17) -- (257.64,102.17);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.95,107.30) -- (257.64,107.30);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.95,112.42) -- (257.64,112.42);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.95,117.55) -- (257.64,117.55);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.95,122.68) -- (257.64,122.68);
\end{scope}
\begin{scope}
\path[clip] ( 33.14,126.48) rectangle (138.05,140.04);
\definecolor{fillColor}{gray}{0.85}

\path[fill=fillColor] ( 33.14,126.48) rectangle (138.05,140.04);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 85.60,130.78) {Bijections};
\end{scope}
\begin{scope}
\path[clip] (142.55,126.48) rectangle (247.45,140.04);
\definecolor{fillColor}{gray}{0.85}

\path[fill=fillColor] (142.55,126.48) rectangle (247.45,140.04);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (195.00,130.78) {Friends};
\end{scope}
\begin{scope}
\path[clip] (251.95,126.48) rectangle (356.85,140.04);
\definecolor{fillColor}{gray}{0.85}

\path[fill=fillColor] (251.95,126.48) rectangle (356.85,140.04);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (304.40,130.78) {Functions};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 53.17, 55.31) --
	( 53.17, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 79.87, 55.31) --
	( 79.87, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (106.58, 55.31) --
	(106.58, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (133.28, 55.31) --
	(133.28, 57.56);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 53.17, 48.55) {$2^{5}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 79.87, 48.55) {$2^{12}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (106.58, 48.55) {$2^{19}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (133.28, 48.55) {$2^{26}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] (162.57, 55.31) --
	(162.57, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (189.28, 55.31) --
	(189.28, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (215.98, 55.31) --
	(215.98, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (242.68, 55.31) --
	(242.68, 57.56);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (162.57, 48.55) {$2^{5}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (189.28, 48.55) {$2^{12}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (215.98, 48.55) {$2^{19}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (242.68, 48.55) {$2^{26}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] (271.97, 55.31) --
	(271.97, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (298.68, 55.31) --
	(298.68, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (325.38, 55.31) --
	(325.38, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (352.08, 55.31) --
	(352.08, 57.56);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (271.97, 48.55) {$2^{5}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (298.68, 48.55) {$2^{12}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (325.38, 48.55) {$2^{19}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (352.08, 48.55) {$2^{26}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 29.09, 68.93) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 29.09, 84.31) {16};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 29.09, 99.69) {128};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 29.09,115.07) {1024};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 30.89, 71.41) --
	( 33.14, 71.41);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 30.89, 86.79) --
	( 33.14, 86.79);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 30.89,102.17) --
	( 33.14,102.17);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 30.89,117.55) --
	( 33.14,117.55);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (195.00, 38.70) {Domain size};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 10.70, 92.02) {Total runtime (s)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 51.80,  4.50) rectangle (338.20, 27.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] ( 60.80,  9.00) rectangle ( 75.25, 23.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 62.24, 16.23) -- ( 73.81, 16.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{fillColor}{RGB}{27,158,119}

\path[fill=fillColor] ( 68.02, 16.23) circle (  1.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (132.24,  9.00) rectangle (146.69, 23.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (133.68, 16.23) -- (145.25, 16.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{fillColor}{RGB}{217,95,2}

\path[fill=fillColor] (139.46, 19.28) --
	(142.11, 14.70) --
	(136.82, 14.70) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (188.77,  9.00) rectangle (203.22, 23.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] (190.22, 16.23) -- (201.78, 16.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{fillColor}{RGB}{117,112,179}

\path[fill=fillColor] (194.03, 14.26) --
	(197.96, 14.26) --
	(197.96, 18.19) --
	(194.03, 18.19) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (257.35,  9.00) rectangle (271.81, 23.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (258.80, 16.23) -- (270.36, 16.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.81, 16.23) -- (267.35, 16.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (264.58, 13.45) -- (264.58, 19.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 79.75, 13.75) {\textsc{FastWFOMC}};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at (151.19, 13.75) {\textsc{ForcLift}};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at (207.72, 13.75) {\textsc{Gantry-BFS}};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at (276.31, 13.75) {\textsc{Gantry-Greedy}};
\end{scope}
\end{tikzpicture}

\end{document}