### 2025-06-05 doc/paper/plot.tex
Potentially the final (camera-ready) version

% Created by tikzDevice version 0.12.6 on 2025-06-03 07:37:32
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (361.35,144.54);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.5pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (361.35,144.54);
\end{scope}
\begin{scope}
\path[clip] ( 33.14, 57.56) rectangle (138.05,126.48);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] ( 33.14, 57.56) rectangle (138.05,126.48);
\definecolor{drawColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 33.14, 63.71) --
	(138.05, 63.71);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 33.14, 79.10) --
	(138.05, 79.10);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 33.14, 94.48) --
	(138.05, 94.48);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 33.14,109.86) --
	(138.05,109.86);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 33.14,125.24) --
	(138.05,125.24);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 39.82, 57.56) --
	( 39.82,126.48);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 66.52, 57.56) --
	( 66.52,126.48);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 93.22, 57.56) --
	( 93.22,126.48);

\path[draw=drawColor,line width= 0.2pt,line join=round] (119.93, 57.56) --
	(119.93,126.48);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 33.14, 71.41) --
	(138.05, 71.41);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 33.14, 86.79) --
	(138.05, 86.79);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 33.14,102.17) --
	(138.05,102.17);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 33.14,117.55) --
	(138.05,117.55);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 53.17, 57.56) --
	( 53.17,126.48);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 79.87, 57.56) --
	( 79.87,126.48);

\path[draw=drawColor,line width= 0.5pt,line join=round] (106.58, 57.56) --
	(106.58,126.48);

\path[draw=drawColor,line width= 0.5pt,line join=round] (133.28, 57.56) --
	(133.28,126.48);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 37.91, 81.38) --
	( 41.73, 81.38) --
	( 45.54, 81.38) --
	( 49.36, 81.38) --
	( 53.17, 81.38) --
	( 56.99, 81.38) --
	( 60.80, 81.39) --
	( 64.62, 81.41) --
	( 68.43, 81.49) --
	( 72.24, 81.81) --
	( 76.06, 83.15) --
	( 79.87, 87.02);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 37.91, 89.74) --
	( 41.73, 84.24) --
	( 45.54, 84.24) --
	( 49.36, 84.34) --
	( 53.17, 87.56) --
	( 56.99,109.01);
\definecolor{fillColor}{RGB}{27,158,119}

\path[fill=fillColor] ( 37.91, 81.38) circle (  1.96);

\path[fill=fillColor] ( 41.73, 81.38) circle (  1.96);

\path[fill=fillColor] ( 45.54, 81.38) circle (  1.96);

\path[fill=fillColor] ( 49.36, 81.38) circle (  1.96);

\path[fill=fillColor] ( 53.17, 81.38) circle (  1.96);

\path[fill=fillColor] ( 56.99, 81.38) circle (  1.96);

\path[fill=fillColor] ( 60.80, 81.39) circle (  1.96);

\path[fill=fillColor] ( 64.62, 81.41) circle (  1.96);

\path[fill=fillColor] ( 68.43, 81.49) circle (  1.96);

\path[fill=fillColor] ( 72.24, 81.81) circle (  1.96);

\path[fill=fillColor] ( 76.06, 83.15) circle (  1.96);

\path[fill=fillColor] ( 79.87, 87.02) circle (  1.96);
\definecolor{fillColor}{RGB}{117,112,179}

\path[fill=fillColor] ( 35.95, 87.78) --
	( 39.87, 87.78) --
	( 39.87, 91.70) --
	( 35.95, 91.70) --
	cycle;

\path[fill=fillColor] ( 39.77, 82.28) --
	( 43.69, 82.28) --
	( 43.69, 86.20) --
	( 39.77, 86.20) --
	cycle;

\path[fill=fillColor] ( 43.58, 82.28) --
	( 47.50, 82.28) --
	( 47.50, 86.20) --
	( 43.58, 86.20) --
	cycle;

\path[fill=fillColor] ( 47.39, 82.38) --
	( 51.32, 82.38) --
	( 51.32, 86.30) --
	( 47.39, 86.30) --
	cycle;

\path[fill=fillColor] ( 51.21, 85.60) --
	( 55.13, 85.60) --
	( 55.13, 89.53) --
	( 51.21, 89.53) --
	cycle;

\path[fill=fillColor] ( 55.02,107.05) --
	( 58.95,107.05) --
	( 58.95,110.97) --
	( 55.02,110.97) --
	cycle;
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 34.10, 57.56) -- ( 34.10, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 37.91, 57.56) -- ( 37.91, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 41.73, 57.56) -- ( 41.73, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 45.54, 57.56) -- ( 45.54, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 49.36, 57.56) -- ( 49.36, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 53.17, 57.56) -- ( 53.17, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 56.99, 57.56) -- ( 56.99, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 60.80, 57.56) -- ( 60.80, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 64.62, 57.56) -- ( 64.62, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 68.43, 57.56) -- ( 68.43, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 72.24, 57.56) -- ( 72.24, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 76.06, 57.56) -- ( 76.06, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 79.87, 57.56) -- ( 79.87, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 83.69, 57.56) -- ( 83.69, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 87.50, 57.56) -- ( 87.50, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 91.32, 57.56) -- ( 91.32, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 95.13, 57.56) -- ( 95.13, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 98.95, 57.56) -- ( 98.95, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (102.76, 57.56) -- (102.76, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (106.58, 57.56) -- (106.58, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (110.39, 57.56) -- (110.39, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (114.21, 57.56) -- (114.21, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (118.02, 57.56) -- (118.02, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (121.83, 57.56) -- (121.83, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (125.65, 57.56) -- (125.65, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (129.46, 57.56) -- (129.46, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (133.28, 57.56) -- (133.28, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (137.09, 57.56) -- (137.09, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14, 61.15) -- ( 38.84, 61.15);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14, 66.28) -- ( 38.84, 66.28);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14, 71.41) -- ( 38.84, 71.41);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14, 76.53) -- ( 38.84, 76.53);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14, 81.66) -- ( 38.84, 81.66);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14, 86.79) -- ( 38.84, 86.79);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14, 91.91) -- ( 38.84, 91.91);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14, 97.04) -- ( 38.84, 97.04);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,102.17) -- ( 38.84,102.17);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,107.30) -- ( 38.84,107.30);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,112.42) -- ( 38.84,112.42);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,117.55) -- ( 38.84,117.55);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,122.68) -- ( 38.84,122.68);
\end{scope}
\begin{scope}
\path[clip] (142.55, 57.56) rectangle (247.45,126.48);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (142.55, 57.56) rectangle (247.45,126.48);
\definecolor{drawColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.2pt,line join=round] (142.55, 63.71) --
	(247.45, 63.71);

\path[draw=drawColor,line width= 0.2pt,line join=round] (142.55, 79.10) --
	(247.45, 79.10);

\path[draw=drawColor,line width= 0.2pt,line join=round] (142.55, 94.48) --
	(247.45, 94.48);

\path[draw=drawColor,line width= 0.2pt,line join=round] (142.55,109.86) --
	(247.45,109.86);

\path[draw=drawColor,line width= 0.2pt,line join=round] (142.55,125.24) --
	(247.45,125.24);

\path[draw=drawColor,line width= 0.2pt,line join=round] (149.22, 57.56) --
	(149.22,126.48);

\path[draw=drawColor,line width= 0.2pt,line join=round] (175.92, 57.56) --
	(175.92,126.48);

\path[draw=drawColor,line width= 0.2pt,line join=round] (202.63, 57.56) --
	(202.63,126.48);

\path[draw=drawColor,line width= 0.2pt,line join=round] (229.33, 57.56) --
	(229.33,126.48);

\path[draw=drawColor,line width= 0.5pt,line join=round] (142.55, 71.41) --
	(247.45, 71.41);

\path[draw=drawColor,line width= 0.5pt,line join=round] (142.55, 86.79) --
	(247.45, 86.79);

\path[draw=drawColor,line width= 0.5pt,line join=round] (142.55,102.17) --
	(247.45,102.17);

\path[draw=drawColor,line width= 0.5pt,line join=round] (142.55,117.55) --
	(247.45,117.55);

\path[draw=drawColor,line width= 0.5pt,line join=round] (162.57, 57.56) --
	(162.57,126.48);

\path[draw=drawColor,line width= 0.5pt,line join=round] (189.28, 57.56) --
	(189.28,126.48);

\path[draw=drawColor,line width= 0.5pt,line join=round] (215.98, 57.56) --
	(215.98,126.48);

\path[draw=drawColor,line width= 0.5pt,line join=round] (242.68, 57.56) --
	(242.68,126.48);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (147.31, 72.23) --
	(151.13, 72.23) --
	(154.94, 72.23) --
	(158.76, 72.23) --
	(162.57, 72.23) --
	(166.39, 72.23) --
	(170.20, 72.23) --
	(174.02, 72.27) --
	(177.83, 72.70) --
	(181.65, 76.95) --
	(185.46, 89.66) --
	(189.28,106.37) --
	(193.09,123.35);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (147.31, 72.11) --
	(151.13, 72.11) --
	(154.94, 72.11) --
	(158.76, 72.11) --
	(162.57, 72.11) --
	(166.39, 72.11) --
	(170.20, 72.11) --
	(174.02, 72.15) --
	(177.83, 72.59) --
	(181.65, 76.89) --
	(185.46, 89.67) --
	(189.28,106.37) --
	(193.09,123.35);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] (147.31, 84.74) --
	(151.13, 84.06) --
	(154.94, 83.93) --
	(158.76, 83.97) --
	(162.57, 84.20) --
	(166.39, 84.27) --
	(170.20, 84.38) --
	(174.02, 86.20) --
	(177.83, 93.38) --
	(181.65,111.72);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (147.31, 62.86) --
	(151.13, 60.69) --
	(154.94, 60.69) --
	(158.76, 60.69);
\definecolor{fillColor}{RGB}{117,112,179}

\path[fill=fillColor] (145.35, 82.78) --
	(149.28, 82.78) --
	(149.28, 86.70) --
	(145.35, 86.70) --
	cycle;

\path[fill=fillColor] (149.17, 82.10) --
	(153.09, 82.10) --
	(153.09, 86.02) --
	(149.17, 86.02) --
	cycle;

\path[fill=fillColor] (152.98, 81.97) --
	(156.91, 81.97) --
	(156.91, 85.90) --
	(152.98, 85.90) --
	cycle;

\path[fill=fillColor] (156.80, 82.01) --
	(160.72, 82.01) --
	(160.72, 85.93) --
	(156.80, 85.93) --
	cycle;

\path[fill=fillColor] (160.61, 82.23) --
	(164.54, 82.23) --
	(164.54, 86.16) --
	(160.61, 86.16) --
	cycle;

\path[fill=fillColor] (164.43, 82.30) --
	(168.35, 82.30) --
	(168.35, 86.23) --
	(164.43, 86.23) --
	cycle;

\path[fill=fillColor] (168.24, 82.41) --
	(172.16, 82.41) --
	(172.16, 86.34) --
	(168.24, 86.34) --
	cycle;

\path[fill=fillColor] (172.05, 84.24) --
	(175.98, 84.24) --
	(175.98, 88.16) --
	(172.05, 88.16) --
	cycle;

\path[fill=fillColor] (175.87, 91.42) --
	(179.79, 91.42) --
	(179.79, 95.34) --
	(175.87, 95.34) --
	cycle;

\path[fill=fillColor] (179.68,109.76) --
	(183.61,109.76) --
	(183.61,113.68) --
	(179.68,113.68) --
	cycle;
\definecolor{fillColor}{RGB}{217,95,2}

\path[fill=fillColor] (147.31, 75.16) --
	(149.96, 70.58) --
	(144.67, 70.58) --
	cycle;

\path[fill=fillColor] (151.13, 75.16) --
	(153.77, 70.58) --
	(148.49, 70.58) --
	cycle;

\path[fill=fillColor] (154.94, 75.16) --
	(157.59, 70.58) --
	(152.30, 70.58) --
	cycle;

\path[fill=fillColor] (158.76, 75.16) --
	(161.40, 70.58) --
	(156.12, 70.58) --
	cycle;

\path[fill=fillColor] (162.57, 75.16) --
	(165.22, 70.58) --
	(159.93, 70.58) --
	cycle;

\path[fill=fillColor] (166.39, 75.16) --
	(169.03, 70.58) --
	(163.75, 70.58) --
	cycle;

\path[fill=fillColor] (170.20, 75.16) --
	(172.84, 70.58) --
	(167.56, 70.58) --
	cycle;

\path[fill=fillColor] (174.02, 75.20) --
	(176.66, 70.62) --
	(171.37, 70.62) --
	cycle;

\path[fill=fillColor] (177.83, 75.64) --
	(180.47, 71.06) --
	(175.19, 71.06) --
	cycle;

\path[fill=fillColor] (181.65, 79.94) --
	(184.29, 75.37) --
	(179.00, 75.37) --
	cycle;

\path[fill=fillColor] (185.46, 92.72) --
	(188.10, 88.15) --
	(182.82, 88.15) --
	cycle;

\path[fill=fillColor] (189.28,109.42) --
	(191.92,104.84) --
	(186.63,104.84) --
	cycle;

\path[fill=fillColor] (193.09,126.40) --
	(195.73,121.82) --
	(190.45,121.82) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (144.54, 62.86) -- (150.09, 62.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (147.31, 60.09) -- (147.31, 65.64);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (148.35, 60.69) -- (153.90, 60.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (151.13, 57.92) -- (151.13, 63.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (152.17, 60.69) -- (157.72, 60.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (154.94, 57.92) -- (154.94, 63.47);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (155.98, 60.69) -- (161.53, 60.69);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (158.76, 57.92) -- (158.76, 63.47);
\definecolor{fillColor}{RGB}{27,158,119}

\path[fill=fillColor] (147.31, 72.23) circle (  1.96);

\path[fill=fillColor] (151.13, 72.23) circle (  1.96);

\path[fill=fillColor] (154.94, 72.23) circle (  1.96);

\path[fill=fillColor] (158.76, 72.23) circle (  1.96);

\path[fill=fillColor] (162.57, 72.23) circle (  1.96);

\path[fill=fillColor] (166.39, 72.23) circle (  1.96);

\path[fill=fillColor] (170.20, 72.23) circle (  1.96);

\path[fill=fillColor] (174.02, 72.27) circle (  1.96);

\path[fill=fillColor] (177.83, 72.70) circle (  1.96);

\path[fill=fillColor] (181.65, 76.95) circle (  1.96);

\path[fill=fillColor] (185.46, 89.66) circle (  1.96);

\path[fill=fillColor] (189.28,106.37) circle (  1.96);

\path[fill=fillColor] (193.09,123.35) circle (  1.96);
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (143.50, 57.56) -- (143.50, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (147.31, 57.56) -- (147.31, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (151.13, 57.56) -- (151.13, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (154.94, 57.56) -- (154.94, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (158.76, 57.56) -- (158.76, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (162.57, 57.56) -- (162.57, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (166.39, 57.56) -- (166.39, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (170.20, 57.56) -- (170.20, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (174.02, 57.56) -- (174.02, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (177.83, 57.56) -- (177.83, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (181.65, 57.56) -- (181.65, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (185.46, 57.56) -- (185.46, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (189.28, 57.56) -- (189.28, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (193.09, 57.56) -- (193.09, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (196.90, 57.56) -- (196.90, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (200.72, 57.56) -- (200.72, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (204.53, 57.56) -- (204.53, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (208.35, 57.56) -- (208.35, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (212.16, 57.56) -- (212.16, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (215.98, 57.56) -- (215.98, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (219.79, 57.56) -- (219.79, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (223.61, 57.56) -- (223.61, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (227.42, 57.56) -- (227.42, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (231.24, 57.56) -- (231.24, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (235.05, 57.56) -- (235.05, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (238.87, 57.56) -- (238.87, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (242.68, 57.56) -- (242.68, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (246.49, 57.56) -- (246.49, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (142.55, 61.15) -- (148.24, 61.15);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (142.55, 66.28) -- (148.24, 66.28);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (142.55, 71.41) -- (148.24, 71.41);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (142.55, 76.53) -- (148.24, 76.53);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (142.55, 81.66) -- (148.24, 81.66);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (142.55, 86.79) -- (148.24, 86.79);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (142.55, 91.91) -- (148.24, 91.91);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (142.55, 97.04) -- (148.24, 97.04);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (142.55,102.17) -- (148.24,102.17);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (142.55,107.30) -- (148.24,107.30);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (142.55,112.42) -- (148.24,112.42);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (142.55,117.55) -- (148.24,117.55);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (142.55,122.68) -- (148.24,122.68);
\end{scope}
\begin{scope}
\path[clip] (251.95, 57.56) rectangle (356.85,126.48);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (251.95, 57.56) rectangle (356.85,126.48);
\definecolor{drawColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.2pt,line join=round] (251.95, 63.71) --
	(356.85, 63.71);

\path[draw=drawColor,line width= 0.2pt,line join=round] (251.95, 79.10) --
	(356.85, 79.10);

\path[draw=drawColor,line width= 0.2pt,line join=round] (251.95, 94.48) --
	(356.85, 94.48);

\path[draw=drawColor,line width= 0.2pt,line join=round] (251.95,109.86) --
	(356.85,109.86);

\path[draw=drawColor,line width= 0.2pt,line join=round] (251.95,125.24) --
	(356.85,125.24);

\path[draw=drawColor,line width= 0.2pt,line join=round] (258.62, 57.56) --
	(258.62,126.48);

\path[draw=drawColor,line width= 0.2pt,line join=round] (285.33, 57.56) --
	(285.33,126.48);

\path[draw=drawColor,line width= 0.2pt,line join=round] (312.03, 57.56) --
	(312.03,126.48);

\path[draw=drawColor,line width= 0.2pt,line join=round] (338.73, 57.56) --
	(338.73,126.48);

\path[draw=drawColor,line width= 0.5pt,line join=round] (251.95, 71.41) --
	(356.85, 71.41);

\path[draw=drawColor,line width= 0.5pt,line join=round] (251.95, 86.79) --
	(356.85, 86.79);

\path[draw=drawColor,line width= 0.5pt,line join=round] (251.95,102.17) --
	(356.85,102.17);

\path[draw=drawColor,line width= 0.5pt,line join=round] (251.95,117.55) --
	(356.85,117.55);

\path[draw=drawColor,line width= 0.5pt,line join=round] (271.97, 57.56) --
	(271.97,126.48);

\path[draw=drawColor,line width= 0.5pt,line join=round] (298.68, 57.56) --
	(298.68,126.48);

\path[draw=drawColor,line width= 0.5pt,line join=round] (325.38, 57.56) --
	(325.38,126.48);

\path[draw=drawColor,line width= 0.5pt,line join=round] (352.08, 57.56) --
	(352.08,126.48);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (256.72, 73.16) --
	(260.53, 73.16) --
	(264.35, 73.16) --
	(268.16, 73.16) --
	(271.97, 73.16) --
	(275.79, 73.16) --
	(279.60, 73.16) --
	(283.42, 73.16) --
	(287.23, 73.16) --
	(291.05, 73.16) --
	(294.86, 73.16) --
	(298.68, 73.16) --
	(302.49, 73.17) --
	(306.31, 73.17) --
	(310.12, 73.20) --
	(313.94, 73.25) --
	(317.75, 73.38) --
	(321.56, 73.71) --
	(325.38, 74.42) --
	(329.19, 76.07) --
	(333.01, 78.80) --
	(336.82, 82.57) --
	(340.64, 87.78) --
	(344.45, 94.24) --
	(348.27,100.04) --
	(352.08,106.25);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (256.72, 72.63) --
	(260.53, 72.63) --
	(264.35, 72.63) --
	(268.16, 72.63) --
	(271.97, 72.63) --
	(275.79, 72.63) --
	(279.60, 72.63) --
	(283.42, 72.63) --
	(287.23, 72.63) --
	(291.05, 72.63) --
	(294.86, 72.63) --
	(298.68, 72.63) --
	(302.49, 72.64) --
	(306.31, 72.64) --
	(310.12, 72.67) --
	(313.94, 72.73) --
	(317.75, 72.87) --
	(321.56, 73.21) --
	(325.38, 73.96) --
	(329.19, 75.58) --
	(333.01, 78.49) --
	(336.82, 82.32) --
	(340.64, 87.62) --
	(344.45, 94.08) --
	(348.27, 99.98) --
	(352.08,106.31);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] (256.72, 85.43) --
	(260.53, 83.72) --
	(264.35, 83.89) --
	(268.16, 83.69) --
	(271.97, 85.93) --
	(275.79,108.62);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (256.72, 61.86) --
	(260.53, 61.00) --
	(264.35, 61.00) --
	(268.16, 61.00) --
	(271.97, 61.00) --
	(275.79, 61.15) --
	(279.60, 61.15);
\definecolor{fillColor}{RGB}{27,158,119}

\path[fill=fillColor] (256.72, 73.16) circle (  1.96);

\path[fill=fillColor] (260.53, 73.16) circle (  1.96);

\path[fill=fillColor] (264.35, 73.16) circle (  1.96);

\path[fill=fillColor] (268.16, 73.16) circle (  1.96);

\path[fill=fillColor] (271.97, 73.16) circle (  1.96);

\path[fill=fillColor] (275.79, 73.16) circle (  1.96);

\path[fill=fillColor] (279.60, 73.16) circle (  1.96);

\path[fill=fillColor] (283.42, 73.16) circle (  1.96);

\path[fill=fillColor] (287.23, 73.16) circle (  1.96);

\path[fill=fillColor] (291.05, 73.16) circle (  1.96);

\path[fill=fillColor] (294.86, 73.16) circle (  1.96);

\path[fill=fillColor] (298.68, 73.16) circle (  1.96);

\path[fill=fillColor] (302.49, 73.17) circle (  1.96);

\path[fill=fillColor] (306.31, 73.17) circle (  1.96);

\path[fill=fillColor] (310.12, 73.20) circle (  1.96);

\path[fill=fillColor] (313.94, 73.25) circle (  1.96);

\path[fill=fillColor] (317.75, 73.38) circle (  1.96);

\path[fill=fillColor] (321.56, 73.71) circle (  1.96);

\path[fill=fillColor] (325.38, 74.42) circle (  1.96);

\path[fill=fillColor] (329.19, 76.07) circle (  1.96);

\path[fill=fillColor] (333.01, 78.80) circle (  1.96);

\path[fill=fillColor] (336.82, 82.57) circle (  1.96);

\path[fill=fillColor] (340.64, 87.78) circle (  1.96);

\path[fill=fillColor] (344.45, 94.24) circle (  1.96);

\path[fill=fillColor] (348.27,100.04) circle (  1.96);

\path[fill=fillColor] (352.08,106.25) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (253.94, 61.86) -- (259.49, 61.86);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (256.72, 59.08) -- (256.72, 64.63);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (257.76, 61.00) -- (263.31, 61.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (260.53, 58.23) -- (260.53, 63.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.57, 61.00) -- (267.12, 61.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (264.35, 58.23) -- (264.35, 63.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (265.39, 61.00) -- (270.94, 61.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (268.16, 58.23) -- (268.16, 63.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (269.20, 61.00) -- (274.75, 61.00);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (271.97, 58.23) -- (271.97, 63.78);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (273.01, 61.15) -- (278.56, 61.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (275.79, 58.38) -- (275.79, 63.93);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (276.83, 61.15) -- (282.38, 61.15);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (279.60, 58.38) -- (279.60, 63.93);
\definecolor{fillColor}{RGB}{117,112,179}

\path[fill=fillColor] (254.75, 83.46) --
	(258.68, 83.46) --
	(258.68, 87.39) --
	(254.75, 87.39) --
	cycle;

\path[fill=fillColor] (258.57, 81.76) --
	(262.49, 81.76) --
	(262.49, 85.68) --
	(258.57, 85.68) --
	cycle;

\path[fill=fillColor] (262.38, 81.92) --
	(266.31, 81.92) --
	(266.31, 85.85) --
	(262.38, 85.85) --
	cycle;

\path[fill=fillColor] (266.20, 81.73) --
	(270.12, 81.73) --
	(270.12, 85.66) --
	(266.20, 85.66) --
	cycle;

\path[fill=fillColor] (270.01, 83.97) --
	(273.94, 83.97) --
	(273.94, 87.89) --
	(270.01, 87.89) --
	cycle;

\path[fill=fillColor] (273.83,106.66) --
	(277.75,106.66) --
	(277.75,110.59) --
	(273.83,110.59) --
	cycle;
\definecolor{fillColor}{RGB}{217,95,2}

\path[fill=fillColor] (256.72, 75.68) --
	(259.36, 71.10) --
	(254.07, 71.10) --
	cycle;

\path[fill=fillColor] (260.53, 75.68) --
	(263.17, 71.10) --
	(257.89, 71.10) --
	cycle;

\path[fill=fillColor] (264.35, 75.68) --
	(266.99, 71.10) --
	(261.70, 71.10) --
	cycle;

\path[fill=fillColor] (268.16, 75.68) --
	(270.80, 71.10) --
	(265.52, 71.10) --
	cycle;

\path[fill=fillColor] (271.97, 75.68) --
	(274.62, 71.10) --
	(269.33, 71.10) --
	cycle;

\path[fill=fillColor] (275.79, 75.68) --
	(278.43, 71.10) --
	(273.15, 71.10) --
	cycle;

\path[fill=fillColor] (279.60, 75.68) --
	(282.25, 71.10) --
	(276.96, 71.10) --
	cycle;

\path[fill=fillColor] (283.42, 75.68) --
	(286.06, 71.10) --
	(280.78, 71.10) --
	cycle;

\path[fill=fillColor] (287.23, 75.68) --
	(289.88, 71.10) --
	(284.59, 71.10) --
	cycle;

\path[fill=fillColor] (291.05, 75.68) --
	(293.69, 71.10) --
	(288.41, 71.10) --
	cycle;

\path[fill=fillColor] (294.86, 75.68) --
	(297.51, 71.10) --
	(292.22, 71.10) --
	cycle;

\path[fill=fillColor] (298.68, 75.68) --
	(301.32, 71.10) --
	(296.03, 71.10) --
	cycle;

\path[fill=fillColor] (302.49, 75.69) --
	(305.13, 71.11) --
	(299.85, 71.11) --
	cycle;

\path[fill=fillColor] (306.31, 75.69) --
	(308.95, 71.12) --
	(303.66, 71.12) --
	cycle;

\path[fill=fillColor] (310.12, 75.72) --
	(312.76, 71.14) --
	(307.48, 71.14) --
	cycle;

\path[fill=fillColor] (313.94, 75.78) --
	(316.58, 71.20) --
	(311.29, 71.20) --
	cycle;

\path[fill=fillColor] (317.75, 75.92) --
	(320.39, 71.34) --
	(315.11, 71.34) --
	cycle;

\path[fill=fillColor] (321.56, 76.27) --
	(324.21, 71.69) --
	(318.92, 71.69) --
	cycle;

\path[fill=fillColor] (325.38, 77.01) --
	(328.02, 72.44) --
	(322.74, 72.44) --
	cycle;

\path[fill=fillColor] (329.19, 78.63) --
	(331.84, 74.06) --
	(326.55, 74.06) --
	cycle;

\path[fill=fillColor] (333.01, 81.55) --
	(335.65, 76.97) --
	(330.37, 76.97) --
	cycle;

\path[fill=fillColor] (336.82, 85.37) --
	(339.47, 80.80) --
	(334.18, 80.80) --
	cycle;

\path[fill=fillColor] (340.64, 90.67) --
	(343.28, 86.09) --
	(338.00, 86.09) --
	cycle;

\path[fill=fillColor] (344.45, 97.14) --
	(347.10, 92.56) --
	(341.81, 92.56) --
	cycle;

\path[fill=fillColor] (348.27,103.03) --
	(350.91, 98.45) --
	(345.62, 98.45) --
	cycle;

\path[fill=fillColor] (352.08,109.36) --
	(354.72,104.79) --
	(349.44,104.79) --
	cycle;
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (252.90, 57.56) -- (252.90, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (256.72, 57.56) -- (256.72, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (260.53, 57.56) -- (260.53, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (264.35, 57.56) -- (264.35, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (268.16, 57.56) -- (268.16, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (271.97, 57.56) -- (271.97, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (275.79, 57.56) -- (275.79, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (279.60, 57.56) -- (279.60, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (283.42, 57.56) -- (283.42, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (287.23, 57.56) -- (287.23, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (291.05, 57.56) -- (291.05, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (294.86, 57.56) -- (294.86, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (298.68, 57.56) -- (298.68, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (302.49, 57.56) -- (302.49, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (306.31, 57.56) -- (306.31, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (310.12, 57.56) -- (310.12, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (313.94, 57.56) -- (313.94, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (317.75, 57.56) -- (317.75, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (321.56, 57.56) -- (321.56, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (325.38, 57.56) -- (325.38, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (329.19, 57.56) -- (329.19, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (333.01, 57.56) -- (333.01, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (336.82, 57.56) -- (336.82, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (340.64, 57.56) -- (340.64, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (344.45, 57.56) -- (344.45, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (348.27, 57.56) -- (348.27, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (352.08, 57.56) -- (352.08, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (355.90, 57.56) -- (355.90, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.95, 61.15) -- (257.64, 61.15);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.95, 66.28) -- (257.64, 66.28);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.95, 71.41) -- (257.64, 71.41);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.95, 76.53) -- (257.64, 76.53);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.95, 81.66) -- (257.64, 81.66);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.95, 86.79) -- (257.64, 86.79);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.95, 91.91) -- (257.64, 91.91);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.95, 97.04) -- (257.64, 97.04);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.95,102.17) -- (257.64,102.17);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.95,107.30) -- (257.64,107.30);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.95,112.42) -- (257.64,112.42);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.95,117.55) -- (257.64,117.55);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (251.95,122.68) -- (257.64,122.68);
\end{scope}
\begin{scope}
\path[clip] ( 33.14,126.48) rectangle (138.05,140.04);
\definecolor{fillColor}{gray}{0.85}

\path[fill=fillColor] ( 33.14,126.48) rectangle (138.05,140.04);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 85.60,130.78) {Bijections};
\end{scope}
\begin{scope}
\path[clip] (142.55,126.48) rectangle (247.45,140.04);
\definecolor{fillColor}{gray}{0.85}

\path[fill=fillColor] (142.55,126.48) rectangle (247.45,140.04);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (195.00,130.78) {Friends \& Smokers};
\end{scope}
\begin{scope}
\path[clip] (251.95,126.48) rectangle (356.85,140.04);
\definecolor{fillColor}{gray}{0.85}

\path[fill=fillColor] (251.95,126.48) rectangle (356.85,140.04);
\definecolor{drawColor}{gray}{0.10}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (304.40,130.78) {Functions};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 53.17, 55.31) --
	( 53.17, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 79.87, 55.31) --
	( 79.87, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (106.58, 55.31) --
	(106.58, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (133.28, 55.31) --
	(133.28, 57.56);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 53.17, 48.55) {$2^{5}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 79.87, 48.55) {$2^{12}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (106.58, 48.55) {$2^{19}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (133.28, 48.55) {$2^{26}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] (162.57, 55.31) --
	(162.57, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (189.28, 55.31) --
	(189.28, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (215.98, 55.31) --
	(215.98, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (242.68, 55.31) --
	(242.68, 57.56);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (162.57, 48.55) {$2^{5}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (189.28, 48.55) {$2^{12}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (215.98, 48.55) {$2^{19}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (242.68, 48.55) {$2^{26}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] (271.97, 55.31) --
	(271.97, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (298.68, 55.31) --
	(298.68, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (325.38, 55.31) --
	(325.38, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (352.08, 55.31) --
	(352.08, 57.56);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (271.97, 48.55) {$2^{5}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (298.68, 48.55) {$2^{12}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (325.38, 48.55) {$2^{19}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (352.08, 48.55) {$2^{26}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 29.09, 68.93) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 29.09, 84.31) {16};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 29.09, 99.69) {128};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 29.09,115.07) {1024};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 30.89, 71.41) --
	( 33.14, 71.41);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 30.89, 86.79) --
	( 33.14, 86.79);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 30.89,102.17) --
	( 33.14,102.17);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 30.89,117.55) --
	( 33.14,117.55);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (195.00, 38.70) {Domain size};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 10.70, 92.02) {Total runtime (s)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 52.31,  4.50) rectangle (337.68, 27.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] ( 61.31,  9.00) rectangle ( 75.77, 23.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 62.76, 16.23) -- ( 74.32, 16.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{fillColor}{RGB}{27,158,119}

\path[fill=fillColor] ( 68.54, 16.23) circle (  1.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (129.38,  9.00) rectangle (143.83, 23.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (130.83, 16.23) -- (142.39, 16.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{fillColor}{RGB}{217,95,2}

\path[fill=fillColor] (136.61, 19.28) --
	(139.25, 14.70) --
	(133.97, 14.70) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (209.71,  9.00) rectangle (224.16, 23.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] (211.15, 16.23) -- (222.72, 16.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{fillColor}{RGB}{117,112,179}

\path[fill=fillColor] (214.97, 14.26) --
	(218.90, 14.26) --
	(218.90, 18.19) --
	(214.97, 18.19) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (281.15,  9.00) rectangle (295.60, 23.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (282.59, 16.23) -- (294.16, 16.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (285.60, 16.23) -- (291.15, 16.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (288.38, 13.45) -- (288.38, 19.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 80.27, 13.75) {\textsc{Crane2-BFS}};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at (148.33, 13.75) {\textsc{Crane2-Greedy}};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at (228.66, 13.75) {\textsc{FastWFOMC}};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (361.35,144.54);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at (300.10, 13.75) {\textsc{ForcLift}};
\end{scope}
\end{tikzpicture}

\end{document}