### 2025-06-13 doc/talk/bijections.tex
Add experimental results

% Created by tikzDevice version 0.12.6 on 2025-06-13 10:38:09
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (307.15,216.81);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.5pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (307.15,216.81);
\end{scope}
\begin{scope}
\path[clip] ( 29.55, 57.56) rectangle (302.65,212.31);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] ( 29.55, 57.56) rectangle (302.65,212.31);
\definecolor{drawColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 29.55, 66.00) --
	(302.65, 66.00);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 29.55,118.22) --
	(302.65,118.22);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 29.55,170.44) --
	(302.65,170.44);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 53.24, 57.56) --
	( 53.24,212.31);

\path[draw=drawColor,line width= 0.2pt,line join=round] (120.96, 57.56) --
	(120.96,212.31);

\path[draw=drawColor,line width= 0.2pt,line join=round] (188.67, 57.56) --
	(188.67,212.31);

\path[draw=drawColor,line width= 0.2pt,line join=round] (256.38, 57.56) --
	(256.38,212.31);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 29.55, 92.11) --
	(302.65, 92.11);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 29.55,144.33) --
	(302.65,144.33);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 29.55,196.55) --
	(302.65,196.55);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 87.10, 57.56) --
	( 87.10,212.31);

\path[draw=drawColor,line width= 0.5pt,line join=round] (154.81, 57.56) --
	(154.81,212.31);

\path[draw=drawColor,line width= 0.5pt,line join=round] (222.52, 57.56) --
	(222.52,212.31);

\path[draw=drawColor,line width= 0.5pt,line join=round] (290.23, 57.56) --
	(290.23,212.31);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.96, 64.59) --
	( 64.53, 64.59) --
	( 87.10, 64.59) --
	(109.67, 64.59) --
	(132.24, 64.59) --
	(154.81, 64.60) --
	(177.38, 64.62) --
	(199.95, 64.73) --
	(222.52, 65.13) --
	(245.09, 66.75) --
	(267.66, 73.60) --
	(290.23, 93.28);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 7pt off 3pt ,line join=round] ( 41.96,107.13) --
	( 64.53, 79.15) --
	( 87.10, 79.15) --
	(109.67, 79.64) --
	(132.24, 96.07) --
	(154.81,205.28);
\definecolor{fillColor}{RGB}{27,158,119}

\path[fill=fillColor] ( 41.96, 64.59) circle (  1.96);

\path[fill=fillColor] ( 64.53, 64.59) circle (  1.96);

\path[fill=fillColor] ( 87.10, 64.59) circle (  1.96);

\path[fill=fillColor] (109.67, 64.59) circle (  1.96);

\path[fill=fillColor] (132.24, 64.59) circle (  1.96);

\path[fill=fillColor] (154.81, 64.60) circle (  1.96);

\path[fill=fillColor] (177.38, 64.62) circle (  1.96);

\path[fill=fillColor] (199.95, 64.73) circle (  1.96);

\path[fill=fillColor] (222.52, 65.13) circle (  1.96);

\path[fill=fillColor] (245.09, 66.75) circle (  1.96);

\path[fill=fillColor] (267.66, 73.60) circle (  1.96);

\path[fill=fillColor] (290.23, 93.28) circle (  1.96);
\definecolor{fillColor}{RGB}{117,112,179}

\path[fill=fillColor] ( 40.00,105.17) --
	( 43.92,105.17) --
	( 43.92,109.10) --
	( 40.00,109.10) --
	cycle;

\path[fill=fillColor] ( 62.57, 77.18) --
	( 66.49, 77.18) --
	( 66.49, 81.11) --
	( 62.57, 81.11) --
	cycle;

\path[fill=fillColor] ( 85.14, 77.18) --
	( 89.06, 77.18) --
	( 89.06, 81.11) --
	( 85.14, 81.11) --
	cycle;

\path[fill=fillColor] (107.71, 77.68) --
	(111.63, 77.68) --
	(111.63, 81.60) --
	(107.71, 81.60) --
	cycle;

\path[fill=fillColor] (130.28, 94.10) --
	(134.20, 94.10) --
	(134.20, 98.03) --
	(130.28, 98.03) --
	cycle;

\path[fill=fillColor] (152.85,203.31) --
	(156.77,203.31) --
	(156.77,207.24) --
	(152.85,207.24) --
	cycle;
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 41.96, 57.56) -- ( 41.96, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 64.53, 57.56) -- ( 64.53, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 87.10, 57.56) -- ( 87.10, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (109.67, 57.56) -- (109.67, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (132.24, 57.56) -- (132.24, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (154.81, 57.56) -- (154.81, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (177.38, 57.56) -- (177.38, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (199.95, 57.56) -- (199.95, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (222.52, 57.56) -- (222.52, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (245.09, 57.56) -- (245.09, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (267.66, 57.56) -- (267.66, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (290.23, 57.56) -- (290.23, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 29.55, 66.00) -- ( 35.24, 66.00);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 29.55, 92.11) -- ( 35.24, 92.11);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 29.55,118.22) -- ( 35.24,118.22);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 29.55,144.33) -- ( 35.24,144.33);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 29.55,170.44) -- ( 35.24,170.44);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 29.55,196.55) -- ( 35.24,196.55);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 25.50, 89.63) {16};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 25.50,141.85) {64};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 25.50,194.07) {256};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 27.30, 92.11) --
	( 29.55, 92.11);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 27.30,144.33) --
	( 29.55,144.33);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 27.30,196.55) --
	( 29.55,196.55);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 87.10, 55.31) --
	( 87.10, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (154.81, 55.31) --
	(154.81, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (222.52, 55.31) --
	(222.52, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (290.23, 55.31) --
	(290.23, 57.56);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 87.10, 48.55) {$2^{3}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (154.81, 48.55) {$2^{6}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (222.52, 48.55) {$2^{9}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (290.23, 48.55) {$2^{12}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (166.10, 38.70) {Domain size};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 10.70,134.94) {Total runtime (s)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 91.84,  4.50) rectangle (240.35, 27.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (100.84,  9.00) rectangle (115.30, 23.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] (102.29, 16.23) -- (113.85, 16.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{fillColor}{RGB}{27,158,119}

\path[fill=fillColor] (108.07, 16.23) circle (  1.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (168.91,  9.00) rectangle (183.36, 23.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 7pt off 3pt ,line join=round] (170.36, 16.23) -- (181.92, 16.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{fillColor}{RGB}{117,112,179}

\path[fill=fillColor] (174.17, 14.26) --
	(178.10, 14.26) --
	(178.10, 18.19) --
	(174.17, 18.19) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at (119.80, 13.75) {\textsc{Crane2-BFS}};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at (187.86, 13.75) {\textsc{FastWFOMC}};
\end{scope}
\end{tikzpicture}

\end{document}