### 2025-06-13 doc/talk/friends.tex

% Created by tikzDevice version 0.12.6 on 2025-06-13 10:42:02
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (307.15,216.81);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.5pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (307.15,216.81);
\end{scope}
\begin{scope}
\path[clip] ( 33.14, 57.56) rectangle (302.65,212.31);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] ( 33.14, 57.56) rectangle (302.65,212.31);
\definecolor{drawColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 33.14, 71.38) --
	(302.65, 71.38);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 33.14,105.92) --
	(302.65,105.92);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 33.14,140.45) --
	(302.65,140.45);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 33.14,174.99) --
	(302.65,174.99);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 33.14,209.53) --
	(302.65,209.53);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 55.60, 57.56) --
	( 55.60,212.31);

\path[draw=drawColor,line width= 0.2pt,line join=round] (116.85, 57.56) --
	(116.85,212.31);

\path[draw=drawColor,line width= 0.2pt,line join=round] (178.10, 57.56) --
	(178.10,212.31);

\path[draw=drawColor,line width= 0.2pt,line join=round] (239.36, 57.56) --
	(239.36,212.31);

\path[draw=drawColor,line width= 0.2pt,line join=round] (300.61, 57.56) --
	(300.61,212.31);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 33.14, 88.65) --
	(302.65, 88.65);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 33.14,123.18) --
	(302.65,123.18);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 33.14,157.72) --
	(302.65,157.72);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 33.14,192.26) --
	(302.65,192.26);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 86.23, 57.56) --
	( 86.23,212.31);

\path[draw=drawColor,line width= 0.5pt,line join=round] (147.48, 57.56) --
	(147.48,212.31);

\path[draw=drawColor,line width= 0.5pt,line join=round] (208.73, 57.56) --
	(208.73,212.31);

\path[draw=drawColor,line width= 0.5pt,line join=round] (269.98, 57.56) --
	(269.98,212.31);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 45.39, 90.51) --
	( 65.81, 90.50) --
	( 86.23, 90.50) --
	(106.65, 90.50) --
	(127.06, 90.50) --
	(147.48, 90.50) --
	(167.90, 90.51) --
	(188.31, 90.59) --
	(208.73, 91.56) --
	(229.15,101.10) --
	(249.56,129.63) --
	(269.98,167.15) --
	(290.40,205.28);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 45.39, 90.23) --
	( 65.81, 90.22) --
	( 86.23, 90.22) --
	(106.65, 90.22) --
	(127.06, 90.22) --
	(147.48, 90.22) --
	(167.90, 90.23) --
	(188.31, 90.31) --
	(208.73, 91.30) --
	(229.15,100.97) --
	(249.56,129.66) --
	(269.98,167.14) --
	(290.40,205.28);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 45.39,118.59) --
	( 65.81,117.07) --
	( 86.23,116.78) --
	(106.65,116.86) --
	(127.06,117.36) --
	(147.48,117.53) --
	(167.90,117.77) --
	(188.31,121.87) --
	(208.73,137.98) --
	(229.15,179.16);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] ( 45.39, 69.46) --
	( 65.81, 64.59) --
	( 86.23, 64.59) --
	(106.65, 64.59);
\definecolor{fillColor}{RGB}{117,112,179}

\path[fill=fillColor] ( 43.43,116.62) --
	( 47.36,116.62) --
	( 47.36,120.55) --
	( 43.43,120.55) --
	cycle;

\path[fill=fillColor] ( 63.85,115.10) --
	( 67.77,115.10) --
	( 67.77,119.03) --
	( 63.85,119.03) --
	cycle;

\path[fill=fillColor] ( 84.27,114.82) --
	( 88.19,114.82) --
	( 88.19,118.74) --
	( 84.27,118.74) --
	cycle;

\path[fill=fillColor] (104.68,114.89) --
	(108.61,114.89) --
	(108.61,118.82) --
	(104.68,118.82) --
	cycle;

\path[fill=fillColor] (125.10,115.40) --
	(129.02,115.40) --
	(129.02,119.33) --
	(125.10,119.33) --
	cycle;

\path[fill=fillColor] (145.52,115.56) --
	(149.44,115.56) --
	(149.44,119.49) --
	(145.52,119.49) --
	cycle;

\path[fill=fillColor] (165.93,115.81) --
	(169.86,115.81) --
	(169.86,119.73) --
	(165.93,119.73) --
	cycle;

\path[fill=fillColor] (186.35,119.91) --
	(190.28,119.91) --
	(190.28,123.83) --
	(186.35,123.83) --
	cycle;

\path[fill=fillColor] (206.77,136.02) --
	(210.69,136.02) --
	(210.69,139.94) --
	(206.77,139.94) --
	cycle;

\path[fill=fillColor] (227.18,177.20) --
	(231.11,177.20) --
	(231.11,181.13) --
	(227.18,181.13) --
	cycle;
\definecolor{fillColor}{RGB}{217,95,2}

\path[fill=fillColor] ( 45.39, 93.28) --
	( 48.04, 88.70) --
	( 42.75, 88.70) --
	cycle;

\path[fill=fillColor] ( 65.81, 93.27) --
	( 68.45, 88.70) --
	( 63.17, 88.70) --
	cycle;

\path[fill=fillColor] ( 86.23, 93.27) --
	( 88.87, 88.70) --
	( 83.59, 88.70) --
	cycle;

\path[fill=fillColor] (106.65, 93.27) --
	(109.29, 88.70) --
	(104.00, 88.70) --
	cycle;

\path[fill=fillColor] (127.06, 93.27) --
	(129.70, 88.70) --
	(124.42, 88.70) --
	cycle;

\path[fill=fillColor] (147.48, 93.27) --
	(150.12, 88.70) --
	(144.84, 88.70) --
	cycle;

\path[fill=fillColor] (167.90, 93.28) --
	(170.54, 88.70) --
	(165.25, 88.70) --
	cycle;

\path[fill=fillColor] (188.31, 93.36) --
	(190.96, 88.79) --
	(185.67, 88.79) --
	cycle;

\path[fill=fillColor] (208.73, 94.35) --
	(211.37, 89.77) --
	(206.09, 89.77) --
	cycle;

\path[fill=fillColor] (229.15,104.02) --
	(231.79, 99.44) --
	(226.50, 99.44) --
	cycle;

\path[fill=fillColor] (249.56,132.72) --
	(252.21,128.14) --
	(246.92,128.14) --
	cycle;

\path[fill=fillColor] (269.98,170.20) --
	(272.62,165.62) --
	(267.34,165.62) --
	cycle;

\path[fill=fillColor] (290.40,208.33) --
	(293.04,203.75) --
	(287.75,203.75) --
	cycle;

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 42.62, 69.46) -- ( 48.17, 69.46);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 45.39, 66.69) -- ( 45.39, 72.24);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 63.04, 64.59) -- ( 68.59, 64.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 65.81, 61.82) -- ( 65.81, 67.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 83.45, 64.59) -- ( 89.00, 64.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 86.23, 61.82) -- ( 86.23, 67.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (103.87, 64.59) -- (109.42, 64.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (106.65, 61.82) -- (106.65, 67.37);
\definecolor{fillColor}{RGB}{27,158,119}

\path[fill=fillColor] ( 45.39, 90.51) circle (  1.96);

\path[fill=fillColor] ( 65.81, 90.50) circle (  1.96);

\path[fill=fillColor] ( 86.23, 90.50) circle (  1.96);

\path[fill=fillColor] (106.65, 90.50) circle (  1.96);

\path[fill=fillColor] (127.06, 90.50) circle (  1.96);

\path[fill=fillColor] (147.48, 90.50) circle (  1.96);

\path[fill=fillColor] (167.90, 90.51) circle (  1.96);

\path[fill=fillColor] (188.31, 90.59) circle (  1.96);

\path[fill=fillColor] (208.73, 91.56) circle (  1.96);

\path[fill=fillColor] (229.15,101.10) circle (  1.96);

\path[fill=fillColor] (249.56,129.63) circle (  1.96);

\path[fill=fillColor] (269.98,167.15) circle (  1.96);

\path[fill=fillColor] (290.40,205.28) circle (  1.96);
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 45.39, 57.56) -- ( 45.39, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 65.81, 57.56) -- ( 65.81, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 86.23, 57.56) -- ( 86.23, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (106.65, 57.56) -- (106.65, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (127.06, 57.56) -- (127.06, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (147.48, 57.56) -- (147.48, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (167.90, 57.56) -- (167.90, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (188.31, 57.56) -- (188.31, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (208.73, 57.56) -- (208.73, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (229.15, 57.56) -- (229.15, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (249.56, 57.56) -- (249.56, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (269.98, 57.56) -- (269.98, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (290.40, 57.56) -- (290.40, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14, 65.62) -- ( 38.84, 65.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14, 77.13) -- ( 38.84, 77.13);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14, 88.65) -- ( 38.84, 88.65);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,100.16) -- ( 38.84,100.16);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,111.67) -- ( 38.84,111.67);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,123.18) -- ( 38.84,123.18);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,134.70) -- ( 38.84,134.70);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,146.21) -- ( 38.84,146.21);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,157.72) -- ( 38.84,157.72);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,169.23) -- ( 38.84,169.23);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,180.75) -- ( 38.84,180.75);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,192.26) -- ( 38.84,192.26);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 33.14,203.77) -- ( 38.84,203.77);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 29.09, 86.17) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 29.09,120.71) {16};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 29.09,155.24) {128};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 29.09,189.78) {1024};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 30.89, 88.65) --
	( 33.14, 88.65);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 30.89,123.18) --
	( 33.14,123.18);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 30.89,157.72) --
	( 33.14,157.72);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 30.89,192.26) --
	( 33.14,192.26);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 86.23, 55.31) --
	( 86.23, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (147.48, 55.31) --
	(147.48, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (208.73, 55.31) --
	(208.73, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (269.98, 55.31) --
	(269.98, 57.56);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 86.23, 48.55) {$2^{3}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (147.48, 48.55) {$2^{6}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (208.73, 48.55) {$2^{9}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (269.98, 48.55) {$2^{12}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (167.90, 38.70) {Domain size};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 10.70,134.94) {Total runtime (s)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 25.21,  4.50) rectangle (310.58, 27.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] ( 34.21,  9.00) rectangle ( 48.67, 23.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 35.66, 16.23) -- ( 47.22, 16.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{fillColor}{RGB}{27,158,119}

\path[fill=fillColor] ( 41.44, 16.23) circle (  1.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (102.28,  9.00) rectangle (116.73, 23.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (103.72, 16.23) -- (115.29, 16.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{fillColor}{RGB}{217,95,2}

\path[fill=fillColor] (109.51, 19.28) --
	(112.15, 14.70) --
	(106.86, 14.70) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (182.61,  9.00) rectangle (197.06, 23.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] (184.05, 16.23) -- (195.62, 16.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{fillColor}{RGB}{117,112,179}

\path[fill=fillColor] (187.87, 14.26) --
	(191.80, 14.26) --
	(191.80, 18.19) --
	(187.87, 18.19) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (254.05,  9.00) rectangle (268.50, 23.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (255.49, 16.23) -- (267.06, 16.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (258.50, 16.23) -- (264.05, 16.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (261.27, 13.45) -- (261.27, 19.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 53.17, 13.75) {\textsc{Crane2-BFS}};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at (121.23, 13.75) {\textsc{Crane2-Greedy}};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at (201.56, 13.75) {\textsc{FastWFOMC}};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at (273.00, 13.75) {\textsc{ForcLift}};
\end{scope}
\end{tikzpicture}

\end{document}