### 2025-06-13 doc/talk/functions.tex

% Created by tikzDevice version 0.12.6 on 2025-06-13 10:38:23
% !TEX encoding = UTF-8 Unicode
\documentclass[10pt]{article}
\usepackage{tikz}

\usepackage[active,tightpage,psfixbb]{preview}

\PreviewEnvironment{pgfpicture}

\setlength\PreviewBorder{0pt}
\begin{document}

\begin{tikzpicture}[x=1pt,y=1pt]
\definecolor{fillColor}{RGB}{255,255,255}
\path[use as bounding box,fill=fillColor,fill opacity=0.00] (0,0) rectangle (307.15,216.81);
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{RGB}{255,255,255}
\definecolor{fillColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.5pt,line join=round,line cap=round,fill=fillColor] (  0.00,  0.00) rectangle (307.15,216.81);
\end{scope}
\begin{scope}
\path[clip] ( 29.55, 57.56) rectangle (302.65,212.31);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] ( 29.55, 57.56) rectangle (302.65,212.31);
\definecolor{drawColor}{RGB}{255,255,255}

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 29.55, 72.61) --
	(302.65, 72.61);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 29.55,118.05) --
	(302.65,118.05);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 29.55,163.49) --
	(302.65,163.49);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 29.55,208.93) --
	(302.65,208.93);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 32.03, 57.56) --
	( 32.03,212.31);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 46.92, 57.56) --
	( 46.92,212.31);

\path[draw=drawColor,line width= 0.2pt,line join=round] ( 76.72, 57.56) --
	( 76.72,212.31);

\path[draw=drawColor,line width= 0.2pt,line join=round] (106.51, 57.56) --
	(106.51,212.31);

\path[draw=drawColor,line width= 0.2pt,line join=round] (136.30, 57.56) --
	(136.30,212.31);

\path[draw=drawColor,line width= 0.2pt,line join=round] (166.10, 57.56) --
	(166.10,212.31);

\path[draw=drawColor,line width= 0.2pt,line join=round] (195.89, 57.56) --
	(195.89,212.31);

\path[draw=drawColor,line width= 0.2pt,line join=round] (225.68, 57.56) --
	(225.68,212.31);

\path[draw=drawColor,line width= 0.2pt,line join=round] (255.48, 57.56) --
	(255.48,212.31);

\path[draw=drawColor,line width= 0.2pt,line join=round] (285.27, 57.56) --
	(285.27,212.31);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 29.55, 95.33) --
	(302.65, 95.33);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 29.55,140.77) --
	(302.65,140.77);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 29.55,186.21) --
	(302.65,186.21);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 61.82, 57.56) --
	( 61.82,212.31);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 91.61, 57.56) --
	( 91.61,212.31);

\path[draw=drawColor,line width= 0.5pt,line join=round] (121.41, 57.56) --
	(121.41,212.31);

\path[draw=drawColor,line width= 0.5pt,line join=round] (151.20, 57.56) --
	(151.20,212.31);

\path[draw=drawColor,line width= 0.5pt,line join=round] (180.99, 57.56) --
	(180.99,212.31);

\path[draw=drawColor,line width= 0.5pt,line join=round] (210.79, 57.56) --
	(210.79,212.31);

\path[draw=drawColor,line width= 0.5pt,line join=round] (240.58, 57.56) --
	(240.58,212.31);

\path[draw=drawColor,line width= 0.5pt,line join=round] (270.37, 57.56) --
	(270.37,212.31);

\path[draw=drawColor,line width= 0.5pt,line join=round] (300.16, 57.56) --
	(300.16,212.31);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 41.96,100.51) --
	( 51.89,100.51) --
	( 61.82,100.51) --
	( 71.75,100.51) --
	( 81.68,100.51) --
	( 91.61,100.52) --
	(101.55,100.51) --
	(111.48,100.51) --
	(121.41,100.51) --
	(131.34,100.52) --
	(141.27,100.52) --
	(151.20,100.52) --
	(161.13,100.54) --
	(171.06,100.55) --
	(180.99,100.64) --
	(190.92,100.79) --
	(200.85,101.17) --
	(210.79,102.14) --
	(220.72,104.25) --
	(230.65,109.09) --
	(240.58,117.17) --
	(250.51,128.32) --
	(260.44,143.70) --
	(270.37,162.78) --
	(280.30,179.90) --
	(290.23,198.27);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] ( 41.96, 98.94) --
	( 51.89, 98.94) --
	( 61.82, 98.94) --
	( 71.75, 98.94) --
	( 81.68, 98.94) --
	( 91.61, 98.94) --
	(101.55, 98.95) --
	(111.48, 98.94) --
	(121.41, 98.94) --
	(131.34, 98.95) --
	(141.27, 98.95) --
	(151.20, 98.95) --
	(161.13, 98.96) --
	(171.06, 98.98) --
	(180.99, 99.07) --
	(190.92, 99.24) --
	(200.85, 99.65) --
	(210.79,100.67) --
	(220.72,102.88) --
	(230.65,107.66) --
	(240.58,116.27) --
	(250.51,127.58) --
	(260.44,143.23) --
	(270.37,162.33) --
	(280.30,179.73) --
	(290.23,198.45);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] ( 41.96,136.75) --
	( 51.89,131.71) --
	( 61.82,132.20) --
	( 71.75,131.63) --
	( 81.68,138.24) --
	( 91.61,205.28);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] ( 41.96, 67.12) --
	( 51.89, 64.59) --
	( 61.82, 64.59) --
	( 71.75, 64.59) --
	( 81.68, 64.59) --
	( 91.61, 65.04) --
	(101.55, 65.04);
\definecolor{fillColor}{RGB}{27,158,119}

\path[fill=fillColor] ( 41.96,100.51) circle (  1.96);

\path[fill=fillColor] ( 51.89,100.51) circle (  1.96);

\path[fill=fillColor] ( 61.82,100.51) circle (  1.96);

\path[fill=fillColor] ( 71.75,100.51) circle (  1.96);

\path[fill=fillColor] ( 81.68,100.51) circle (  1.96);

\path[fill=fillColor] ( 91.61,100.52) circle (  1.96);

\path[fill=fillColor] (101.55,100.51) circle (  1.96);

\path[fill=fillColor] (111.48,100.51) circle (  1.96);

\path[fill=fillColor] (121.41,100.51) circle (  1.96);

\path[fill=fillColor] (131.34,100.52) circle (  1.96);

\path[fill=fillColor] (141.27,100.52) circle (  1.96);

\path[fill=fillColor] (151.20,100.52) circle (  1.96);

\path[fill=fillColor] (161.13,100.54) circle (  1.96);

\path[fill=fillColor] (171.06,100.55) circle (  1.96);

\path[fill=fillColor] (180.99,100.64) circle (  1.96);

\path[fill=fillColor] (190.92,100.79) circle (  1.96);

\path[fill=fillColor] (200.85,101.17) circle (  1.96);

\path[fill=fillColor] (210.79,102.14) circle (  1.96);

\path[fill=fillColor] (220.72,104.25) circle (  1.96);

\path[fill=fillColor] (230.65,109.09) circle (  1.96);

\path[fill=fillColor] (240.58,117.17) circle (  1.96);

\path[fill=fillColor] (250.51,128.32) circle (  1.96);

\path[fill=fillColor] (260.44,143.70) circle (  1.96);

\path[fill=fillColor] (270.37,162.78) circle (  1.96);

\path[fill=fillColor] (280.30,179.90) circle (  1.96);

\path[fill=fillColor] (290.23,198.27) circle (  1.96);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 39.18, 67.12) -- ( 44.73, 67.12);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 41.96, 64.34) -- ( 41.96, 69.89);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 49.12, 64.59) -- ( 54.66, 64.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 51.89, 61.82) -- ( 51.89, 67.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 59.05, 64.59) -- ( 64.60, 64.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 61.82, 61.82) -- ( 61.82, 67.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 68.98, 64.59) -- ( 74.53, 64.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 71.75, 61.82) -- ( 71.75, 67.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 78.91, 64.59) -- ( 84.46, 64.59);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 81.68, 61.82) -- ( 81.68, 67.37);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 88.84, 65.04) -- ( 94.39, 65.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 91.61, 62.26) -- ( 91.61, 67.81);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] ( 98.77, 65.04) -- (104.32, 65.04);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (101.55, 62.26) -- (101.55, 67.81);
\definecolor{fillColor}{RGB}{117,112,179}

\path[fill=fillColor] ( 40.00,134.78) --
	( 43.92,134.78) --
	( 43.92,138.71) --
	( 40.00,138.71) --
	cycle;

\path[fill=fillColor] ( 49.93,129.75) --
	( 53.85,129.75) --
	( 53.85,133.67) --
	( 49.93,133.67) --
	cycle;

\path[fill=fillColor] ( 59.86,130.24) --
	( 63.78,130.24) --
	( 63.78,134.16) --
	( 59.86,134.16) --
	cycle;

\path[fill=fillColor] ( 69.79,129.67) --
	( 73.71,129.67) --
	( 73.71,133.59) --
	( 69.79,133.59) --
	cycle;

\path[fill=fillColor] ( 79.72,136.28) --
	( 83.65,136.28) --
	( 83.65,140.20) --
	( 79.72,140.20) --
	cycle;

\path[fill=fillColor] ( 89.65,203.31) --
	( 93.58,203.31) --
	( 93.58,207.24) --
	( 89.65,207.24) --
	cycle;
\definecolor{fillColor}{RGB}{217,95,2}

\path[fill=fillColor] ( 41.96,101.99) --
	( 44.60, 97.41) --
	( 39.32, 97.41) --
	cycle;

\path[fill=fillColor] ( 51.89,101.99) --
	( 54.53, 97.41) --
	( 49.25, 97.41) --
	cycle;

\path[fill=fillColor] ( 61.82,101.99) --
	( 64.46, 97.41) --
	( 59.18, 97.41) --
	cycle;

\path[fill=fillColor] ( 71.75,101.99) --
	( 74.39, 97.41) --
	( 69.11, 97.41) --
	cycle;

\path[fill=fillColor] ( 81.68,101.99) --
	( 84.33, 97.41) --
	( 79.04, 97.41) --
	cycle;

\path[fill=fillColor] ( 91.61,101.99) --
	( 94.26, 97.41) --
	( 88.97, 97.41) --
	cycle;

\path[fill=fillColor] (101.55,102.00) --
	(104.19, 97.42) --
	( 98.90, 97.42) --
	cycle;

\path[fill=fillColor] (111.48,101.99) --
	(114.12, 97.41) --
	(108.83, 97.41) --
	cycle;

\path[fill=fillColor] (121.41,101.99) --
	(124.05, 97.41) --
	(118.76, 97.41) --
	cycle;

\path[fill=fillColor] (131.34,102.00) --
	(133.98, 97.42) --
	(128.70, 97.42) --
	cycle;

\path[fill=fillColor] (141.27,102.00) --
	(143.91, 97.42) --
	(138.63, 97.42) --
	cycle;

\path[fill=fillColor] (151.20,102.00) --
	(153.84, 97.42) --
	(148.56, 97.42) --
	cycle;

\path[fill=fillColor] (161.13,102.02) --
	(163.77, 97.44) --
	(158.49, 97.44) --
	cycle;

\path[fill=fillColor] (171.06,102.03) --
	(173.70, 97.46) --
	(168.42, 97.46) --
	cycle;

\path[fill=fillColor] (180.99,102.12) --
	(183.64, 97.54) --
	(178.35, 97.54) --
	cycle;

\path[fill=fillColor] (190.92,102.29) --
	(193.57, 97.71) --
	(188.28, 97.71) --
	cycle;

\path[fill=fillColor] (200.85,102.70) --
	(203.50, 98.12) --
	(198.21, 98.12) --
	cycle;

\path[fill=fillColor] (210.79,103.72) --
	(213.43, 99.15) --
	(208.14, 99.15) --
	cycle;

\path[fill=fillColor] (220.72,105.94) --
	(223.36,101.36) --
	(218.07,101.36) --
	cycle;

\path[fill=fillColor] (230.65,110.72) --
	(233.29,106.14) --
	(228.01,106.14) --
	cycle;

\path[fill=fillColor] (240.58,119.32) --
	(243.22,114.75) --
	(237.94,114.75) --
	cycle;

\path[fill=fillColor] (250.51,130.63) --
	(253.15,126.06) --
	(247.87,126.06) --
	cycle;

\path[fill=fillColor] (260.44,146.28) --
	(263.08,141.70) --
	(257.80,141.70) --
	cycle;

\path[fill=fillColor] (270.37,165.38) --
	(273.01,160.80) --
	(267.73,160.80) --
	cycle;

\path[fill=fillColor] (280.30,182.78) --
	(282.95,178.21) --
	(277.66,178.21) --
	cycle;

\path[fill=fillColor] (290.23,201.50) --
	(292.88,196.92) --
	(287.59,196.92) --
	cycle;
\definecolor{drawColor}{gray}{0.70}

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 32.03, 57.56) -- ( 32.03, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 41.96, 57.56) -- ( 41.96, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 51.89, 57.56) -- ( 51.89, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 61.82, 57.56) -- ( 61.82, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 71.75, 57.56) -- ( 71.75, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 81.68, 57.56) -- ( 81.68, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 91.61, 57.56) -- ( 91.61, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (101.55, 57.56) -- (101.55, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (111.48, 57.56) -- (111.48, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (121.41, 57.56) -- (121.41, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (131.34, 57.56) -- (131.34, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (141.27, 57.56) -- (141.27, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (151.20, 57.56) -- (151.20, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (161.13, 57.56) -- (161.13, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (171.06, 57.56) -- (171.06, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (180.99, 57.56) -- (180.99, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (190.92, 57.56) -- (190.92, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (200.85, 57.56) -- (200.85, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (210.79, 57.56) -- (210.79, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (220.72, 57.56) -- (220.72, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (230.65, 57.56) -- (230.65, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (240.58, 57.56) -- (240.58, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (250.51, 57.56) -- (250.51, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (260.44, 57.56) -- (260.44, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (270.37, 57.56) -- (270.37, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (280.30, 57.56) -- (280.30, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (290.23, 57.56) -- (290.23, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] (300.16, 57.56) -- (300.16, 63.25);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 29.55, 65.04) -- ( 35.24, 65.04);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 29.55, 80.18) -- ( 35.24, 80.18);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 29.55, 95.33) -- ( 35.24, 95.33);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 29.55,110.48) -- ( 35.24,110.48);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 29.55,125.62) -- ( 35.24,125.62);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 29.55,140.77) -- ( 35.24,140.77);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 29.55,155.92) -- ( 35.24,155.92);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 29.55,171.06) -- ( 35.24,171.06);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 29.55,186.21) -- ( 35.24,186.21);

\path[draw=drawColor,line width= 0.6pt,line join=round,line cap=round] ( 29.55,201.36) -- ( 35.24,201.36);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 25.50, 92.85) {2};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 25.50,138.29) {16};

\node[text=drawColor,anchor=base east,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 25.50,183.73) {128};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 27.30, 95.33) --
	( 29.55, 95.33);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 27.30,140.77) --
	( 29.55,140.77);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 27.30,186.21) --
	( 29.55,186.21);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{gray}{0.20}

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 61.82, 55.31) --
	( 61.82, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] ( 91.61, 55.31) --
	( 91.61, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (121.41, 55.31) --
	(121.41, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (151.20, 55.31) --
	(151.20, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (180.99, 55.31) --
	(180.99, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (210.79, 55.31) --
	(210.79, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (240.58, 55.31) --
	(240.58, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (270.37, 55.31) --
	(270.37, 57.56);

\path[draw=drawColor,line width= 0.5pt,line join=round] (300.16, 55.31) --
	(300.16, 57.56);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{gray}{0.30}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 61.82, 48.55) {$2^{3}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 91.61, 48.55) {$2^{6}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (121.41, 48.55) {$2^{9}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (151.20, 48.55) {$2^{12}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (180.99, 48.55) {$2^{15}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (210.79, 48.55) {$2^{18}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (240.58, 48.55) {$2^{21}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (270.37, 48.55) {$2^{24}$};

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.72] at (300.16, 48.55) {$2^{27}$};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at (166.10, 38.70) {Domain size};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,rotate= 90.00,anchor=base,inner sep=0pt, outer sep=0pt, scale=  0.90] at ( 10.70,134.94) {Total runtime (s)};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{fillColor}{RGB}{255,255,255}

\path[fill=fillColor] ( 23.41,  4.50) rectangle (308.78, 27.95);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] ( 32.41,  9.00) rectangle ( 46.87, 23.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{RGB}{27,158,119}

\path[draw=drawColor,line width= 0.6pt,line join=round] ( 33.86, 16.23) -- ( 45.42, 16.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{fillColor}{RGB}{27,158,119}

\path[fill=fillColor] ( 39.64, 16.23) circle (  1.96);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (100.48,  9.00) rectangle (114.93, 23.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{RGB}{217,95,2}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 2pt off 2pt ,line join=round] (101.93, 16.23) -- (113.49, 16.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{fillColor}{RGB}{217,95,2}

\path[fill=fillColor] (107.71, 19.28) --
	(110.35, 14.70) --
	(105.06, 14.70) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (180.81,  9.00) rectangle (195.26, 23.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{RGB}{117,112,179}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 2pt ,line join=round] (182.25, 16.23) -- (193.82, 16.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{fillColor}{RGB}{117,112,179}

\path[fill=fillColor] (186.07, 14.26) --
	(190.00, 14.26) --
	(190.00, 18.19) --
	(186.07, 18.19) --
	cycle;
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{fillColor}{gray}{0.92}

\path[fill=fillColor] (252.25,  9.00) rectangle (266.70, 23.45);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.6pt,dash pattern=on 4pt off 4pt ,line join=round] (253.69, 16.23) -- (265.26, 16.23);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{RGB}{231,41,138}

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (256.70, 16.23) -- (262.25, 16.23);

\path[draw=drawColor,line width= 0.4pt,line join=round,line cap=round] (259.48, 13.45) -- (259.48, 19.00);
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at ( 51.37, 13.75) {\textsc{Crane2-BFS}};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at (119.43, 13.75) {\textsc{Crane2-Greedy}};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at (199.76, 13.75) {\textsc{FastWFOMC}};
\end{scope}
\begin{scope}
\path[clip] (  0.00,  0.00) rectangle (307.15,216.81);
\definecolor{drawColor}{RGB}{0,0,0}

\node[text=drawColor,anchor=base west,inner sep=0pt, outer sep=0pt, scale=  0.72] at (271.20, 13.75) {\textsc{ForcLift}};
\end{scope}
\end{tikzpicture}

\end{document}